### 2023-11-02 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@stefanbringuier}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2023-11-03 src/tex/ms.tex
Environments, Rules, & Scripts for EOS. Works, produces plot.

@@
 \usepackage{showyourwork}
+\usepackage[version=4]{mhchem}
 
@@
 % Title
-\title{An open source scientific article}
+\title{Assessing Phonon Properties of Shape Memory Alloys using Graph Neural Network Potentials}
 
@@
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
+
 \end{abstract}
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+\section{Results}
+\label{sec:results}
+\subsection{Equation of States}
+\label{subsec:eos}
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+The thermodynamic equation of state (EOS) is a standard assement of the phase stability of materials at a give condition.
+In the context of atomic calculations the equation of state can be determined at ground-state (i.e., 0K) to asses the phase stability of the different crystals.
+The equation of state can be represented as a pressure vs.
+volume or more commonly as the energy per atom  vs the ratio of the isotropically strained volume to that of the equilibrium volume.
+One criteria for a interatomic potential is that it accurately refelcts the EOS curves for different phases.
+This entails the construction of a convex hull which represents the minimum energy curve accross the structures for the volume strain conditions.
+\par
+
+The EOS curves for the \ce{NiTi} structures \textit{B2}, \textit{B19}, \textit{B19'}, \textit{B33}, and \textit{Pbcm}  for the different potentials studied in this work are shown in fig.~\ref{fig:eos}
+\begin{figure}[ht!]
+    \script{PlotNiTiEOS.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/NiTi_EOS_Comparison.png}
+        \caption{
+            The equation of state curves for different \ce{NiTi} structures for (a) Mutter (EAM) (b) Zhong (EAM) (c) M3GNet (d) CHGNet (e) MACE and (f) ALIGNN. 
+        }
+        \label{fig:eos}
+    \end{centering}
+\end{figure}
+
+\appendix
+\section{Reproducibilty}
+This study was carried out using the reproducibility software \href{https://github.com/showyourwork/showyourwork}{\showyourwork} \citep{Luger2021}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}, create the figures, and compile the manuscript.
+Each figure caption contains two links: one to the dataset stored on zenodo used in the corresponding figure, and the other to the script used to make the figure (at the commit corresponding to the current build of the manuscript).
+The git repository associated to this study is publicly available at \url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP} and allows anyone to re-build the entire manuscript.
+The datasets generated by this paper are stored at \url{xyz}.
 

### 2023-11-03 src/scripts/GenerateEOS.py

@@ -1,0 +1,156 @@
+import argparse
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+#from pathlib import Path
+import paths
+
+
+
+from ase.optimize import BFGS, FIRE, GPMin
+from ase.db import connect
+from ase.spacegroup import crystal
+from ase.calculators.eam import EAM
+from ase.units import kJ
+from ase.eos import EquationOfState
+from ase.constraints import StrainFilter, UnitCellFilter
+from ase.spacegroup.symmetrize import FixSymmetry, check_symmetry
+
+from NiTiStructures import *
+from Calculators import *
+
+
+def calc_EOS(structure,
+             potential,
+             of,
+             dbname="NiTi_EOS.db",
+             strain=np.linspace(-0.10,0.10,100),
+             eos_eq='sj',
+             fmax=5.0e-3,
+             ):
+    """
+    Calculate the equation of state for different potentials on a specified structure. 
+    The structure is strained using the StrainFilter class from the ASE.
+    """
+
+
+    potname = potential[0]
+    structure.calc =  potential[1]
+    vol_intial = structure.get_cell().volume
+    
+    db = connect(dbname)
+
+    sf = StrainFilter(structure)
+    energies = []
+    volumes = []
+    for e in strain:
+        isotropic_strain = np.array([e,e,e, 0, 0, 0])
+        sf.set_positions(isotropic_strain)
+        energy = sf.get_potential_energy()
+        volume = sf.get_cell().volume
+        energies.append(energy)
+        volumes.append(volume)
+        
+    eos = EquationOfState(volumes, energies,eos=eos_eq)
+    v0, e0, B = eos.fit()
+
+
+        
+    # Plot and save EOS
+    fig = plt.figure(1, figsize=(7, 4))
+    ax = fig.add_subplot(1, 1, 1)
+    structure_name = structure.info["structure_name"]
+    eos.plot(ax=ax, filename=os.path.join(of,f'{potname}_NiTi_{structure_name}_EOS_{eos_eq}.png'))
+    fig.clear()
+    plt.close(fig)  
+
+
+
+    # Now find the ground-state structure using the StrainFilter and BFGS optimization
+    # Then store the ground-state optimal structure and the results for the EOS fit.
+    structure_copy = structure.copy()
+    structure_copy.calc = potential[1]
+
+    # Apply SymmetryFilter and UnitCellFilter for optimization
[…]

### 2023-11-03 src/scripts/NiTiStructures.py

@@ -1,0 +1,158 @@
+from ase.spacegroup import crystal
+
+def NiTi_B2_Structure(
+    a=3.04, b=3.04, c=3.04, alpha=90.0, beta=90.0, gamma=90.0
+):
+    """
+    Austenite phase.
+    """
+    B2 = crystal(
+        ["Ni", "Ti"],
+        [(0, 0, 0), (0.5, 0.5, 0.5)],
+        spacegroup=221,
+        cellpar=[a, b, c, alpha, beta, gamma],
+        pbc=True,
+    )
+
+    return B2
+
+
+def NiTi_B33_Structure(
+    a=2.888, b=9.398, c=4.042, alpha=90.0, beta=90.0, gamma=90.0
+):
+    """
+    Structure from: Liu et al., Phys. Rev. B 87, 140104(R) 2018. 
+    Cmcm (B33/BCO) @ 0 GPa
+    """
+    B33 = crystal(
+        ["Ni", "Ti"],
+        [(0.0,0.5847,0.25),(0.0,0.1440,0.75)],
+        cellpar=[a, b, c, alpha, beta, gamma],
+        pbc=True,
+        spacegroup=63,
+    )
+    return B33
+
+def NiTi_Cmcm_Structure(
+        a=2.90, b=3.99, c=4.90, alpha = 90.0, beta=107.01, gamma=90
+        ):
+    Cmcm = crystal(
+        ["Ni", "Ti"],
+        [(0.915362,0.75,0.829275),(0.641384,0.75,0.284766)],
+        cellpar=[a, b, c, alpha, beta, gamma],
+        pbc=True,
+        spacegroup=11, #Should be 63 according to mp data
+    )
+    return Cmcm
+    
+
+def NiTi_B19P_Structure(
+    a=2.89, b=3.97, c=4.83, alpha=90.0, beta=105.23, gamma=90.0
+):
+    """
+    From Comp. Mat Sci S. Gur paper
+    """
+    B19P = crystal(
+        ["Ni", "Ti"],
+        [(0.076622, 0.25, 0.671102), (0.369548, 0.25, 0.217074)],
+        cellpar=[a, b, c, alpha, beta, gamma],
+        pbc=True,
+        spacegroup=11,
+    )
+    return B19P
+
+def NiTi_B19_Structure(
+    a=2.8, b=4.17, c=4.67, alpha=90.0, beta=95.55, gamma=90.0
+):
+    """
+    Structure: https://materialsproject.org/materials/mp-1179013
+    """
+    B19 = crystal(
[…]

### 2023-11-03 src/scripts/PlotNiTiEOS.py

@@ -1,0 +1,106 @@
+
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+from matplotlib.ticker import MultipleLocator, AutoLocator
+from cycler import cycler
+import seaborn as sns  # For color palette
+from ase.db import connect
+import numpy as np
+import os
+import paths
+
+SUBPLOT_ORDER = {'a': 'Mutter',
+                 'b': 'Zhong',
+                 'c': 'M3GNet',
+                 'd': 'CHGNet',
+                 'e': 'MACE',
+                 'f': 'ALIGNN',
+                 }
+
+def plot_eos(db_path, figsize=(3.4, 5.75), dpi=600, xlim=(0.725, 1.325), ignore_list=['B32', 'R_Phase'], 
+             color_palette="muted", n_colors=10, wspace=0.3, hspace=0.3, linewidth=1, linestyle='-', 
+             fontsize=8, tick_fontsize=6, legend_fontsize=8, legend_ncol=3):
+   
+    # Connect to the ASE database
+    db = connect(db_path)
+    
+    # Gather all unique calculator models and structure names
+#    calculator_models = set(row.model_name for row in db.select())
+
+    calculator_models = sorted(set(row.model_name for row in db.select()),
+                               key=lambda model: list(SUBPLOT_ORDER.values()).index(model))
+
+    structure_names = set(row.structure_name for row in db.select())
+
+    
+    # Create a figure with a 2x3 grid of subplots
+    fig, axs = plt.subplots(3, 2, figsize=figsize)
+    axs = axs.ravel()  # Flatten the array of axes for easier indexing
+
+    # Create a color cycle iterator using a slightly modified color palette
+    color_cycle = sns.color_palette(color_palette, n_colors=n_colors)
+    
+    label_file_path = paths.figures / 'NiTi_EOS_Comparison_label.txt'
+    with open(label_file_path, 'w') as label_file:
+        # Iterate over calculator models
+        for i, calculator_model in enumerate(calculator_models):
+
+            if calculator_model not in SUBPLOT_ORDER.values():
+                raise ValueError(f"{calculator_model} model is not expected")
+            
+            ax = axs[i]  # Select the current axis
+            alphabetic_label = f'({chr(97 + i)})'
+            ax.set_title(alphabetic_label, loc='left', fontsize=fontsize)  # Set title with alphabetic label
+            label_file.write(f'{alphabetic_label}: {calculator_model}\n')  # Write label to file
+            
[…]

### 2023-11-03 src/scripts/SetMetadata.py

@@ -1,0 +1,44 @@
+import paths
+from ase.db import connect
+
+def set_db_metadata(dbname):
+    db = connect(dbname, append=True)
+    db.metadata = {
+        "title": "NiTi Equation of State Calculations",
+        "key_descriptions": {
+            "vi": ("Initial Volume", "Initial volume of structure", "Å^3"),
+            "v0": ("Equilibrium Volume", "Equilibrium volume from EOS fit", "Å^3"),
+            "e0": ("Equilibrium Energy", "Equilibrium energy from EOS fit", "eV"),
+            "bulk_modulus": ("Bulk Modulus", "Bulk modulus from EOS fit", "GPa"),
+            "fit_eq": ("Fit Equation", "Equation used for the EOS fit", "string"),
+            "strain": ("Strain", "Isotropic Strain", "fractional value"),
+            "potential": ("Potential", "Potential used in the calculation", "string"),
+            "spacegroup": ("Spacgroup #", "International spacegroup number", "number"),
+            "ecoh" :("Cohesive energy", "Cohesive energy per atom", "eV/atom"),
+            "structure_name":("Structure Name", "Prototype structure name", "string"),
+            "model_name": ("Model Name", "Name of the model/parameters", "string"),
+        },
+        "default_columns": [
+            "id",
+            "calculator",
+            "formula",
+            "potential",
+            "vi",
+            "v0",
+            "e0",
+            "ecoh",
+            "bulk_modulus",
+            "fit_eq",
+            "strain",
+            "volume",
+            "structure_name",
+            "model_name",
+            "spacegroup",
+        ],
+    }
+    return None
+
+
+if __name__ == "__main__":
+    db_path_file = paths.data / "NiTi_EOS.json"
+    set_db_metadata(db_path_file)  

### 2023-11-23 src/tex/ms.tex
Workflow runs as intended. Figure scripts started. Some text added to manuscript draft.

@@
 % Define document class
-\documentclass[twocolumn]{aastex631}
+\documentclass[preprint]{elsarticle}
 \usepackage{showyourwork}
+
+%% Packages
+\usepackage[utf8]{inputenc}
+\usepackage{fontspec}
+\usepackage{amsmath}
+\usepackage{siunitx}
 \usepackage[version=4]{mhchem}
+\usepackage{pdflscape}
 
@@
 % Author list
-\author{@stefanbringuier}
+\author[1]{Stefan Bringuier\corref{cor1}}
+\ead{stefanbringuier@gmail.com}
+\cortext[cor1]{Corresponding author}
+\affiliation[1]{organization={Indpendent Researcher},
+  city={San Diego, CA},
+  country={United States},
+}
 
@@
 \begin{abstract}
+Shape Memory Alloys (SMAs) such as NiTi are pivotal in vibrational dampening,robotic arms, prothestic hands, and mechanical vales as a result of their unique displacive phase transformation properties.
+However, assessing their phonon characteristics, which provide insight into the displacive phase transformation, by using conventional Density Functional Theory (DFT) methods is computationally challenging.
+In this paper we assess the use of recent Graph Neural Network (GNN) potentials as a viable alternative.
+These GNN potentials have leveraged large DFT-computed databases for training and provide a good approximation to the potential energy surfaces which enable inference on energies and forces in lattice and atomic dynamic calculations.
+We compare GNN potentials with the Embedded Atom Method (EAM) in predicting phonon properties of NiTi.
+We find that in general the EAM potentials are slightly more performant than the GN in that they appear to more accurately describe equation of state, phonon dispersions, and elastic constants.
+However, the MACE potential shows significant promise as an exploritory potential for designing and characterizing more complex SMAs.
+Our findings reveal GNN potentials as a promising avenue for advancing SMA research and fine-tuning of pre-trained GNN models with more relevant DFT data (i.e., force-constants) may permit better outcomes.
+\end{abstract}
 
-\end{abstract}
+\begin{keyword}
+  shape memory alloys \sep machine learning potentials \sep phonons
+\end{keyword}
+
+\maketitle
 
@@
 
+\section{Computation Methodology}
+\label{sec:methods}
+
+\begin{itemize}
+\item Discuss EAM potentials Mutter and Zhong.
+\item GNN potentials M3GNet, CHGNet, MACE, ALIGNN
+\item EOS calcs
+\item Phonon calcs with strain.
+
+  The calculation of the phonon dispersion is based on the dynamical matrix which in term is obtained from the fourier transform of the force-constant matrix.
+There are different numerical methods to obtain the force-constant matrix.
+Here we employ the finite-difference small displacement method (cite alfe), whereby atoms in a supercell are displaced by a small amount to ensure they remain within the harmonic (i.e., linear-response) regime.
+Then the force-constants are calculated.
+To obtain the dynamical matrix we sample reciprocal space on a given mesh within the irreducible Brillouin zone (BZ).
+Finally, we diagonalize the dynamical matrix along high-symmetry paths to obtain the phonon bands and eigenfrequencies.
+\par
+
+  Since the the finite-difference is a numerical approximation technique, one has to check convergence of the calculation for various models.
+In this work we have found that for most models and the avaible resources a supercell of $8\times8\times8$ is sufficient to converge, we also utlize the acoustic sum rule (add ref).
+The only stand-out for using that supercell is with use of the ALIGNN model, we found that supercells any larger than $5\times5\times5$ had considerable memory and we the required resources were unavilable.
+Despite this we find suitable convergence, althoug the ALIGNN model in general was not good at reproducing the NiTi phonon bandstructures.
+\par
+
+  The displacement was choosen for all calculations to be $0.01\AA$ and the number of points along the BZ path in reciprocal space was 150.
+These values seemed to give the best trade-off of convergence for all the models.
+We also calculate the phonon does on a $30\times30\time30$ reciprocal space grid and use lorentz smoothing, the DOS can be found in the supplemental material.\par
+
+\item elastic constants.
+\end{itemize}
+
 \section{Results}
 \label{sec:results}
-\subsection{Equation of States}
-\label{subsec:eos}
 
+\subsection{NiTi: Ground State Structures}
+\label{subsec:nitigs}
+
+%\comment{Display results in tables}
+
+\variable{output/Table_NiTi_Equilibrium_Structures.tex}
+
+\subsection{NiTi: Equation of States}
+\label{subsec:nitieos}
+\cite{Luger2021}
 The thermodynamic equation of state (EOS) is a standard assement of the phase stability of materials at a give condition.
@@
 
-The EOS curves for the \ce{NiTi} structures \textit{B2}, \textit{B19}, \textit{B19'}, \textit{B33}, and \textit{Pbcm}  for the different potentials studied in this work are shown in fig.~\ref{fig:eos}
+The EOS curves for the \ce{NiTi} structures \textit{B2}, \textit{B19}, \textit{B19'}, \textit{B33}, and \textit{Pbcm}  for the different potentials studied in this work are shown in fig.~\ref{fig:eos}.
+The two EAM potentials, show that at 0 strain ($V/V_o = 1.0$) both the \textit{B19P} and \textit{B33} phases are stable.
+I believe this is the first reporting of the \textit{B33} phase for these two fittings of the EAM potential for \ce{NiTi}.
+Its important to note that these potentials have been developed to capture the displacive phase transformation between \textit{B19P} to \textit{B2}.
+This is shown in fig.~\ref{fig:eos}(a,b) with the stability indicated upon compressive isotropic strain.
+A mininum appears around $\approx 0.9$.
+At tensile isotropic strains the \textit{B19P} phase is stable up till around $1.18$ where the \textit{Pbcm} phase shows stability.
+An interesting note is that these two fittings of the EAM potential show a double well curve for the \textit{B2} phase.
+\par
+
+In the UIP-GNN potentials M3GNet, CHGNet, and MACE, the equation of states shown in fig.~\ref{fig:eos}(c-e) look fairly different from thos in fig.~\ref{fig:eos}(a,b).
+For the M3GNet,  at no strain, the \textit{B33} appears to be the predicted stable phase and the \textit{B2} phase becomes stable around $0.9$ (see fig.~\ref{fig:eos}(c).
+The stability for \textit{B19} and \textit{B19P} exit, in trade-off, between approximately $1.025$ and $1.1$ tensile isotropic strain.
+Interestingly beyond a strain of $1.1$ the \textit{Pbcm} is dominate.
+The double-well curve shown for the EAM potentials is not present for the B2-phase.
+The CHGNet potential results shown in fig.~\ref{fig:eos}(d) are similar to the M3GNet, but the stability window of the \textit{B2} occurs at a lower compressive isotropic strain.
+\par
+
+The MACE potential shown in fig.~\ref{fig:eos}(e) has a very different set of EOS curves compard to the M3GNet and CHGNet.
+Here many of the phases exhibit a stablity window, except the \textit{B33} phase, which was stable in M3GNet and CHGNet.
+Interestingly, the \text{B19} phase, not the \text{B19P} phase is shown to be the stable zero strain phase.
+One unique perspective of the the UIP-GNN models is that they provide a different displacive phase transformation pathway then that shown by the EAM potentials.
+However, this does not include the meta-stable R-phase or other potential meta-stable phases which can also offer such transformation mechanisms.
+\par
+
+Finally, the ALIGNN potential EOS was generated, but a few issues arose during the calculations.
+The most prominant was the difficulty in convergence of the structural optimzation to the specified tolerance (see \ref{sec:methods}.
+Secondly the calculated cohesive energies differs from the UIP-GNN by a factor $\approx 2$ and by $1$ eV for the EAM potentials.
+The \textit{B19P} phase is the stable at zero strain and for a considerable portion of the compressive, but all of the tensile.
+At compressive strain beyond $\approx 0.94$ the \textit{B2} phase is stable.
+
+It is observed that for all but the ALIGNN potential, that upon high tensile isotropic strain, the \textit{Pbcm} phase is the most stable. \par
+
 \begin{figure}[ht!]
@@
         \caption{
-            The equation of state curves for different \ce{NiTi} structures for (a) Mutter (EAM) (b) Zhong (EAM) (c) M3GNet (d) CHGNet (e) MACE and (f) ALIGNN. 
+            The equation of state curves for different \ce{NiTi} structures for (a) Mutter (EAM) (b) Zhong (EAM) (c) M3GNet (d) CHGNet (e) MACE and (f) ALIGNN interatomic potentials.
+The x-axis is the ratio of the isotropic strained unit cell volume, $V$,  with respect to the ground-state unit cell volume, $V_o$.
         }
@@
 
-\appendix
-\section{Reproducibilty}
-This study was carried out using the reproducibility software \href{https://github.com/showyourwork/showyourwork}{\showyourwork} \citep{Luger2021}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}, create the figures, and compile the manuscript.
+\subsection{NiTi: Phonon Instability}
+\label{subsec:niphonons}
+
+\subsubsection{B2 Phase}
+\label{subsubsec:b2}
+
+\begin{figure}[ht!]
+    \script{PlotNoStrainPhonons.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/B2_combined_models_phonon_bandstructures.png}
+        \caption{
+          Ground-state, no isotropic strain, Phonon bandstructure of B2 phase for all models.
+The Mutter and Zhong EAM potentials differ in that Mutter predicts no dynamic instability of the B2 phase at 0K.
+        }
+        \label{fig:allmodels_no_strainphonon_b2}
+    \end{centering}
+\end{figure}
+
+\begin{figure}[ht!]
+    \script{PlotEAMPhonons.py}
+    \begin{centering}
+        \includegraphics[width=0.5\linewidth]{figures/Mutter_B2_combined_strain_phonon_bandstructures.png}
+        \vspace{1mm}
+        \includegraphics[width=0.5\linewidth]{figures/Zhong_B2_combined_strain_phonon_bandstructures.png}
+        \caption{
+           Phonon dispersion of B2 structure for the Mutter (top) and Zhong (bottom)  EAM potentials at different isotropic strains.
+The optical branches have been made more transparent to emphasis the behavior in the TA$_1$ and TA$_2$ branches.
+The color indicates direction from compressive to tensile strain.
+Note that the strain percent shown is with respect to isotropic changes (i.e.
+volume).
+        }
+        \label{fig:mutter_zhong_phonon_b2}
+    \end{centering}
+\end{figure}
+
+%ADD Plot of zero strain EAM and first-pricinpals results
+
+\begin{figure}[ht!]
+    \script{PlotGNNPhonons.py}
+    \begin{centering}
+      \includegraphics[width=0.5\linewidth]{figures/M3GNet_B2_combined_strain_phonon_bandstructures.png}
+      \vspace{1mm}
+      \includegraphics[width=0.5\linewidth]{figures/CHGNet_B2_combined_strain_phonon_bandstructures.png}
+      \vspace{1mm}
+      \includegraphics[width=0.5\linewidth]{figures/MACE_B2_combined_strain_phonon_bandstructures.png}
+      \caption{
+        Phonon dispersion of B2 structure for the (a) M3GNet, (b) CHGNet, and (c) MACE potentials for different isotropic strains.
+Similar transcparency is used as done in \ref{fig:mutter_zhong_phonon_b2}.
+      }
+      \label{fig:gnn_phonon_b2}
+    \end{centering}
+\end{figure}
+
+\subsection{NiTiCu: Equation of States}
+
+\bibliographystyle{naturemag}
+\bibliography{bib}
+
+\section*{CR\lowercase{e}d\lowercase{i}T author statement}
+
+S.B: Conceptualization, Formal Analysis, Investigation, Methodology, Software, Visualisation, Writing---original draft.
+
+\section*{Reproducibilty and Data Availability}
+This study was carried out using the reproducibility software \href{https://github.com/showyourwork/showyourwork}{\showyourwork} \cite{Luger2021}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}, create the figures, and compile the manuscript.
 Each figure caption contains two links: one to the dataset stored on zenodo used in the corresponding figure, and the other to the script used to make the figure (at the commit corresponding to the current build of the manuscript).
@@
 
-\bibliography{bib}
+\section*{Acknowledgements}
+
+\appendix
 

### 2023-11-23 src/scripts/CalculateEOS.py

@@ -1,0 +1,134 @@
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+from ase.optimize import FIRE
+from ase.db import connect
+from ase.units import kJ
+from ase.eos import EquationOfState
+from ase.constraints import FixAtoms
+from ase.constraints import StrainFilter, UnitCellFilter
+from ase.spacegroup.symmetrize import FixSymmetry, check_symmetry
+
+
+import copy
+
+
+def to_scalar(value):
+    if isinstance(value, np.ndarray) and value.size == 1:
+        return value.item()
+    else:
+        return value
+
+
+def CalculateEOS(structure,
+             potential,
+             of,
+             dbname="NiTi_EOS.db",
+             strain=np.linspace(-0.10,0.10,100),
+             eos_eq='sj',
+             fmax=2.0e-3,
+             fstep=5000,
+             ):
+    """
+    Calculate the equation of state for different potentials on a specified structure. 
+    The structure is strained using the StrainFilter class from the ASE.
+    
+    TODO: Need to handle the sensitivities of the LAMMPSlib calc
+    """
+    db = connect(dbname)
+#    db.serial = True
+    potname = potential[0]
+    structure.set_calculator(potential[1])
+    vol_intial = structure.get_cell().volume
+
+    # Store the symmetry
+    sym_before = check_symmetry(structure,1.0e-6, verbose=False)
+    # Apply SymmetryFilter and UnitCellFilter for optimization
+    structure.set_constraint(FixSymmetry(structure,symprec=1.0e-4))
+    ucf_sym = UnitCellFilter(structure)
+    # FIRE is used as best performance seen.
+    opt = FIRE(ucf_sym)
+    opt.run(fmax=fmax,steps=fstep)
+    #ISSUE: Need to turn of constraints for ASE db
+    structure.set_constraint(None)
+    etotal = structure.get_total_energy()
+    stress = structure.get_stress()
+    forces = structure.get_forces()
+    ecoh = structure.get_potential_energy() / structure.get_global_number_of_atoms()
+    nextrow = db.write(structure,
+                       relaxed=True,
+                       vi=vol_intial,
+                       ecoh = ecoh,
+                       spacegroup=sym_before["number"],
+                       structure_name=structure.info['structure_name'],
+                       model_name = structure.info['model_name'],
+                       )
+
+             
+
[…]

### 2023-11-23 src/scripts/CalculatePhonons.py

@@ -1,0 +1,265 @@
+import argparse
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+import paths
+
+from ase.phonons import Phonons
+from ase.db.core import Database, connect
+from ase.spacegroup import get_spacegroup
+from ase.spacegroup import crystal
+from ase.io.trajectory import Trajectory
+from ase.io import write
+from ase.constraints import StrainFilter, UnitCellFilter
+from ase.spacegroup.symmetrize import FixSymmetry, check_symmetry
+
+
+EV_to_THz = 241.799050402293e0
+
+
+def PlotBandstructure(bs, dos, info, ymaxlim=13.0, of="."):
+    """
+    Plots a single phonon band structure and DOS.
+    """
+    fig = plt.figure(1, figsize=(7, 4))
+    ax = fig.add_axes([0.12, 0.07, 0.67, 0.85])
+    dosax = fig.add_axes([0.8, 0.07, 0.17, 0.85])
+
+    xticks = bs.get_labels()[1].tolist()
+    xlabels = bs.get_labels()[2]
+
+    for xt in xticks:
+        ax.axvline(x=xt, color="lightgray")
+    ax.axhline(y=0.0, color="black")
+
+    q_points = bs.get_labels()[0]
+    energies_THz = EV_to_THz * bs.energies
+    for n in range(len(energies_THz[0, 1, :])):
+        omega = energies_THz[0, :, n]
+        ax.plot(q_points, omega, "k-", lw=2)
+
+    ax.set_title(
+        f"Lattice Parameter: {info['a']:.3f}Å\
+        Strain: {info['strain_percent']:.2f}% (a0={info['a0']:.3f}Å)\
+        Potential: {info['potname']}",
+        fontsize=10,
+    )
+
+    ax.set_xticks(xticks)
+    ax.set_xticklabels(xlabels)
+
+    ax.set_xlim((0.0, q_points[-1]))
+    ax.set_ylim((-0.5, ymaxlim))
+    ax.set_ylabel("Frequency ($\mathrm{THz}$)", fontsize=22)
+
+    dosax.fill_between(
+        dos.get_weights(),
+        EV_to_THz * dos.get_energies(),
+        y2=0,
+        color="grey",
+        edgecolor="k",
+        lw=1,
+    )
+
+    dosax.set_ylim((0, ymaxlim))
+    dosax.set_yticks([])
+    dosax.set_xticks([])
[…]

### 2023-11-23 src/scripts/Calculators.py

@@ -2,38 +2,52 @@ import paths
 import numpy as np
 from ase.calculators.eam import EAM
 
-
+# No longer needed using conda LAMMPS, which provides the needed stress.
 class CustomEAM(EAM):
-    '''
-    Override EAM to zero out stress for ASE Database support
-    '''
-    def __init__(self, **kwargs):
-        super().__init__(**kwargs)
-    
-    def get_stress(self, atoms=None, **kwargs):
-        return np.zeros(6)
+   '''
+   Override EAM to zero out stress for ASE Database support
+   '''
+   def __init__(self, **kwargs):
+       super().__init__(**kwargs)
+   
+   def get_stress(self, atoms=None, **kwargs):
+       return np.zeros(6)
 
 def GetCalculator(args):
     if args.model == 'Mutter':
-        asecalc = CustomEAM(potential= str(paths.static / "NiTi_mutter.eam.fs"))
+       potential_file = str(paths.static / "NiTi_mutter.eam.fs")
+       #asecalc = CustomEAM(potential= str(paths.static / "NiTi_mutter.eam.fs"))
+       from ase.calculators.lammpslib import LAMMPSlib 
+       cmds = ["pair_style eam/fs",f"pair_coeff * *  {potential_file} Ni Ti"]
+       amds=["thermo_style custom etotal lx ly lz vol pxx pyy pzz pxy pxz pyz press"]
+       asecalc = LAMMPSlib(lmpcmds=cmds,amendments=amds,keep_alive=True)
+       #from ase.calculators.lammpsrun import LAMMPS
+       #parameters = {'pair_style': 'eam/fs','pair_coeff': [f'* * {potential_file} Ni Ti']}
+       #files = [potential_file]
+       #asecalc = LAMMPS(parameters=parameters,files=files)
     elif args.model == 'Zhong':
-        asecalc = CustomEAM(potential= str(paths.static / "NiTiZhong.eam.fs"))
+       potential_file = str(paths.static / "NiTiZhong.eam.fs")
+       #asecalc = CustomEAM(potential= str(paths.static / "NiTiZhong.eam.fs"))
+       from ase.calculators.lammpslib import LAMMPSlib
+       cmds = ["pair_style eam/fs",f"pair_coeff * *  {potential_file} Ni Ti"]
+       amds=["thermo_style custom etotal lx ly lz vol pxx pyy pzz pxy pxz pyz press"]
[…]

### 2023-11-23 src/scripts/GenerateEOS.py

@@ -1,156 +1,0 @@
-import argparse
-import os
-import numpy as np
-import matplotlib.pyplot as plt
-#from pathlib import Path
-import paths
-
-
-
-from ase.optimize import BFGS, FIRE, GPMin
-from ase.db import connect
-from ase.spacegroup import crystal
-from ase.calculators.eam import EAM
-from ase.units import kJ
-from ase.eos import EquationOfState
-from ase.constraints import StrainFilter, UnitCellFilter
-from ase.spacegroup.symmetrize import FixSymmetry, check_symmetry
-
-from NiTiStructures import *
-from Calculators import *
-
-
-def calc_EOS(structure,
-             potential,
-             of,
-             dbname="NiTi_EOS.db",
-             strain=np.linspace(-0.10,0.10,100),
-             eos_eq='sj',
-             fmax=5.0e-3,
-             ):
-    """
-    Calculate the equation of state for different potentials on a specified structure. 
-    The structure is strained using the StrainFilter class from the ASE.
-    """
-
-
-    potname = potential[0]
-    structure.calc =  potential[1]
-    vol_intial = structure.get_cell().volume
-    
-    db = connect(dbname)
-
-    sf = StrainFilter(structure)
-    energies = []
-    volumes = []
-    for e in strain:
-        isotropic_strain = np.array([e,e,e, 0, 0, 0])
-        sf.set_positions(isotropic_strain)
-        energy = sf.get_potential_energy()
-        volume = sf.get_cell().volume
-        energies.append(energy)
-        volumes.append(volume)
-        
-    eos = EquationOfState(volumes, energies,eos=eos_eq)
-    v0, e0, B = eos.fit()
-
-
-        
-    # Plot and save EOS
-    fig = plt.figure(1, figsize=(7, 4))
-    ax = fig.add_subplot(1, 1, 1)
-    structure_name = structure.info["structure_name"]
-    eos.plot(ax=ax, filename=os.path.join(of,f'{potname}_NiTi_{structure_name}_EOS_{eos_eq}.png'))
-    fig.clear()
-    plt.close(fig)  
-
-
-
-    # Now find the ground-state structure using the StrainFilter and BFGS optimization
-    # Then store the ground-state optimal structure and the results for the EOS fit.
-    structure_copy = structure.copy()
-    structure_copy.calc = potential[1]
-
-    # Apply SymmetryFilter and UnitCellFilter for optimization
[…]

### 2023-11-23 src/scripts/NewASEDatabase.py

@@ -1,0 +1,20 @@
+import sys
+import paths
+from ase.db import connect
+from SetMetadata import set_db_metadata
+
+dbname = sys.argv[1]
+db = connect(paths.data / dbname, append=False)
+
+#Need to write something otherwise it doesn't work
+from ase import Atoms
+from ase.calculators.emt import EMT
+h2 = Atoms('H2', [(0, 0, 0), (0, 0, 0.7)])
+h2.calc = EMT()
+h2.get_total_energy()
+h2.get_forces()
+db.write(h2)
+db.delete(ids=[1])
+set_db_metadata(paths.data / dbname)
+
+exit()

### 2023-11-23 src/scripts/PlotEAMPhonons.py

@@ -1,0 +1,14 @@
+import sys
+import paths
+import json
+from PlotPhonons import *
+
+
+eam_models = ["Mutter", "Zhong"]
+dbname = paths.data / sys.argv[1]
+nstrains = int(sys.argv[2])
+structures = sys.argv[3:]
+
+for m in eam_models:
+    for s in structures:
+        PlotPhonons(dbname, m, s, num_strains=nstrains)

### 2023-11-23 src/scripts/PlotGNNPhonons.py

@@ -1,0 +1,14 @@
+import sys
+import paths
+import json
+from PlotPhonons import *
+
+
+gnn_models = ["M3GNet","CHGNet","MACE"]
+dbname = paths.data / sys.argv[1]
+nstrains = int(sys.argv[2])
+structures = sys.argv[3:]
+
+for m in gnn_models:
+    for s in structures:
+        PlotPhonons(dbname, m, s, num_strains=nstrains)

### 2023-11-23 src/scripts/PlotNiTiEOS.py

@@ -1,3 +1,9 @@
+import sys
+
+import numpy as np
+import os
+import paths
+
 
 import matplotlib.pyplot as plt
 import matplotlib.ticker as ticker
@@ -5,9 +11,6 @@ from matplotlib.ticker import MultipleLocator, AutoLocator
 from cycler import cycler
 import seaborn as sns  # For color palette
 from ase.db import connect
-import numpy as np
-import os
-import paths
 
 SUBPLOT_ORDER = {'a': 'Mutter',
                  'b': 'Zhong',
@@ -17,7 +20,7 @@ SUBPLOT_ORDER = {'a': 'Mutter',
                  'f': 'ALIGNN',
                  }
 
-def plot_eos(db_path, figsize=(3.4, 5.75), dpi=600, xlim=(0.725, 1.325), ignore_list=['B32', 'R_Phase'], 
+def plot_eos(db_path, figsize=(3.4, 5.75), dpi=600, xlim=(0.75, 1.25), ignore_list=['B32', 'R_Phase'], 
              color_palette="muted", n_colors=10, wspace=0.3, hspace=0.3, linewidth=1, linestyle='-', 
              fontsize=8, tick_fontsize=6, legend_fontsize=8, legend_ncol=3):
    
@@ -83,7 +86,7 @@ def plot_eos(db_path, figsize=(3.4, 5.75), dpi=600, xlim=(0.725, 1.325), ignore_
                     ax.plot(vv_ratio, energy_per_atom, linewidth=linewidth, linestyle=linestyle, label=structure_name)
 
                     # Update the global y-axis limits
-                    ymin =  min(energy_per_atom) - 0.15
+                    ymin =  min(energy_per_atom) - 0.1
                     ymax =  max(energy_per_atom)
                     ax.set_ylim((ymin, ymax))
 
@@ -92,7 +95,7 @@ def plot_eos(db_path, figsize=(3.4, 5.75), dpi=600, xlim=(0.725, 1.325), ignore_
                    fontsize=legend_fontsize, edgecolor='black', 
                    fancybox=True)
 
-        fig.text(0.5, 0.05, 'v/v$_o$', ha='center', fontsize=fontsize)  # Single x-label
+        fig.text(0.5, 0.05, 'V/V$_o$', ha='center', fontsize=fontsize)  # Single x-label
         fig.text(0.01, 0.5, 'Energy per atom (eV)', va='center', rotation='vertical', fontsize=fontsize)  # Single y-label
[…]

### 2023-11-23 src/scripts/PlotNoStrainPhonons.py

@@ -1,0 +1,11 @@
+import sys
+import paths
+from PlotPhonons import *
+
+
+models = ["Mutter", "Zhong","M3GNet","CHGNet","MACE","ALIGNN"]
+dbname = paths.data / sys.argv[1]
+structures = sys.argv[2:]
+
+for s in structures:
+    PlotAllModelPhonons(dbname, models, s, strain=0.00)

### 2023-11-23 src/scripts/PlotPhonons.py

@@ -1,0 +1,257 @@
+import paths
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec
+from ase.db import connect
+from scipy.interpolate import interp1d
+
+
+def normalize_q_points(q_points):
+    return (q_points - np.min(q_points)) / (np.max(q_points) - np.min(q_points))
+
+
+def get_color(value, cmap_name="viridis", vmin=-0.12, vmax=0.12):
+    norm = plt.Normalize(vmin=vmin, vmax=vmax)
+    cmap = plt.get_cmap(cmap_name)
+    return cmap(norm(value))
+
+
+def PlotPhonons(dbname, model_name, structure_name, num_strains=None):
+    """
+    This function plots individual phonon bandstructures for different strains and
+    then all the strains on one plot with the optical branches semi-transparent.
+
+    TODO: confirm strain is in bars and thus conver to MPa
+    """
+    with connect(dbname) as db:
+        for row in db.select(structure_name=structure_name, model_name=model_name):
+            data = row.data
+            for strain, pdata in data["strain_phonons"].items():
+                fig, ax = plt.subplots()
+                for b_idx, energies in pdata["bandstructure"]["band_index"].items():
+                    q_points = pdata["bandstructure"]["q_points"]
+                    ax.plot(q_points, energies, color="k")
+
+                ax.set_xticks(pdata["bandstructure"]["label_locs"])
+                ax.set_xticklabels(pdata["bandstructure"]["labels"])
+                _ = [
+                    ax.axvline(x=loc, color="gray", linestyle="-", alpha=0.5)
+                    for loc in pdata["bandstructure"]["label_locs"]
+                ]
+                stress = (
+                    np.mean(pdata["stress"][:3]) * 0.1
+                )  # Confirm units are bar -> MPa
+                ax.set_title(f"Strain: {float(strain)*100:.3f}%, Stress: {stress:.3f} MPa")
+                ax.set_xlim((0, pdata["bandstructure"]["label_locs"][-1]))
+                ax.set_ylabel("Frequency [THz]")
+                ax.axhline(y=0.0, color="gray", linestyle="-", alpha=0.5)
+                fname = (
+                    model_name.upper()
+                    + "/"
+                    + f"{structure_name}_strain_{float(strain):0.3f}_phonon_bandstructure.png"
+                )
+                plt.savefig(paths.data / fname)
+                plt.close(fig)
+
[…]

### 2023-11-23 src/scripts/bck.PlotPhonons.py

@@ -1,0 +1,74 @@
+import paths
+import numpy as np
+import matplotlib.pyplot as plt
+from ase.db import connect
+
+
+def normalize_q_points(q_points):
+    return (q_points - np.min(q_points)) / (np.max(q_points) - np.min(q_points))
+
+def get_color(value, cmap_name='viridis', vmin=-0.12, vmax=0.03):
+    norm = plt.Normalize(vmin=vmin, vmax=vmax)
+    cmap = plt.get_cmap(cmap_name)
+    return cmap(norm(value))
+#def get_color(value, cmap_name='Blues', vmin=-0.11, vmax=0.01, start=0.1, stop=0.9):
+#    norm = plt.Normalize(vmin=vmin, vmax=vmax)
+#    cmap = plt.get_cmap(cmap_name)
+#    scaled_value = start + (stop - start) * norm(value)
+#    return cmap(scaled_value)
+
+def PlotPhonons(dbname, model_name, structure_name):
+    with connect(dbname) as db:
+        for row in db.select(structure_name=structure_name, model_name=model_name):
+            data = row.data
+            for strain, pdata in data['strain_phonons'].items():
+                fig, ax = plt.subplots()
+                for b_idx, energies in pdata['bandstructure']['band_index'].items():
+                    q_points = pdata['bandstructure']['q_points']
+                    ax.plot(q_points, energies, color='k')
+                
+                ax.set_xticks(pdata['bandstructure']['label_locs'])
+                ax.set_xticklabels(pdata['bandstructure']['labels'])
+                _ = [ax.axvline(x=loc, color='gray', linestyle='-', alpha=0.5) for loc in pdata['bandstructure']['label_locs']]
+                stress = np.mean(pdata["stress"][:3]) * 0.1  # Confirm units are bar -> MPa
+                ax.set_title(f'Strain: {float(strain)*100}%, Stress: {stress:.3f} MPa')
+                ax.set_xlim((0, pdata['bandstructure']['label_locs'][-1]))
+                ax.set_ylabel("Frequency [THz]")
+                plt.savefig(f'/tmp/{structure_name}_phonon_bandstructure_{strain}.png')
+                plt.close(fig)
+    
+    fig, ax = plt.subplots()
+    with connect(dbname) as db:
+        first_row = db.get(structure_name=structure_name, model_name=model_name)
[…]

### 2023-12-05 src/tex/ms.tex
Changes to scripts, Snakefile. Adding more calcs.

@@
 \usepackage{pdflscape}
+%% Handles multipage table. Not sure allowed by pub.
+\usepackage{longtable}
 
@@
 \begin{figure}[ht!]
+    \script{PlotNoStrainPhonons.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/B19P_combined_models_phonon_bandstructures.png}
+        \caption{
+          Phonon bandstructure of B19P phase for all models.
+        }
+        \label{fig:allmodels_no_strainphonon_B19P}
+    \end{centering}
+\end{figure}
+
+\begin{figure}[ht!]
+    \script{PlotNoStrainPhonons.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/B19_combined_models_phonon_bandstructures.png}
+        \caption{
+          Phonon bandstructure of B19 phase for all models.
+        }
+        \label{fig:allmodels_no_strainphonon_B19P}
+    \end{centering}
+\end{figure}
+
+\begin{figure}[ht!]
+    \script{PlotNoStrainPhonons.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/B19P_combined_models_phonon_bandstructures.png}
+        \caption{
+          Phonon bandstructure of BCO phase for all models.
+        }
+        \label{fig:allmodels_no_strainphonon_B19P}
+    \end{centering}
+\end{figure}
+
+\begin{figure}[ht!]
     \script{PlotEAMPhonons.py}
@@
 
+\section{Conclusion}
+\label{sec:conclusion}
+This word aimed to provide an assement of the of GNN univeral interatomic potentials on the ability to capture the stable ground-state phases of binary and ternary SMA.
+In addition, the dynamic stability of these SMA is crucially evaluated to understand the suitability in observing the onset of displacive phase transformations via the softening of TA branch modes.
+In the case of NiTi we find that all the GNN potentials predict either the Monoclinic or BCO phases as stable over the B2 structure.
+However, not all agree with the recent works by (cite) which have indicated that it is the BCO phase which is the ground-state structure.
+
+We observe that the M3GNet and CHGnet capture the M-point instability of the B2 phase, however, no stablizing isotropic strain is observed.
+With regard to the MACE, very good agrement is found with the phonon bandstructure and that from ab-initio methods.
+The ALIGNN potential was found not to be suitable of NiTi.
+
 \bibliographystyle{naturemag}

### 2023-12-05 src/scripts/CalculateGruneisen.py

@@ -1,0 +1,134 @@
+import paths
+import numpy as np
+import numpy.polynomial.polynomial as poly
+from ase.db import connect
+
+
+def FindIndex(data, label):
+    """
+        locates the index for the q-point value that is closest to the
+    point of interest.
+    """
+    if label not in data["bandstructure"]["labels"]:
+        return "Label not found"
+
+    label_index = data["bandstructure"]["labels"].index(label)
+    label_loc = data["bandstructure"]["label_locs"][label_index]
+    q_points = data["bandstructure"]["q_points"]
+    closest_index = min(
+        range(len(q_points)), key=lambda i: abs(q_points[i] - label_loc)
+    )
+
+    return closest_index
+
+
+def ModeGruneisen(phonons, q, n=3):
+    """
+    Calculate the specific mode Gruneisen parameter via a polynomial fit.
+    """
+
+    volumes = []
+    ef_list = []
+    for k in phonons.keys():
+        volumes.append(phonons[k]["volume"])
+        mode_index = FindIndex(phonons[k], q)
+        mode_bands = []
+        for band in phonons[k]["bandstructure"]["band_index"].values():
+            mode_bands.append(band[mode_index])
+        ef_list.append(mode_bands)
+    ef = np.array(ef_list)
+
+    v0 = phonons["0.00"]["volume"]
+    domega_dv0 = np.zeros(ef.shape[1])
+    for i in range(ef.shape[1]):
+        coefs = poly.polyfit(volumes, ef[:, i], n)
+        derivative = poly.polyder(coefs)
+        # Evaluate the derivative at the equilibrium volume
+        domega_dv0[i] = poly.polyval(v0, derivative)
+
+    return domega_dv0
+
+
+def Gruneisen(phonons, n=3):
+    """
+    Calculate the Gruneisen parameters via a polynomial fit.
+    """
+
+    volumes = []
+    ef_list = []
+    for k in phonons.keys():
+        volumes.append(phonons[k]["volume"])
+        bands = []
+        for band in phonons[k]["bandstructure"]["band_index"].values():
+            bands.append(band)
+        ef_list.append(bands)
+    ef = np.array(ef_list)
+
+    v0 = phonons["0.00"]["volume"]
+    domega_dv0 = np.zeros((ef.shape[1], ef.shape[2]))
+    for i in range(ef.shape[1]):
+        for j in range(ef.shape[2]):
+            # Fit a polynomial to the i-th mode
+            coefs = poly.polyfit(volumes, ef[:, i, j], n)
[…]

### 2023-12-05 src/scripts/CalculatePhonons.py

@@ -14,10 +14,12 @@ from ase.constraints import StrainFilter, UnitCellFilter
 from ase.spacegroup.symmetrize import FixSymmetry, check_symmetry
 
 
+from ConfigsUtils import *
+
 EV_to_THz = 241.799050402293e0
 
 
-def PlotBandstructure(bs, dos, info, ymaxlim=13.0, of="."):
+def PlotBandstructure(bs, dos, info, ymaxlim=9.0, of="."):
     """
     Plots a single phonon band structure and DOS.
     """
@@ -39,8 +41,8 @@ def PlotBandstructure(bs, dos, info, ymaxlim=13.0, of="."):
         ax.plot(q_points, omega, "k-", lw=2)
 
     ax.set_title(
-        f"Lattice Parameter: {info['a']:.3f}Å\
-        Strain: {info['strain_percent']:.2f}% (a0={info['a0']:.3f}Å)\
+        f"Volume: {info['vol']:.3f}Å^3\
+        Strain: {info['strain']:.2f}% \
         Potential: {info['potname']}",
         fontsize=10,
     )
@@ -66,7 +68,10 @@ def PlotBandstructure(bs, dos, info, ymaxlim=13.0, of="."):
     dosax.set_xticks([])
     dosax.set_xlabel("DOS", fontsize=18)
     fig.savefig(
-        os.path.join(of, f"{info['potname']}_a={info['a']:.3f}_NiTi-B2_Phonons.png")
+        os.path.join(
+            of,
+            f"{info['structure']}_{info['potname']}_{info['strain']:.3f}_Phonons.png",
+        )
     )
     fig.clear()
 
@@ -94,7 +99,7 @@ def WriteModeWisual(
 
     NOT TESTED with showyourwork flow
     """
-    
+
     # Write modes for specific q-vector to trajectory files:
     bandpath = structure.cell.bandpath(npoints=pts)
     p = bandpath.special_points[point]
@@ -128,19 +133,23 @@ def WriteModeWisual(
 def GetPhonons(
     structure,
     potential,
+    bandpath,
     supercell=(8, 8, 8),
-    bspts=150,
+    # bspts=150,
     doskpts=(30, 30, 30),
     dospts=150,
     displacement=0.01,
+    center_refcell=True,
     lorentz_width=4.5e-4,
 ):
     """
     displacement is in Angstrom and is atom movement.
     """
-    bandpath = structure.get_cell().bandpath(npoints=bspts)
[…]

### 2023-12-05 src/scripts/ConfigsUtils.py

@@ -1,0 +1,62 @@
+ORDER = {
+    "Mutter": 1,
+    "Zhong": 2,
+    "M3GNet": 3,
+    "CHGNet": 4,
+    "MACE": 5,
+    "ALIGNN": 6,
+}
+
+
+SPACEGROUP_MAP = {"B2" : 221,
+                  "B19P" : 11,
+                  "B19" : 11,
+                  "BCO" : 63,
+                  }
+
+def GetPhononCalcConfig(structure_name, potential):
+    """
+    Settings have been manually calibrated.
+    """
+    # Supercell, displacement (ang.), band path points
+    configurations = {
+        "B2": {
+            "ALIGNN": ((6, 6, 6), 0.025, 200),
+            "Mutter": ((4, 4, 4), 0.025, 200),
+            "Zhong": ((5, 5, 5), 0.025, 200),
+            "M3GNet": ((5, 5, 5), 0.01, 200),
+            "CHGNet": ((5, 5, 5), 0.050, 200),
+            "MACE": ((5, 5, 5), 0.01, 200),
+        },
+        "B19": {
+            "Mutter": ((4, 4, 4), 0.050, 200),
+            "Zhong": ((5, 5, 5), 0.050, 200),
+            "M3GNet": ((4, 4, 4), 0.010, 200),
+            "CHGNet": ((5, 5, 5), 0.010, 200),
+            "MACE": ((4, 4, 4), 0.01, 200),
+            "ALIGNN": ((4, 4, 4), 0.025, 200),
+        },
+        "B19P": {
+            "Mutter": ((4, 4, 4), 0.050, 200),
+            "Zhong": ((5, 5, 5), 0.050, 200),
+            "M3GNet": ((4, 4, 4), 0.010, 200),
+            "CHGNet": ((5, 5, 5), 0.010, 200),
+            "MACE": ((4, 4, 4), 0.01, 200),
+            "ALIGNN": ((4, 4, 4), 0.025, 200),
+        },
+        "BCO": {
+            "Mutter": ((2, 2, 2), 0.050, 200),
+            "Zhong": ((2, 2, 2), 0.050, 200),
+            "M3GNet": ((2, 2, 2), 0.010, 200),
+            "CHGNet": ((2, 2, 2), 0.010, 200),
+            "MACE": ((2, 2, 2), 0.01, 200),
+            "ALIGNN": ((2, 2, 2), 0.025, 200),
+        },
+    }
+
+    # Validate structure_name and potential_key
[…]

### 2023-12-05 src/scripts/GenerateMGruneisenTable.py

@@ -1,0 +1,57 @@
+import sys
+from ase.db import connect
+import paths
+from CalculateGruneisen import *
+from ConfigsUtils import *
+
+
+def GenerateMModeGruneisen(dbname, output_filename):
+    db = connect(dbname)
+    unique_structure_types = set(entry.structure_name for entry in db.select())
+
+    with open(output_filename, "w") as file:
+        file.write("\\begin{table}[h]\n")
+        file.write("\\centering\n")
+
+        # Begin tabular environment
+        file.write("\\begin{tabular}{|c|c|}\n")
+        file.write("\\hline\n")
+        file.write("Structure & \\textbf{M}-Mode Gr\\\"{u}neisen Parameters \\\\\n")
+        file.write("\\hline\n")
+
+        for structure_type in unique_structure_types:
+            if structure_type != "B2":
+                continue
+            structure_name = structure_type.replace("_", "-")
+            file.write(f"{structure_name} & ")
+            file.write("\\begin{tabular}{c c c c c c c}\n")  # Inner table header
+            file.write("Model & TA$_1$ & TA$_2$ & LA & TO$_1$ & TO$_2$ & LO \\\\\n")
+            file.write("\\hline\n")
+            entries = db.select(structure_name=structure_type)
+            sentries = sorted(entries, key=lambda entry: ORDER[entry.model_name])
+            for entry in sentries:
+                model_name =  entry.get('model_name', 'NA')
+                if model_name == "ALIGNN":
+                    if structure_name != "B2":
+                        UserWarning(f"Skipping: {structure_name} {model_name}")
+                        continue
+
+                _, gamma = CalculateModeGruneisen(entry)
+                file.write(
+                    f"{model_name} & {gamma[0]:.2f} & {gamma[1]:.2f} & {gamma[2]:.2f} & {gamma[3]:.2f} & {gamma[4]:.2f} & {gamma[5]:.2f} \\\\\n"
+                )
+
+            file.write("\\hline\n")
+            file.write("\\end{tabular}\\\\\n")
+            file.write("\\hline\n")
+        file.write("\\end{tabular}\n")
+        file.write(
+            "\\caption{Mode Gruneisen parameters for different NiTi structures.}\n"
+        )
+        file.write("\\label{tab:mode_gruneisen_niti}\n")
+        file.write("\\end{table}\n")
+
+
+if __name__ == "__main__":
[…]

### 2023-12-05 src/scripts/GenerateNiTiEquilTable.py

@@ -11,20 +11,12 @@ from ase.spacegroup import get_basis
 from ase.spacegroup import Spacegroup
 from ase.spacegroup import Spacegroup, get_spacegroup
 
-
-ORDER = {'Mutter':1,
-         'Zhong':2,
-         'M3GNet':3,
-         'CHGNet':4,
-         'MACE':5,
-         'ALIGNN':6,
-         }
-
+from ConfigsUtils import *
 
 def GetElementBasisList(structure, spacegroup, tol=1e-3):
     """
     NOT CONFIRMED
-    Provides the basis/wycoff sites with element labels
+    Provides the basis/wykcoff sites with element labels
     """
     sg = Spacegroup(spacegroup)
     rotations = sg.rotations
@@ -63,22 +55,25 @@ def GenerateStructureTable(dbname, output_filename):
     unique_structure_types = set(entry.structure_name for entry in db.select())
     
     with open(output_filename, 'w') as file:
-        #file.write("\\begin{landscape}\n")
-        file.write("\\begin{table}[h]\n")
-        file.write("\\centering\n")
-        
-        # Outer table header
-        file.write("\\begin{tabular}{|l|c|}\n")
+        file.write("\\begin{longtable}{|l|c|}\n")
+        file.write("\\caption{Equilibrium structures for NiTi.} \\label{tab:equil_niti}\\\\\n")
         file.write("\\hline\n")
         file.write("Structure & Unit Cell \\\\\n")
         file.write("\\hline\n")
-        
+        file.write("\\endfirsthead\n")
+        file.write("\\caption[]{Equilibrium structures for NiTi. (Continued)}\\\\\n")
+        file.write("\\hline\n")
+        file.write("Structure\n(Spacegroup \#) & Unit Cell \\\\\n")
+        file.write("\\endhead\n")
+        file.write("\\hline\n")
+        file.write("\\endfoot\n")
+
         for structure_type in unique_structure_types:
             structure_name = structure_type.replace("_","-")
-            file.write(f"{structure_name} & ")
+            spg_num = SPACEGROUP_MAP[structure_name]
+            file.write(f"{structure_name}\n({spg_num}) & ")
             file.write("\\begin{tabular}{c c c c c c c}\n")  # Inner table header
             file.write("Model & $a$ (\AA) & $b$ (\AA) & $c$ (\AA) & $\\alpha$ ($^\circ$) & $\\beta$ ($^\circ$) & $\\gamma$ ($^\circ$)\\\\\n")
-            file.write("\\hline\n")
[…]

### 2023-12-05 src/scripts/NiTiStructures.py

@@ -1,8 +1,7 @@
 from ase.spacegroup import crystal
 
-def NiTi_B2_Structure(
-    a=3.04, b=3.04, c=3.04, alpha=90.0, beta=90.0, gamma=90.0
-):
+
+def NiTi_B2_Structure(a=3.04, b=3.04, c=3.04, alpha=90.0, beta=90.0, gamma=90.0):
     """
     Austenite phase.
     """
@@ -17,40 +16,55 @@ def NiTi_B2_Structure(
     return B2
 
 
-def NiTi_B33_Structure(
-    a=2.888, b=9.398, c=4.042, alpha=90.0, beta=90.0, gamma=90.0
-):
+def NiTi_B33_Structure(a=2.888, b=9.398, c=4.042, alpha=90.0, beta=90.0, gamma=90.0):
     """
-    Structure from: Liu et al., Phys. Rev. B 87, 140104(R) 2018. 
+    Structure from: Liu et al., Phys. Rev. B 87, 140104(R) 2018.
     Cmcm (B33/BCO) @ 0 GPa
     """
     B33 = crystal(
         ["Ni", "Ti"],
-        [(0.0,0.5847,0.25),(0.0,0.1440,0.75)],
+        [(0.0, 0.5847, 0.25), (0.0, 0.1440, 0.75)],
         cellpar=[a, b, c, alpha, beta, gamma],
         pbc=True,
         spacegroup=63,
     )
     return B33
 
-def NiTi_Cmcm_Structure(
-        a=2.90, b=3.99, c=4.90, alpha = 90.0, beta=107.01, gamma=90
-        ):
+
+def NiTi_Cmcm_Structure(a=2.90, b=3.99, c=4.90, alpha=90.0, beta=107.01, gamma=90.0):
+    """
+    Add reference paper, confirm!
+    """
     Cmcm = crystal(
         ["Ni", "Ti"],
-        [(0.915362,0.75,0.829275),(0.641384,0.75,0.284766)],
+        [(0.915362, 0.75, 0.829275), (0.641384, 0.75, 0.284766)],
         cellpar=[a, b, c, alpha, beta, gamma],
         pbc=True,
-        spacegroup=11, #Should be 63 according to mp data
+        spacegroup=11,  # Should be 63 according to mp data
     )
     return Cmcm
-    
+
+
+def NiTi_BCO_Structure(a=2.8987, b=9.3619, c=3.9870, alpha=90.0, beta=90.0, gamma=90.0):
+    """
[…]

### 2023-12-05 src/scripts/PlotGruneisen.py

@@ -1,0 +1,48 @@
+
+import matplotlib.pyplot as plt
+
+
+def PlotGruneisen(gruneisen_data):
+    if gruneisen_data is None:
+        print("No data to plot.")
+        return
+
+    q_points = gruneisen_data['q_points']
+    num_bands = len(gruneisen_data['band_index'])
+
+    plt.figure(figsize=(10, 6))
+    for band_index in range(1, num_bands + 1):
+        gamma_values = gruneisen_data['band_index'][band_index]
+        plt.plot(q_points, gamma_values, label=f'Band {band_index}')
+    plt.ylim((-50,50))
+    plt.xlabel('Q-point')
+    plt.ylabel('Gruneisen Parameter')
+    plt.title('Mode Gruneisen Parameters for Each Band')
+    plt.legend()
+    plt.grid(True)
+    plt.savefig("/tmp/GruneisenAll.png")
+
+def PlotGruneisenFrequency(ph,gruneisen_data):
+    if gruneisen_data is None:
+        print("No data to plot.")
+        return
+
+    q_points = gruneisen_data['q_points']
+    num_bands = len(gruneisen_data['band_index'])
+    
+    plt.figure(figsize=(10, 6))
+    for band_index in range(1, num_bands + 1):
+        gamma_values = gruneisen_data['band_index'][band_index]
+        frequencies = ph["0.00"]["bandstructure"]["band_index"][band_index-1]
+
+        # Scatter plot for each band
+        plt.scatter(frequencies, gamma_values, label=f'Band {band_index}')
+
+    plt.ylim((-50, 50))
+    plt.xlabel('Frequency')
+    plt.ylabel('Gruneisen Parameter')
+    plt.title('Gruneisen Parameters vs. Frequency')
+    plt.legend()
+    plt.grid(True)
+    plt.savefig("/tmp/Gruneisen_vs_Frequency.png")
+

### 2023-12-05 src/scripts/PlotNiTiEOS.py

@@ -78,7 +78,10 @@ def plot_eos(db_path, figsize=(3.4, 5.75), dpi=600, xlim=(0.75, 1.25), ignore_li
                 for row in db.select(model_name=calculator_model, structure_name=structure_name):
                     volumes = np.array(row.data['volumes'])
                     energies = np.array(row.data['energies'])
-                    vi = row.vi
+                    #NOTE: We can't use EOS fit volume, as the fits aren't always that good.
+                    #vi = row.v0
+                    min_energy_index = np.argmin(energies)
+                    vi = volumes[min_energy_index]
                     natoms = row.natoms
                     
                     vv_ratio = volumes / vi
@@ -87,7 +90,7 @@ def plot_eos(db_path, figsize=(3.4, 5.75), dpi=600, xlim=(0.75, 1.25), ignore_li
 
                     # Update the global y-axis limits
                     ymin =  min(energy_per_atom) - 0.1
-                    ymax =  max(energy_per_atom)
+                    ymax =  max(energy_per_atom) * 1.05
                     ax.set_ylim((ymin, ymax))
 
         handles, labels = axs[-1].get_legend_handles_labels()

### 2023-12-13 src/tex/ms.tex
Add elastic calcs, Gruneisen param, calibrated ConfigsUtils, added reference.bib, etc.

@@
 % Author list
-\author[1]{Stefan Bringuier\corref{cor1}}
+\author[1]{Stefan Bringuier\corref{cor1}}[role=Lead Researcher,orcid=0000-0001-6753-1437]
 \ead{stefanbringuier@gmail.com}
@@
 \label{sec:intro}
+Shape memory alloys derive their unique displacive phase transformations from the dynamic instablity that occurs during the application of temperature, strain, or stress.
+The origin of this instability occurs due to softening or symmetry breaking of the phonon modes in the phase.
+Therefore, improved understanding of the phonon characteristics of the different phases that can occur for a given SMA is particular important.
+One of the challenges is in performing such calculations, which typically employ density functional theory in order to determine the atomic forces, which in turn are used to construct the dynamical matrix.
+However, the use of classical interatomic potentials can alleviate some of this computational burden.
+\par
+
+Recently, there as been a growing effort in the development of machine learning potentials.
+These deviate from traditional physics approaches to interatomic potentials, which have historically been inspired from more accurate methods like tight-binding method.
+In machine learning potentials, one describes the atoms and structure using descriptors of features, these are typically learned or prescribed.
+
+The main inquiry of this work is to asses the GNN interatomic potentials on binary SMA NiTi and PbTi primarily through the phonon predictions.
+Our analysis aims to provide a qualitative perspective based on previously studied phonon phenomena of these systems.
+The findings here will provide guidnace for future use of GNN interatomic potentials for the development other binary, and more so, ternary and quaternary SMAs.
+IMportantly, the use of pretrained base models for these systems acts as an initial seed and can be used to fine-tune future GNN potentials for better SMA disocovery works.
+\par
 
@@
 \label{sec:methods}
+
+\subsection{Workflow}
+\label{sec:workflow}
+The ground-state structures for the NiTi phases \textit{B2}, \textit{B19}, \textit{B19'}, and \textit{BCO} are generated using the atomic simulation environment (ASE) package (cite).
+This requires that the unit cell parameters, basis atoms, and spacegroup number be provided.
+These initial reference parameters for these structures have been taken from the materials project (cite specific DOIs for the structures).
+The energy and forces on the atoms are determined using the calculator interface routines which enables the assesment of different interatomic potentials for the same ASE structure.
+Similarly, for all EOS, phonon, and elastic constant claculations each step is performed for a specific interatomic potential via the ASE calculator interface.
+We not that this is done in different python computing environments using \href{https://github.com/showyourwork/showyourwork}{\showyourwork} package.\par
+
+First the structures are optimized using ASE local optimization routines.
+The forces were converged to $10^{-4}\,\text{eV}/\AA$ and utlize the FIRE optimizater, but found other local optimization options found same ground-state structures.
+To ensure that the unit cell parameters were adjusted along with the atomic coordinates and that the crystal symmetry was not broken, we utlized the ASE constraint and filter routines during optimization of the structures.
+\par
+
+To determine the equation of state, we started with the obtained ground-state structures.
+To adjust the volume of the unit cell, we then apply an isotropic strain tensor to the cell and scale the atomic coordinates
+
+\begin{equation}
+  \label{eq:isotropic_strain}
+  T=\begin{bmatrix}
+    \epsilon & 0 & 0 \\
+    0 & \epsilon & 0 \\
+    0 & 0 & \epsilon \\
+  \end{bmatrix}
+\end{equation}
+
+were $\epsilon$ is the fraction strain applied to the unit cell.
+To ensure that the strain is indeed isotropic we provide a quality check on the symmetry of the distorted structure.
+The isotropic strain is applied in compression and tension, from $-11\%$ to $11\%$, which gave a good range to asses the EOS curves and obtain fits.
+We fit the EOS to a standard polynomial which provides values for the equilibrium volume, energy, and bulk modulus.
+In order to compare respective equations of state amongst the different phases, we normalize the volume of each phase with respect to the obtained ground-state volume.
+\par
+
+Phonon calculations are done using the ASE phonons module, which uses the finite-difference method via small displacement of atoms in a reference cell.
+This requires the use of a supercell of the reference structure.
+In order to ensure convergence of this method, one needs to sweep different supercell sizes and how the displacement size impacts the phonon eigenvalues and vectors.
+This process was done initially for each interatomic potential and structure.
+Details on the settings are provided in the appendix.
+\par
+
+The ASE structures provide a routine for sampling the irreducible Brillouin zone for the specific structure, however, a significant amount of additional symmetry points are include in the band path.
+For this reason we only focus on high-symmetry points in the IBZ, given that the monoclinic and othorhombic phases are low symmetry structures.
+The exception is the cubic structure which we sample the full IBZ.
+The IBZ and points symmetry point locations can be seen in appendix \ref{sec:appx_ibz}.
+The final phonon dispersion diagrams are plot along these points.
+\par
+
+To better understand dynamic instability in the various structures, same isotropic strain in eq.~\ref{eq:isotropic_strain} is used and respetive phonon bandstructures are calculated.
+These are then plotted to observed the onset or change of the instability of different branches.
+\par
+
+The elastic constants have been calculated using ....
 
@@
 \label{subsec:nitieos}
-\cite{Luger2021}
+
 The thermodynamic equation of state (EOS) is a standard assement of the phase stability of materials at a give condition.
@@
 \begin{figure}[ht!]
+    \script{PlotNiTiCohesiveEnergy.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/NiTi_CohesiveEnergy.png}
+        \caption{
+          Comparison of predicted cohesive energy for respective ground-state structure and model.
+        }
+        \label{fig:ecoh}
+    \end{centering}
+\end{figure}
+
+\begin{figure}[ht!]
     \script{PlotNiTiEOS.py}
@@
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/B19P_combined_models_phonon_bandstructures.png}
-        \caption{
-          Phonon bandstructure of B19P phase for all models.
-        }
-        \label{fig:allmodels_no_strainphonon_B19P}
-    \end{centering}
-\end{figure}
-
-\begin{figure}[ht!]
-    \script{PlotNoStrainPhonons.py}
-    \begin{centering}
         \includegraphics[width=\linewidth]{figures/B19_combined_models_phonon_bandstructures.png}
         \caption{
-          Phonon bandstructure of B19 phase for all models.
+          Ground-state phonon bandstructure of B19 phase for all models.
         }
@@
         \caption{
-          Phonon bandstructure of BCO phase for all models.
+          Ground-state phonon bandstructure of \textit{B19}$^\prime$ phase for all models.
+        }
+        \label{fig:allmodels_no_strainphonon_B19P}
+    \end{centering}
+\end{figure}
+
+\begin{figure}[ht!]
+    \script{PlotNoStrainPhonons.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/BCO_combined_models_phonon_bandstructures.png}
+        \caption{
+          Ground-state phonon bandstructure of BCO phase for all models.
         }
@@
 
-\subsection{NiTiCu: Equation of States}
+The B2 phase of NiTi exhibits notable phonon instabilities, particularly at the \textit{M} point in the Brillouin zone.
+These instabilities are characterized by imaginary modes, represented as negative phonon frequencies, indicating a dynamical instability in this crystal structure.
+The \textit{M} point, located at the edge of the Brillouin zone, corresponds to a wave vector of $\left[\frac{1}{2}, \frac{1}{2}, 0\right]$ in units of the reciprocal lattice vectors.
+The phonon dispersion curves for NiTi in the B2 phase reveal a softening of the acoustic modes at the \textit{M} point, suggesting a propensity for a structural phase transition.
+\par
+
+The mode Gruneisen parameter, $\gamma$, plays a pivotal role in understanding the stabilization of the NiTi B2 phase under external strain, particularly in the context of phonon instabilities at the M point.
+The mode Gruneisen parameter is defined as,
+
+\begin{equation}
+  \label{eq:modegruneisen}
+  \gamma = -\frac{V}{\omega}\left(\frac{\partial \omega}{\partial V}\right)
+\end{equation},
+
+where $V$ is the volume, usually taken at equilibrium, and $\omega$ is the phonon frequency.
+For the B2 phase of NiTi, the negative phonon frequencies at the M point indicate an instability that can be modulated by applying strain.
+The strain perturbs the lattice volume, thereby altering the phonon frequencies.
+The effect of this strain can be quantitatively assessed using the mode Gruneisen parameter, which describes how the phonon frequencies change with volume.
+A positive $\gamma$ suggests that the phonon frequency increases as the volume decreases under strain, potentially stabilizing the lattice.
+In the case of NiTi, applying compressive strain may increase the frequencies of the soft modes at the M point, thus stabilizing the B2 phase and delaying or suppressing the martensitic transformation.
+This interplay between strain and phonon behavior, quantified by $\gamma$, is crucial in tailoring the mechanical properties of shape memory and superelastic materials.
+
+\variable{output/Table_NiTi_M_ModeGruneisen.tex}
+
+\subsection{PbTi: Equation of States}
 
@@
 \bibliographystyle{naturemag}
-\bibliography{bib}
+\bibliography{references}
 
@@
 \section*{Reproducibilty and Data Availability}
-This study was carried out using the reproducibility software \href{https://github.com/showyourwork/showyourwork}{\showyourwork} \cite{Luger2021}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}, create the figures, and compile the manuscript.
+This study was carried out using the reproducibility software \href{https://github.com/showyourwork/showyourwork}{\showyourwork}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}, create the figures, and compile the manuscript.
 Each figure caption contains two links: one to the dataset stored on zenodo used in the corresponding figure, and the other to the script used to make the figure (at the commit corresponding to the current build of the manuscript).
@@
 \appendix
+\subsection{Irreducible Brillouin Zones}
+\label{sec:appx_ibz}
+
+The irreducible Brillouin zones for the the different crystals investigated in this study are provided below.
+The high-symmetry points are indicated.
+Additionally the coordinates of the high-symmetry points in the IBZ are shown in the tables below.
+In Ref.
+xxx the  specific implementation of the band path sampling and definitions are provided.
+
+\begin{figure}[ht!]                                                                                      
+    \script{AppendixPlotBZ.py}                                                                           
+    \begin{centering}                                                                                    
+      \includegraphics[width=0.5\linewidth]{figures/B2_BrillouinZonePointsSampled.png} 
+      \vspace{1mm}                                                                                       
+      \includegraphics[width=0.5\linewidth]{figures/B19_BrillouinZonePointsSampled.png}
+      \vspace{1mm}
+      \includegraphics[width=0.5\linewidth]{figures/BCO_BrillouinZonePointsSampled.png}
+      \vspace{1mm}                                                                                       
+      \caption{The irreducible Brillouin zones for (top) B2(\#221) (middle) B19/B19P(\#11), and (bottom) BCO(\#63) crystal structures}                                                                                  \label{fig:ibz}
+    \end{centering}                                                                                      
+\end{figure}
+
+\variable{output/B2_SpecialSymmetryPointsBZ.tex}
+\variable{output/B19_SpecialSymmetryPointsBZ.tex}
+\variable{output/BCO_SpecialSymmetryPointsBZ.tex}
 

### 2023-12-13 src/scripts/AppendixPlotBZ.py

@@ -1,0 +1,45 @@
+import paths
+
+import matplotlib.pyplot as plt
+
+from NiTiStructures import *
+
+
+structure_names = ["B2", "B19", "B19P", "BCO"]
+fmap = {
+    "B2": NiTi_B2_Structure,
+    "B19": NiTi_B19_Structure,
+    "B19P": NiTi_B19P_Structure,
+    "BCO": NiTi_BCO_Structure,
+}
+
+for name in structure_names:
+    structure = fmap[name]()
+    structure.info["name"] = name
+    spg = structure.info["spacegroup"].no
+    path, _ = get_bandpath(structure,npoints=1)
+    bz_diagram = f"{name}_BrillouinZonePointsSampled.png"
+    bz_plot = path.plot(vectors=True)
+    bz_plot.figure.savefig(paths.figures / bz_diagram)
+    special_points = path.special_points
+
+    # Write table of special points
+    latex_table = "\\begin{table}[h]\n"
+    latex_table += "\\centering\n"
+    latex_table += "\\begin{tabular}{|c c|}\n"
+    latex_table += "\\hline\n"
+    latex_table += "Point & Coordinates \\\\"
+    latex_table += "\\hline\n"
+
+    for point, coords in special_points.items():
+        coords_str = ", ".join(map(str, coords))
+        latex_table += f"{point} & {coords_str} \\\\\n"
+
+    latex_table += "\\hline\n"
+    latex_table += "\\end{tabular}\n"
+    latex_table += f"\\caption{{Special points sampled in the irreducible Brillouin Zone for space group {spg}}}\n"
+    latex_table += "\\end{table}"
+
+    sp_table = f"{name}_SpecialSymmetryPointsBZ.tex"
+    with open(paths.output / sp_table, "w") as file:
+        file.write(latex_table)

### 2023-12-13 src/scripts/CalculateElastic.py

@@ -5,40 +5,87 @@ import ase.units as units
 from ase.spacegroup import crystal
 
 
-from parcalc import ParCalculate
+#from parcalc import ParCalculate
 from elastic import get_elementary_deformations
 from elastic.elastic import get_cij_order
 from elastic import get_elastic_tensor
 
+from ase.calculators.lammpslib import LAMMPSlib
 
-def CalculateElasticConstants(dbname,
-                  potential,
-                  npoints=10,
-                  displacement=0.33):
+from queue import Empty
 
+from multiprocessing import Process, Queue
+
+import time
+import os
+import tempfile
+import shutil
+from copy import deepcopy
+from subprocess import check_output
+from contextlib import contextmanager
+
+
+def SerialCalculate(systems,calc,cleanup=True,prefix="Calc_"):
+
+    if type(systems) != type([]) :
+        sysl=[systems]
+    else :
+        sysl=systems
+    basedir=os.getcwd()
+    res=[]
+    for n,s in enumerate(sysl):
+        if isinstance(calc, LAMMPSlib):
+            s.set_calculator(calc)
+        else:
+            s.set_calculator(deepcopy(calc))
+        s.get_calculator().block=False
+        place=tempfile.mkdtemp(prefix=prefix, dir=basedir)
+        os.chdir(place)
+        s.get_calculator().working_dir=place
+        #print("Start at :", place)
+        s.get_potential_energy()
+        os.chdir(basedir)
+        time.sleep(0.2)
+        res.append([n,s])
+        print("Workers started:", len(sysl))
+
+    return [r for ns,s in enumerate(sysl) for nr,r in res if nr==ns]
+
+
+def calculate_elastic_constants(
+    dbname, structure, potential, npoints=10, displacement=2.0
+):
+    """
+    displacement: percentage of length and angles
+    npoints: number of iterations along displacement
+    """
+    
     db = connect(dbname)
     potname = potential[0]
     calc = potential[1]
-    for entry in db.select(model_name = potname):
+
+        
+        
+    for entry in db.select(structure_name=structure, model_name=potname):
         spg = entry.spacegroup
-        structure = crystal(entry.toatoms(),spacegroup=spg)
+        structure = crystal(entry.toatoms(), spacegroup=spg)
         # Need to set some defaults
         calc.clean = potential[1].reset
-        #calc.set(isif=2)
+        # calc.set(isif=2)
         structure.calc = calc
 
[…]

### 2023-12-13 src/scripts/ConfigsUtils.py

@@ -21,36 +21,36 @@ def GetPhononCalcConfig(structure_name, potential):
     # Supercell, displacement (ang.), band path points
     configurations = {
         "B2": {
-            "ALIGNN": ((6, 6, 6), 0.025, 200),
-            "Mutter": ((4, 4, 4), 0.025, 200),
-            "Zhong": ((5, 5, 5), 0.025, 200),
-            "M3GNet": ((5, 5, 5), 0.01, 200),
-            "CHGNet": ((5, 5, 5), 0.050, 200),
-            "MACE": ((5, 5, 5), 0.01, 200),
+            "Mutter": ((8, 8, 8), 0.0935, 200),
+            "Zhong": ((8, 8, 8), 0.0935, 200),
+            "M3GNet": ((8, 8, 8), 0.260, 200),
+            "CHGNet": ((8, 8, 8), 0.030, 200),
+            "MACE": ((8, 8, 8), 0.010, 200),
+            "ALIGNN": ((6, 6, 6), 0.050, 200),
         },
         "B19": {
-            "Mutter": ((4, 4, 4), 0.050, 200),
-            "Zhong": ((5, 5, 5), 0.050, 200),
-            "M3GNet": ((4, 4, 4), 0.010, 200),
-            "CHGNet": ((5, 5, 5), 0.010, 200),
-            "MACE": ((4, 4, 4), 0.01, 200),
-            "ALIGNN": ((4, 4, 4), 0.025, 200),
+            "Mutter": ((7, 9, 7), 0.05, 200),
+            "Zhong": ((7, 9, 7), 0.05, 200),
+            "M3GNet": ((7, 9, 7), 0.05, 200),
+            "CHGNet": ((7, 9, 7), 0.05, 200),
+            "MACE": ((7, 9, 7), 0.05, 200),
+            "ALIGNN": ((5, 7, 5), 0.05, 200),
         },
         "B19P": {
-            "Mutter": ((4, 4, 4), 0.050, 200),
-            "Zhong": ((5, 5, 5), 0.050, 200),
-            "M3GNet": ((4, 4, 4), 0.010, 200),
-            "CHGNet": ((5, 5, 5), 0.010, 200),
-            "MACE": ((4, 4, 4), 0.01, 200),
-            "ALIGNN": ((4, 4, 4), 0.025, 200),
+            "Mutter": ((8, 6, 4), 0.05, 200),
[…]

### 2023-12-13 src/scripts/PlotPhonons.py

@@ -5,26 +5,28 @@ from matplotlib.gridspec import GridSpec
 from ase.db import connect
 from scipy.interpolate import interp1d
 
+
 def format_xticklabels(labels, locs):
     # Dictionary to keep track of formatted labels with their locations
     formatted_labels_dict = {}
-    
+
     for loc, label in zip(locs, labels):
         if label == "G":  # Check if label is 'G'
-            label = '$\Gamma$'  # Replace with the Greek letter Gamma
-        
+            label = "$\Gamma$"  # Replace with the Greek letter Gamma
+
         # Check if this location already has a label
         if loc in formatted_labels_dict:
             # Append the new label to it with a comma
-            formatted_labels_dict[loc] += ', ' + label
+            formatted_labels_dict[loc] += ", " + label
         else:
             # Otherwise, just add the label
             formatted_labels_dict[loc] = label
-    
+
     # Now, build the list of formatted labels in the order of locations
     formatted_labels = [formatted_labels_dict[loc] for loc in locs]
-    
+
     return formatted_labels
+
 
 def normalize_q_points(q_points):
     return (q_points - np.min(q_points)) / (np.max(q_points) - np.min(q_points))
@@ -53,15 +55,38 @@ def PlotPhonons(dbname, model_name, structure_name, num_strains=None):
                     ax.plot(q_points, energies, color="k")
 
                 ax.set_xticks(pdata["bandstructure"]["label_locs"])
-                ax.set_xticklabels(format_xticklabels(pdata["bandstructure"]["labels"], pdata["bandstructure"]["label_locs"]))
+                ax.set_xticklabels(
+                    format_xticklabels(
+                        pdata["bandstructure"]["labels"],
+                        pdata["bandstructure"]["label_locs"],
+                    )
+                )
+                # Calculate midpoints for q_vec_labels
+                midpoints = 0.5 * np.array(
+                    pdata["bandstructure"]["label_locs"][:-1]
+                    + pdata["bandstructure"]["label_locs"][1:]
+                )
+                for midpoint, vec_label in zip(
+                    midpoints, pdata["bandstructure"]["q_vec_labels"]
+                ):
+                    ax.text(
+                        midpoint,
+                        ax.get_ylim()[1],
+                        vec_label,
+                        ha="center",
+                        va="bottom",
[…]

### 2023-12-19 src/tex/ms.tex
Calculation scripts good. Some plotting good. Added draft MD simulation script. New Calculator.

@@
 \begin{figure}[ht!]
-    \script{PlotNiTiCohesiveEnergy.py}
+    \script{PlotCohesiveEnergy.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/NiTi_CohesiveEnergy.png}
+        \includegraphics[width=\linewidth]{figures/NiTi_CohesiveEnergyPlot.png}
         \caption{
@@
 \begin{figure}[ht!]
-    \script{PlotNiTiEOS.py}
+    \script{PlotEOS.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/NiTi_EOS_Comparison.png}
+        \includegraphics[width=\linewidth]{figures/NiTi_EquationOfStates.png}
         \caption{
@@
 \begin{figure}[ht!]
-    \script{PlotNoStrainPhonons.py}
+    \script{PlotPhonons.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/B2_combined_models_phonon_bandstructures.png}
+        \includegraphics[width=\linewidth]{figures/NiTi_B2_ModelsPhononBandstructures.png}
         \caption{
@@
         }
-        \label{fig:allmodels_no_strainphonon_b2}
+        \label{fig:allmodels_b2}
     \end{centering}
@@
 \begin{figure}[ht!]
-    \script{PlotNoStrainPhonons.py}
+    \script{PlotPhonons.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/B19_combined_models_phonon_bandstructures.png}
+        \includegraphics[width=\linewidth]{figures/NiTi_B19_ModelsPhononBandstructures.png}
         \caption{
@@
         }
-        \label{fig:allmodels_no_strainphonon_B19P}
+        \label{fig:allmodels_B19P}
     \end{centering}
@@
 \begin{figure}[ht!]
-    \script{PlotNoStrainPhonons.py}
+    \script{PlotPhonons.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/B19P_combined_models_phonon_bandstructures.png}
+        \includegraphics[width=\linewidth]{figures/NiTi_B19P_ModelsPhononBandstructures.png}
         \caption{
@@
 \begin{figure}[ht!]
-    \script{PlotNoStrainPhonons.py}
+    \script{PlotPhonons.py}
     \begin{centering}
-        \includegraphics[width=\linewidth]{figures/BCO_combined_models_phonon_bandstructures.png}
+        \includegraphics[width=\linewidth]{figures/NiTi_BCO_ModelsPhononBandstructures.png}
         \caption{
@@
     \begin{centering}
-        \includegraphics[width=0.5\linewidth]{figures/Mutter_B2_combined_strain_phonon_bandstructures.png}
+        \includegraphics[width=0.5\linewidth]{figures/NiTi_Mutter_B2_StrainsPhononBandstructures.png}
         \vspace{1mm}
-        \includegraphics[width=0.5\linewidth]{figures/Zhong_B2_combined_strain_phonon_bandstructures.png}
+        \includegraphics[width=0.5\linewidth]{figures/NiTi_Zhong_B2_StrainsPhononBandstructures.png}
         \caption{
@@
     \begin{centering}
-      \includegraphics[width=0.5\linewidth]{figures/M3GNet_B2_combined_strain_phonon_bandstructures.png}
+      \includegraphics[width=0.5\linewidth]{figures/NiTi_M3GNet_B2_StrainsPhononBandstructures.png}
       \vspace{1mm}
-      \includegraphics[width=0.5\linewidth]{figures/CHGNet_B2_combined_strain_phonon_bandstructures.png}
+      \includegraphics[width=0.5\linewidth]{figures/NiTi_CHGNet_B2_StrainsPhononBandstructures.png}
       \vspace{1mm}
-      \includegraphics[width=0.5\linewidth]{figures/MACE_B2_combined_strain_phonon_bandstructures.png}
+      \includegraphics[width=0.5\linewidth]{figures/NiTi_MACE_B2_StrainsPhononBandstructures.png}
       \caption{
@@
 \begin{figure}[ht!]                                                                                      
-    \script{AppendixPlotBZ.py}                                                                           
+    \script{AppendixBZ.py}                                                                           
     \begin{centering}                                                                                    

### 2023-12-19 src/scripts/CalculateEOS.py

@@ -9,6 +9,7 @@ from ase.constraints import StrainFilter
 from ase.units import kJ
 from ase.eos import EquationOfState
 
+from MinimizeStructure import minimize_structure
 
 def to_scalar(value):
     if isinstance(value, np.ndarray) and value.size == 1:
@@ -71,7 +72,7 @@ def calculate_eos(structure,
         plot_eos(eos,model,species,name)
         
 
-    return (v0, e0, B, volumes, energies)
+    return (eos_eq, v0, e0, B, volumes, energies)
 
 def plot_eos(eos,potential_name,specie_names,structure_name):
     # Plot and save EOS
@@ -110,7 +111,7 @@ def get_min_structure_and_calculate_eos(structure,
 
     nextrow, opt_structure = minimize_structure(structure,potential,write_db=dbname)
              
-    v0,e0,B,volumes,energies = calculate_eos(opt_structure,potential,strain)
+    eos_eq, v0,e0,B,volumes,energies = calculate_eos(opt_structure,potential,strain)
 
     db = connect(dbname)
     db.update(id=nextrow,

### 2023-12-19 src/scripts/Calculators.py

@@ -35,7 +35,7 @@ def get_ase_calculator(model="MutterASE"):
     if model == "Mutter":
         from ase.calculators.lammpslib import LAMMPSlib
 
-        potential_file = str(paths.static / "NiTi_<utter.eam.fs")
+        potential_file = str(paths.static / "NiTi_Mutter.eam.fs")
         cmds = ["pair_style eam/fs", f"pair_coeff * *  {potential_file} Ni Ti"]
         amds = ["thermo_style custom etotal lx ly lz vol pxx pyy pzz pxy pxz pyz press"]
         asecalc = LAMMPSlib(lmpcmds=cmds, amendments=amds, keep_alive=True)
@@ -47,6 +47,7 @@ def get_ase_calculator(model="MutterASE"):
     elif model == "Zhong":
         from ase.calculators.lammpslib import LAMMPSlib
 
+        potential_file = str(paths.static / "NiTi_Zhong.eam.fs")
         cmds = ["pair_style eam/fs", f"pair_coeff * *  {potential_file} Ni Ti"]
         amds = ["thermo_style custom etotal lx ly lz vol pxx pyy pzz pxy pxz pyz press"]
         asecalc = LAMMPSlib(lmpcmds=cmds, amendments=amds, keep_alive=True)
@@ -62,12 +63,12 @@ def get_ase_calculator(model="MutterASE"):
         m3gnet_params = matgl.load_model("M3GNet-MP-2021.2.8-PES")
         asecalc = M3GNetCalculator(m3gnet_params, stress_weight=1.0)
 
-    elif args.model == "CHGNet":
+    elif model == "CHGNet":
         from chgnet.model.dynamics import CHGNetCalculator
 
         asecalc = CHGNetCalculator(stress_weight=1.0)
 
-    elif args.model == "MACE":
+    elif model == "MACE":
         from mace.calculators import MACECalculator
 
         asecalc = MACECalculator(
@@ -76,12 +77,21 @@ def get_ase_calculator(model="MutterASE"):
             default_dtype="float32",
         )
 
-    elif args.model == "ALIGNN":
+    elif model == "ALIGNN":
         from alignn.ff.ff import AlignnAtomwiseCalculator, default_path
 
         alignn_params = default_path()
         asecalc = AlignnAtomwiseCalculator(path=alignn_params, device="cpu")
 
+    elif model == "Ko":
+        from ase.calculators.lammpslib import LAMMPSlib
+
[…]

### 2023-12-19 src/scripts/PlotEAMPhonons.py

@@ -1,14 +1,0 @@
-import sys
-import paths
-import json
-from PlotPhonons import *
-
-
-eam_models = ["Mutter", "Zhong"]
-dbname = paths.data / sys.argv[1]
-nstrains = int(sys.argv[2])
-structures = sys.argv[3:]
-
-for m in eam_models:
-    for s in structures:
-        PlotPhonons(dbname, m, s, num_strains=nstrains)

### 2023-12-19 src/scripts/PlotEOS.py

@@ -1,0 +1,139 @@
+import sys
+import numpy as np
+import os
+import paths
+
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+from matplotlib.ticker import MultipleLocator, AutoLocator
+from cycler import cycler
+import seaborn as sns  # For color palette
+from ase.db import connect
+
+from Config import SUBPLOT_ORDER
+
+def plot_eos(
+    dbname,
+    output_plot,
+    figsize=(3.4, 5.75),
+    dpi=600,
+    xlim=(0.75, 1.25),
+    ignore_list=["B32", "R_Phase"],
+    color_palette="muted",
+    n_colors=10,
+    wspace=0.3,
+    hspace=0.3,
+    linewidth=1,
+    linestyle="-",
+    fontsize=8,
+    tick_fontsize=6,
+    legend_fontsize=8,
+    legend_ncol=3,
+    label_notes="EOS_Labels.txt",
+):
+
+    db = connect(dbname)
+
+    calculator_models = sorted(
+        set(row.model_name for row in db.select()),
+        key=lambda model: list(SUBPLOT_ORDER.values()).index(model),
+    )
+
+    structure_names = set(row.structure_name for row in db.select())
+
+    # Create a figure with a 2x3 grid of subplots
+    fig, axs = plt.subplots(3, 2, figsize=figsize)
+    axs = axs.ravel()  # Flatten the array of axes for easier indexing
+
+    # Create a color cycle iterator using a slightly modified color palette
+    color_cycle = sns.color_palette(color_palette, n_colors=n_colors)
+
+    label_file_path = paths.figures / label_notes
+    with open(label_file_path, "w") as label_file:
+        for i, calculator_model in enumerate(calculator_models):
+            if calculator_model not in SUBPLOT_ORDER.values():
+                raise ValueError(f"{calculator_model} model is not expected")
+
+            ax = axs[i] 
+            alphabetic_label = f"({chr(97 + i)})"
+            ax.set_title(
+                alphabetic_label, loc="left", fontsize=fontsize
+            ) 
+            label_file.write(
+                f"{alphabetic_label}: {calculator_model}\n"
+            )
+
+          
+            ax.set_prop_cycle(cycler(color=color_cycle))
+            ax.set_xlim(xlim)
+            ax.xaxis.set_major_locator(MultipleLocator(0.1))
+            ax.yaxis.set_major_locator(ticker.MaxNLocator(nbins=6))
[…]

### 2023-12-19 src/scripts/PlotGNNPhonons.py

@@ -1,14 +1,0 @@
-import sys
-import paths
-import json
-from PlotPhonons import *
-
-
-gnn_models = ["M3GNet","CHGNet","MACE"]
-dbname = paths.data / sys.argv[1]
-nstrains = int(sys.argv[2])
-structures = sys.argv[3:]
-
-for m in gnn_models:
-    for s in structures:
-        PlotPhonons(dbname, m, s, num_strains=nstrains)

### 2023-12-19 src/scripts/PlotPhonons.py

@@ -1,9 +1,11 @@
+import sys
 import paths
 import numpy as np
 import matplotlib.pyplot as plt
 from matplotlib.gridspec import GridSpec
+import argparse
+
 from ase.db import connect
-from scipy.interpolate import interp1d
 
 
 def format_xticklabels(labels, locs):
@@ -38,15 +40,84 @@ def get_color(value, cmap_name="viridis", vmin=-1.5, vmax=1.5):
     return cmap(norm(value))
 
 
-def PlotPhonons(dbname, model_name, structure_name, num_strains=None):
+def plot_default_phonons(bs, dos, info, ymaxlim=9.0, of="."):
     """
-    This function plots individual phonon bandstructures for different strains and
-    then all the strains on one plot with the optical branches semi-transparent.
+    Plot results from ASE phonon calculation.
+    """
+    fig = plt.figure(1, figsize=(7, 4))
+    ax = fig.add_axes([0.12, 0.07, 0.67, 0.85])
+    dosax = fig.add_axes([0.8, 0.07, 0.17, 0.85])
 
-    TODO: confirm strain is in bars and thus conver to MPa
+    xticks = bs.get_labels()[1].tolist()
+    xlabels = bs.get_labels()[2]
+
+    for xt in xticks:
+        ax.axvline(x=xt, color="lightgray")
+    ax.axhline(y=0.0, color="black")
+
+    q_points = bs.get_labels()[0]
+    energies_THz = EV_to_THz * bs.energies
+    for n in range(len(energies_THz[0, 1, :])):
+        omega = energies_THz[0, :, n]
+        ax.plot(q_points, omega, "k-", lw=2)
+
+    ax.set_title(
+        f"Volume: {info['vol']:.3f}Å^3\
+        Strain: {info['strain']:.2f}% \
+        Potential: {info['potname']}",
+        fontsize=10,
+    )
+
+    ax.set_xticks(xticks)
+    ax.set_xticklabels(xlabels)
+
+    ax.set_xlim((0.0, q_points[-1]))
+    ax.set_ylim((-2.0, ymaxlim))
+    ax.set_ylabel("Frequency ($\mathrm{THz}$)", fontsize=22)
+
+    dosax.fill_between(
+        dos.get_weights(),
+        EV_to_THz * dos.get_energies(),
+        y2=0,
+        color="grey",
+        edgecolor="k",
+        lw=1,
+    )
+
[…]

### 2023-12-19 src/scripts/Structures.py

@@ -250,7 +250,7 @@ def get_path_directions(path):
             start_coords = special_points[start_point]
             end_coords = special_points[end_point]
 
-            direction = start_coords - end_coords
+            direction = start_coords + end_coords
 
             direction_str = []
             for component in direction:
@@ -348,13 +348,13 @@ def spacegroup63_bandpath(
 
 # Get a path and return the directions
 def get_bandpath(structure, npoints=200):
-    if structure.info["name"] == "B19P":
+    if structure.info["structure_name"] == "B19P":
         path, directions = spacegroup11_bandpath(structure, npoints=npoints)
-    elif structure.info["name"] == "B19":
+    elif structure.info["structure_name"] == "B19":
         path, directions = spacegroup51_bandpath(structure, npoints=npoints)
-    elif structure.info["name"] == "B2":
+    elif structure.info["structure_name"] == "B2":
         path, directions = spacegroup221_bandpath(structure, npoints=npoints)
-    elif structure.info["name"] == "BCO":
+    elif structure.info["structure_name"] == "BCO":
         path, directions = spacegroup63_bandpath(structure, npoints=npoints)
     else:
         raise ValueError("Not valid structure")

### 2023-12-19 src/scripts/VisualizePhonons.py

@@ -1,0 +1,53 @@
+from ase.io.trajectory import Trajectory
+from ase.io import write
+
+def WriteModeWisual(
+    fileprefix,
+    phonons,
+    structure,
+    point="M",
+    branches=[0, 1],
+    repeat=(8, 8, 8),
+    amp_scale=25.7e-3,
+    mode_branches=[
+        ("TA1", "mode.0.traj"),
+        ("TA2", "mode.1.traj"),
+    ],
+    image_rotation="-36x,26.5y,-25z",
+    pts=100,
+    of=".",
+):
+    """
+    Creates trajectory files and then gif images of modes at a given symmetry point.
+
+    NOT TESTED with showyourwork flow
+    """
+
+    # Write modes for specific q-vector to trajectory files:
+    bandpath = structure.cell.bandpath(npoints=pts)
+    p = bandpath.special_points[point]
+    phonons.write_modes(
+        [l / 2 for l in p],
+        branches=branches,
+        repeat=repeat,
+        kT=amp_scale,
+        center=True,
+    )
+
+    for branch in mode_branches:
+        trajfile = phonons.name + "." + branch[1]
+        with Trajectory(trajfile, "r") as traj:
+            write(
+                os.path.join(of, f"{fileprefix}.{point}Point.Mode.{branch[0]}.gif"),
+                traj,
+                interval=50,
+                rotation=image_rotation,
+            )
+            write(
+                os.path.join(of, f"{fileprefix}.{point}Point.Mode.{branch[0]}.extxyz"),
+                traj,
+            )
+
+    # clean-up write modes
+    for branch in mode_branches:
+        os.remove(phonons.name + "." + branch[1])

### 2023-12-29 src/tex/ms.tex
New scripts, potential, rework Snakemake, move folders, added ms text.

@@
 \usepackage{longtable}
+\usepackage{subcaption}
 
@@
 These GNN potentials have leveraged large DFT-computed databases for training and provide a good approximation to the potential energy surfaces which enable inference on energies and forces in lattice and atomic dynamic calculations.
-We compare GNN potentials with the Embedded Atom Method (EAM) in predicting phonon properties of NiTi.
-We find that in general the EAM potentials are slightly more performant than the GN in that they appear to more accurately describe equation of state, phonon dispersions, and elastic constants.
+We compare GNN potentials with the Embedded Atom Method (EAM) and Modified EAM style potentials in predicting phonon properties of NiTi and PtTi.
+We find that in general the EAM/MEAM potentials are slightly more performant than the GNN in that they appear to more accurately describe equation of state, phonon dispersions, and elastic constants.
 However, the MACE potential shows significant promise as an exploritory potential for designing and characterizing more complex SMAs.
-Our findings reveal GNN potentials as a promising avenue for advancing SMA research and fine-tuning of pre-trained GNN models with more relevant DFT data (i.e., force-constants) may permit better outcomes.
+Our findings suggest GNN potentials as a promising initial avenue for advancing SMA research, in particular ternary and quaternary systems, upon fine-tuning of pre-trained GNN models with more relevant DFT data (i.e., force-constants) to permit better outcomes.
 \end{abstract}
@@
 In machine learning potentials, one describes the atoms and structure using descriptors of features, these are typically learned or prescribed.
+The features can be prescribed in isolation or combination of  structural, compositional, and property characteristics of the material data point.
+Futhermore, one may choose to fined new representations via unsurpervised learning.
+\par
 
-The main inquiry of this work is to asses the GNN interatomic potentials on binary SMA NiTi and PbTi primarily through the phonon predictions.
-Our analysis aims to provide a qualitative perspective based on previously studied phonon phenomena of these systems.
-The findings here will provide guidnace for future use of GNN interatomic potentials for the development other binary, and more so, ternary and quaternary SMAs.
-IMportantly, the use of pretrained base models for these systems acts as an initial seed and can be used to fine-tune future GNN potentials for better SMA disocovery works.
+The subsequent step once the featureization process is completed is to determine a suitable neural network archetcture which can learn the target property or properties given the feature inputs.
+In the case of atomistic simulations the target is prediction of the potential energy surface and corresponding gradients (i.e., forces).
+Two of the most early attempts to do this were the works of \citeauthor{Bartok2010} and \citeauthor{Thompson2015}, whereby projection onto higher-dimensional basis, such as bispectrum componenets of local neighbor density, is learned in order to describe the atomic environments.
+The energy per atom is then determined by these learned features by employing linear regression or non-linear Gaussian process regression.
+\par
+
+While these early attempts were successfully, they could only include structural details to learn new features from.
+Futhermore, the training process was typically restricted to a limited number of elemental constieunts and therefore for each chemical system and phases of interest one needs to provide a suitable training dataset.
+To move beyond such limitations, groups began using graph neural network (GNN) architectures and message passing scheme to represent the atomic structures as nodes and edges.
+This enabled two distinctions, first being that the nodes and edges corresponding to the atoms and bonds could include features, and secondly the learning could be done over a large structural and compositional space enabling a generalization\footnote{Typically these are referred to as Universal interatomic potentials or forcefields since they in theory generalize well to any chemical system or phase.}.
+The use of GNNs is efficient in representing structure-composition-property inputs to additional NN layers for predicting potential energy surfaces and gradients and was initially pursued in the CGCNN, MEGNet, and ALIGNN potentials.\cite{Choudhary2021,Chen2019,Xie2018}.
+The differences between the approaches are predominately in atomic and bond features, nearest-neighbor graph, and specific NN archetcture.
+\par
+
+The training of these early GNN utlized large density functional theory calculation repositories and mostly tuned the learning cost function to ground-state structure and energy.
+This resulted in mean-absolute-errors and other measures of performance that could be improved upon.\cite{Riebesell2023} More recently there have been significant improvements in both training data used and the GNN architecture that show promise for truly approaching a general interatomic potential suitable for static and dynamic atomistic calculations of multicomponents and multiphases.
+If succesfull, identifying and designing new potential SMA materials using GNN interatomic potentials could have tremendous impact on technology.
+Therefore, the primary goal is to provide a qualitative assement of predictions of the equations of state, phonon bandstructure, and elastic constants for various phases of NiTi and PtTi using three GNN potentials intended for general use: M3GNet, CHGNet, and MACE.\cite{Chen2022,Deng2023,Batatia2022} We note that the \cite{Riebesell2023} provides a summarized performance over these GNN over materials project database, but does not specifically address SMA and phonon characteristics.
 \par
@@
 \label{sec:workflow}
-The ground-state structures for the NiTi phases \textit{B2}, \textit{B19}, \textit{B19'}, and \textit{BCO} are generated using the atomic simulation environment (ASE) package (cite).
+The ground-state structures for the NiTi phases \textit{B2}, \textit{B19}, \textit{B19'}, and \textit{BCO} are generated using the atomic simulation environment (ASE) package \cite{Larsen2017}.
 This requires that the unit cell parameters, basis atoms, and spacegroup number be provided.
-These initial reference parameters for these structures have been taken from the materials project (cite specific DOIs for the structures).
-The energy and forces on the atoms are determined using the calculator interface routines which enables the assesment of different interatomic potentials for the same ASE structure.
-Similarly, for all EOS, phonon, and elastic constant claculations each step is performed for a specific interatomic potential via the ASE calculator interface.
-We not that this is done in different python computing environments using \href{https://github.com/showyourwork/showyourwork}{\showyourwork} package.\par
+The initial unit cell parameters and basis corrdinates for these structures have been taken from the materials project\cite{Jain2013} and several references \cite{Haskins2016,Kadkhodaei2018}.
+The initial structures for NiTi are shown in fig~\ref{fig:niti_structures} and are adapted for PtTi as well given same spacegroup symmetries.
+The energy and forces on the atoms are determined using the ASE calculator interfaces for EAM, MEAM, M3GNet, CHGNet, and MACE.\cite{Mutter2010,Zhong2011,Ko2015,Kim2017,Chen2022,Deng2023,Batatia2022} The EAM and MEAM potential interfaces are achieved through the use of LAMMPS python library\cite{Thompson2022} and ASE.
+The ALIGNN ASE calculator interface was also implemented, however, we found that the predictions for NiTi and PtTi to be unreliable.\par
 
-First the structures are optimized using ASE local optimization routines.
-The forces were converged to $10^{-4}\,\text{eV}/\AA$ and utlize the FIRE optimizater, but found other local optimization options found same ground-state structures.
-To ensure that the unit cell parameters were adjusted along with the atomic coordinates and that the crystal symmetry was not broken, we utlized the ASE constraint and filter routines during optimization of the structures.
+For each combination of potential/model, chemical system, and phase we perform calcualtions for the structure minimization, equation of state (EOS), phonons, and elastic constants.
+To determine the finite-temperature phase transitions we utlize an isothermal-isobaric ensemble via molecular dynamics simulations.
+The thermostating and barostating is done with Nose-Hoover and Parrinello-Rahman dynamics and the tempertaure is adjusted in a quasi-static manner which entails equlibrating the simulation for 10 ns while decrementing the temperature by 10 Kelvin; the pressure is kept constant at 1 bar.
+We note that in order to achieve the different computing environments required by the GNN potentials we leverage the ability to specify the  python computing environments while using \href{https://github.com/showyourwork/showyourwork}{\showyourwork} package.\par
+
+\begin{figure}[ht!]
+    \script{VisualizedStructures.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/NiTi_VisualizedStructures.png}
+        \caption{
+          Initial unit cells of used in this work for NiTi, (a) \textit{B2}(221), (b) \textit{B19}(11), (c) \textit{B19'}(51), and (d) \textit{BCO}(61).
+The spacegroup numbers are given in the parentheses.
+Also additional atoms (i.e.
+non-basis sites) are shown based on periodic boundary conditions.
+For PtTi (a-b) structures are used with different unit cell parameters and basis coordinates.
+        }
+        \label{fig:niti_structures}
+    \end{centering}
+\end{figure}
+
+Prior to any property calculations or dynamics the structures are optimized using the FIRE\cite{Bitzek2006} local optimizer with the forces converged to $2\cdot 10^{-3}\,\text{eV}/\AA$.
+We observed that other local optimization routines found same ground-state structures.
+To ensure that the unit cell parameters were adjusted along with the atomic coordinates and that the crystal symmetry was not broken, we utlized a constraint and filter during the optimization process.
 \par
 
-To determine the equation of state, we started with the obtained ground-state structures.
-To adjust the volume of the unit cell, we then apply an isotropic strain tensor to the cell and scale the atomic coordinates
+The equation of state of state can be obtained by varying the cell volume and calculating the total system energy without allowing for atomic relaxations.
+From the obtained ground-state structures the unit cell volume is adjusted by applying an isotropic strain,
 
@@
   \end{bmatrix}
-\end{equation}
+\end{equation},
 
-were $\epsilon$ is the fraction strain applied to the unit cell.
-To ensure that the strain is indeed isotropic we provide a quality check on the symmetry of the distorted structure.
-The isotropic strain is applied in compression and tension, from $-11\%$ to $11\%$, which gave a good range to asses the EOS curves and obtain fits.
-We fit the EOS to a standard polynomial which provides values for the equilibrium volume, energy, and bulk modulus.
-In order to compare respective equations of state amongst the different phases, we normalize the volume of each phase with respect to the obtained ground-state volume.
+to the cell and scaling the fractional atomic coordinates.
+Here $\epsilon$ is the fraction strain applied to the unit cell.
+To ensure that the strain is indeed isotropic we provide a quality check on the symmetry of the distorted structure using the spacegroup package \textit{spglib}.\cite{Togo2018} The isotropic strain is applied in compression and tension, from $-11\%$ to $11\%$, which gave a good range to asses the EOS curves and obtain fits.
+We fit the EOS to a third order inverse polynomial,
+
+\begin{equation}
+  \label{eq:eos_fit}
+  E(V) = c_0 + c_1 t + c_2 t^2 + c_3 t^3, \quad \text{where} \quad t = V^{-\frac{1}{3}}
+\end{equation},
+
+where the fit coefficients are used to determine equilibrium volume, system energy, and bulk modulus \cite{Alchagirov2003}.
+In order to compare respective equations of state amongst the different phases, we normalize the strained volume of each phase with respect to the obtained minimized ground-state volume.
 \par
@@
 
-It is observed that for all but the ALIGNN potential, that upon high tensile isotropic strain, the \textit{Pbcm} phase is the most stable. \par
-
 \begin{figure}[ht!]
@@
         \caption{
-            The equation of state curves for different \ce{NiTi} structures for (a) Mutter (EAM) (b) Zhong (EAM) (c) M3GNet (d) CHGNet (e) MACE and (f) ALIGNN interatomic potentials.
+            The equation of state curves for different \ce{NiTi} structures for (a) Mutter (EAM) (b) Zhong (EAM) (c) Ko (2NN-MEAM) (d) M3GNet (e) CHGNet (f) MACE interatomic potentials.
 The x-axis is the ratio of the isotropic strained unit cell volume, $V$,  with respect to the ground-state unit cell volume, $V_o$.
@@
 
-\begin{figure}[ht!]
+\begin{figure}[!htp]
     \script{PlotPhonons.py}
@@
 
-\begin{figure}[ht!]
-    \script{PlotEAMPhonons.py}
+\begin{figure}[!htp]
+    \script{PlotPhonons.py}
     \begin{centering}
@@
         \includegraphics[width=0.5\linewidth]{figures/NiTi_Zhong_B2_StrainsPhononBandstructures.png}
+        \vspace{1mm}
+        \includegraphics[width=0.5\linewidth]{figures/NiTi_Ko_B2_StrainsPhononBandstructures.png}
         \caption{
-           Phonon dispersion of B2 structure for the Mutter (top) and Zhong (bottom)  EAM potentials at different isotropic strains.
+           Phonon dispersion of B2 structure for the  (top) EAM Mutter, (middle) EAM Zhong, and (bottom) 2NN-MEAM Ko potentials at different isotropic strains.
 The optical branches have been made more transparent to emphasis the behavior in the TA$_1$ and TA$_2$ branches.
@@
 
-\begin{figure}[ht!]
-    \script{PlotGNNPhonons.py}
+\begin{figure}[!htp]
+    \script{PlotPhonons.py}
     \begin{centering}
@@
       \caption{
-        Phonon dispersion of B2 structure for the (a) M3GNet, (b) CHGNet, and (c) MACE potentials for different isotropic strains.
+        Phonon dispersion of B2 structure for the (top) M3GNet, (middle) CHGNet, and (bottom) MACE potentials for different isotropic strains.
 Similar transcparency is used as done in \ref{fig:mutter_zhong_phonon_b2}.
@@
 
-\subsection{PbTi: Equation of States}
+\subsection{PtTi: Equation of States}
 
@@
 
-\begin{figure}[ht!]                                                                                      
+\begin{figure}[!htp]                                                                                      
     \script{AppendixBZ.py}                                                                           
@@
       \vspace{1mm}
+      \includegraphics[width=0.5\linewidth]{figures/B19P_BrillouinZonePointsSampled.png}
+      \vspace{1mm}
       \includegraphics[width=0.5\linewidth]{figures/BCO_BrillouinZonePointsSampled.png}
       \vspace{1mm}                                                                                       
-      \caption{The irreducible Brillouin zones for (top) B2(\#221) (middle) B19/B19P(\#11), and (bottom) BCO(\#63) crystal structures}                                                                                  \label{fig:ibz}
+      \caption{The irreducible Brillouin zones for \textit{B2}(\#221), \textit{B19}(\#51), \textit{B19'}(\#11), and \textit{BCO}(\#63) crystal structures}                                                               \label{fig:ibz}
     \end{centering}                                                                                      
@@
 \variable{output/B19_SpecialSymmetryPointsBZ.tex}
+\variable{output/B19P_SpecialSymmetryPointsBZ.tex}
 \variable{output/BCO_SpecialSymmetryPointsBZ.tex}

### 2023-12-29 src/scripts/CalculationRunner.py

@@ -128,6 +128,7 @@ def main():
             "Zhong",
             "ZhongASE",
             "Ko",
+            "Kim",
             "M3GNet",
             "CHGNet",
             "MACE",
@@ -174,6 +175,7 @@ def main():
         help="Maximum strain for phonon calcs, fractional",
     )
     parser.add_argument("--eos_fit", type=str, default="sj", help="EOS fit function")
+    parser.add_argument("--chemsys", type=str, default="NiTi", help="Chemical system")
     parser.add_argument(
         "--calc_type",
         type=str,
@@ -187,7 +189,7 @@ def main():
     outfolder = paths.data / args.model.upper()
     os.makedirs(outfolder, exist_ok=True)
 
-    structure = get_structure(args.structure)
+    structure = get_structure(args.chemsys,args.structure)
     asecalc = get_ase_calculator(args.model)
     structure.calc = asecalc
     structure.info["model_name"] = args.model
@@ -212,7 +214,6 @@ def main():
         elastic_calc(db_path_file, structure, (args.model, asecalc))
 
     return None
-
 
 if __name__ == "__main__":
     main()

### 2023-12-29 src/scripts/Calculators.py

@@ -27,10 +27,14 @@ def get_ase_calculator(model="MutterASE"):
         ASE.calculators
 
     Notes:
-        For the EAM potentials there is the option to select the 
-        internal ASE calculator for EAM. This works but does not 
-        provide the stress there for we return a customEAM object 
+        For the EAM potentials there is the option to select the
+        internal ASE calculator for EAM. This works but does not
+        provide the stress there for we return a customEAM object
         which returns a zero stress array.
+
+        EAM/MEAM potentials are referenced via first author last
+        name, this means no specification of chemical system is
+        required.
     """
     if model == "Mutter":
         from ase.calculators.lammpslib import LAMMPSlib
@@ -55,6 +59,30 @@ def get_ase_calculator(model="MutterASE"):
     elif model == "ZhongASE":
         potential_file = str(paths.static / "NiTi_Zhong.eam.fs")
         asecalc = CustomEAM(potential=potential_file)
+
+    elif model == "Ko":
+        from ase.calculators.lammpslib import LAMMPSlib
+
+        library_file = str(paths.static / "NiTi_Ko.meam.library")
+        potential_file = str(paths.static / "NiTi_Ko.meam.potential")
+        cmds = [
+            "pair_style meam",
+            f"pair_coeff * *  {library_file} Ni Ti {potential_file} Ni Ti",
+        ]
+        amds = ["thermo_style custom etotal lx ly lz vol pxx pyy pzz pxy pxz pyz press"]
+        asecalc = LAMMPSlib(lmpcmds=cmds, amendments=amds, keep_alive=True)
+
+    elif model == "Kim":
+        from ase.calculators.lammpslib import LAMMPSlib
+
+        library_file = str(paths.static / "PtTi_Kim.meam.library")
+        potential_file = str(paths.static / "PtTi_Kim.meam.potential")
+        cmds = [
+            "pair_style meam",
+            f"pair_coeff * *  {library_file} Ni Ti {potential_file} Ni Ti",
+        ]
+        amds = ["thermo_style custom etotal lx ly lz vol pxx pyy pzz pxy pxz pyz press"]
+        asecalc = LAMMPSlib(lmpcmds=cmds, amendments=amds, keep_alive=True)
 
     elif model == "M3GNet":
         import matgl
@@ -83,15 +111,6 @@ def get_ase_calculator(model="MutterASE"):
[…]

### 2023-12-29 src/scripts/GenerateEquilibriumTable.py

@@ -73,7 +73,7 @@ def generate_structure_table(dbname, chemsys, output_filename):
         file.write("\\endfirsthead\n")
         file.write("\\caption[]{Equilibrium structures for NiTi. (Continued)}\\\\\n")
         file.write("\\hline\n")
-        file.write("Structure\n(Spacegroup \#) & Unit Cell \\\\\n")
+        file.write("Structure\\\\(Spacegroup \#) & Unit Cell \\\\\n")
         file.write("\\endhead\n")
         file.write("\\hline\n")
         file.write("\\endfoot\n")

### 2023-12-29 src/scripts/PlotEOS.py

@@ -28,7 +28,7 @@ def plot_eos(
     fontsize=8,
     tick_fontsize=6,
     legend_fontsize=8,
-    legend_ncol=3,
+    legend_ncol=5,
     label_notes="EOS_Labels.txt",
 ):
 

### 2023-12-29 src/scripts/Structures.py

@@ -13,9 +13,29 @@ def NiTi_B2_Structure(a=3.04, b=3.04, c=3.04, alpha=90.0, beta=90.0, gamma=90.0)
         cellpar=[a, b, c, alpha, beta, gamma],
         pbc=True,
     )
-    
+
     B2.info["structure_name"] = "B2"
     B2.info["chemsys"] = "NiTi"
+    return B2
+
+
+def PtTi_B2_Structure(a=3.17, b=3.17, c=3.17, alpha=90.0, beta=90.0, gamma=90.0):
+    """
+    Austenite phase.
+    Ref. Kadkhodaei, S.; van de Walle, Acta Materialia 2018, 147, 296–303.
+    https://doi.org/10.1016/j.actamat.2018.01.025.
+    Suppl. Mater. Appendix A
+    """
+    B2 = crystal(
+        ["Pt", "Ti"],
+        [(0, 0, 0), (0.5, 0.5, 0.5)],
+        spacegroup=221,
+        cellpar=[a, b, c, alpha, beta, gamma],
+        pbc=True,
+    )
+
+    B2.info["structure_name"] = "B2"
+    B2.info["chemsys"] = "PtTi"
     return B2
 
 
@@ -89,6 +109,26 @@ def NiTi_B19_Structure(a=4.09, b=2.86, c=4.59, alpha=90.0, beta=90.0, gamma=90.0
 
     B19.info["structure_name"] = "B19"
     B19.info["chemsys"] = "NiTi"
+    return B19
+
+
+def PbTi_B19_Structure(a=4.61, b=2.76, c=4.86, alpha=90.0, beta=90.0, gamma=90.0):
+    """
+    Pmma (spacegroup 51)
+    Ref. Kadkhodaei, S.; van de Walle, Acta Materialia 2018, 147, 296–303.
+    https://doi.org/10.1016/j.actamat.2018.01.025.
+    Suppl. Mater. Appendix A
+    """
+    B19 = crystal(
+        ["Pb", "Ti"],
+        [(0.25, 0.0, 0.6874), (0.25, 0.50, 0.1958)],
+        cellpar=[a, b, c, alpha, beta, gamma],
+        pbc=True,
+        spacegroup=51,
+    )
+
+    B19.info["structure_name"] = "B19"
+    B19.info["chemsys"] = "PbTi"
     return B19
 
 
@@ -203,18 +243,24 @@ def NiTi_R_Phase_Structure(
     return R_Phase
 
 
[…]

### 2024-01-02 src/tex/ms.tex
Some additional ms text. Added rule to show phonon calc code setting.

@@
 %% Packages
+\usepackage{orcidlink}
 \usepackage[utf8]{inputenc}
 \usepackage{fontspec}
+\usepackage{fontawesome5}
+\usepackage{amssymb}
 \usepackage{amsmath}
@@
 \usepackage{subcaption}
+\usepackage{listings}
+\lstset{
+  basicstyle=\ttfamily,
+  columns=fullflexible,
+  frame=single,
+  breaklines=true,
+  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
+}
 
-% Begin!
+% \journal{Computational Materials Science}
+
 \begin{document}
 
-% Title
 \title{Assessing Phonon Properties of Shape Memory Alloys using Graph Neural Network Potentials}
@@
 % Author list
-\author[1]{Stefan Bringuier\corref{cor1}}[role=Lead Researcher,orcid=0000-0001-6753-1437]
+\author[1]{Stefan Bringuier\corref{cor1}\orcidlink{0000-0001-6753-1437}}
 \ead{stefanbringuier@gmail.com}
@@
 In the case of atomistic simulations the target is prediction of the potential energy surface and corresponding gradients (i.e., forces).
-Two of the most early attempts to do this were the works of \citeauthor{Bartok2010} and \citeauthor{Thompson2015}, whereby projection onto higher-dimensional basis, such as bispectrum componenets of local neighbor density, is learned in order to describe the atomic environments.
+Two of the most early attempts to do this were the works of \citet{Bartok2010} and \citet{Thompson2015}, whereby projection onto higher-dimensional basis, such as bispectrum componenets of local neighbor density, is learned in order to describe the atomic environments.
 The energy per atom is then determined by these learned features by employing linear regression or non-linear Gaussian process regression.
@@
 Futhermore, the training process was typically restricted to a limited number of elemental constieunts and therefore for each chemical system and phases of interest one needs to provide a suitable training dataset.
-To move beyond such limitations, groups began using graph neural network (GNN) architectures and message passing scheme to represent the atomic structures as nodes and edges.
+To move beyond such limitations, groups began using graph neural network (GNN) architectures and message passing scheme to represent the atomic structures as nodes and edges \cite{Reiser2022}.
 This enabled two distinctions, first being that the nodes and edges corresponding to the atoms and bonds could include features, and secondly the learning could be done over a large structural and compositional space enabling a generalization\footnote{Typically these are referred to as Universal interatomic potentials or forcefields since they in theory generalize well to any chemical system or phase.}.
-The use of GNNs is efficient in representing structure-composition-property inputs to additional NN layers for predicting potential energy surfaces and gradients and was initially pursued in the CGCNN, MEGNet, and ALIGNN potentials.\cite{Choudhary2021,Chen2019,Xie2018}.
+The use of GNNs is efficient in representing structure-composition-property inputs to additional NN layers for predicting potential energy surfaces and gradients and was initially pursued in the CGCNN, MEGNet, and ALIGNN potentials \cite{Choudhary2021,Chen2019,Xie2018}.
 The differences between the approaches are predominately in atomic and bond features, nearest-neighbor graph, and specific NN archetcture.
@@
 The training of these early GNN utlized large density functional theory calculation repositories and mostly tuned the learning cost function to ground-state structure and energy.
-This resulted in mean-absolute-errors and other measures of performance that could be improved upon.\cite{Riebesell2023} More recently there have been significant improvements in both training data used and the GNN architecture that show promise for truly approaching a general interatomic potential suitable for static and dynamic atomistic calculations of multicomponents and multiphases.
+This resulted in mean-absolute-errors and other measures of performance that could be improved upon \cite{Riebesell2023}.
+More recently there have been significant improvements in both training data used and the GNN architecture that show promise for truly approaching a general interatomic potential suitable for static and dynamic atomistic calculations of multicomponents and multiphases.
 If succesfull, identifying and designing new potential SMA materials using GNN interatomic potentials could have tremendous impact on technology.
-Therefore, the primary goal is to provide a qualitative assement of predictions of the equations of state, phonon bandstructure, and elastic constants for various phases of NiTi and PtTi using three GNN potentials intended for general use: M3GNet, CHGNet, and MACE.\cite{Chen2022,Deng2023,Batatia2022} We note that the \cite{Riebesell2023} provides a summarized performance over these GNN over materials project database, but does not specifically address SMA and phonon characteristics.
+Therefore, the primary goal is to provide a qualitative assement of predictions of the equations of state, phonon bandstructure, and elastic constants for various phases of NiTi and PtTi using three GNN potentials intended for general use: M3GNet, CHGNet, and MACE \cite{Chen2022,Deng2023,Batatia2022}.
+We note that the \citet{Riebesell2023} provides a summarized performance over these GNN using the materials project database, but does not specifically address SMA and phonon characteristics.
 \par
@@
 This requires that the unit cell parameters, basis atoms, and spacegroup number be provided.
-The initial unit cell parameters and basis corrdinates for these structures have been taken from the materials project\cite{Jain2013} and several references \cite{Haskins2016,Kadkhodaei2018}.
+The initial unit cell parameters and basis corrdinates for these structures have been taken from the materials project \cite{Jain2013} and several references \cite{Haskins2016,Kadkhodaei2018}.
 The initial structures for NiTi are shown in fig~\ref{fig:niti_structures} and are adapted for PtTi as well given same spacegroup symmetries.
-The energy and forces on the atoms are determined using the ASE calculator interfaces for EAM, MEAM, M3GNet, CHGNet, and MACE.\cite{Mutter2010,Zhong2011,Ko2015,Kim2017,Chen2022,Deng2023,Batatia2022} The EAM and MEAM potential interfaces are achieved through the use of LAMMPS python library\cite{Thompson2022} and ASE.
-The ALIGNN ASE calculator interface was also implemented, however, we found that the predictions for NiTi and PtTi to be unreliable.\par
+The energy and forces on the atoms are determined using the ASE calculator interfaces for EAM, MEAM, M3GNet, CHGNet, and MACE.\cite{Mutter2010,Zhong2011,Ko2015,Kim2017,Chen2022,Deng2023,Batatia2022} The EAM and MEAM potential interfaces are achieved through the use of LAMMPS python library \cite{Thompson2022} and ASE.
+The ALIGNN ASE calculator interface was also implemented, however, we found that the predictions for NiTi and PtTi to be unreliable are not included in this study.\par
 
@@
 The thermostating and barostating is done with Nose-Hoover and Parrinello-Rahman dynamics and the tempertaure is adjusted in a quasi-static manner which entails equlibrating the simulation for 10 ns while decrementing the temperature by 10 Kelvin; the pressure is kept constant at 1 bar.
-We note that in order to achieve the different computing environments required by the GNN potentials we leverage the ability to specify the  python computing environments while using \href{https://github.com/showyourwork/showyourwork}{\showyourwork} package.\par
+We note that in order to achieve the different computing environments required by the GNN potentials we leverage the ability to specify the python computing environments using \href{https://github.com/showyourwork/showyourwork}{\showyourwork} package.\par
 
@@
 
-Prior to any property calculations or dynamics the structures are optimized using the FIRE\cite{Bitzek2006} local optimizer with the forces converged to $2\cdot 10^{-3}\,\text{eV}/\AA$.
-We observed that other local optimization routines found same ground-state structures.
-To ensure that the unit cell parameters were adjusted along with the atomic coordinates and that the crystal symmetry was not broken, we utlized a constraint and filter during the optimization process.
+Prior to any property calculations or dynamics the structures are optimized using the FIRE\cite{Bitzek2006} local optimizer with the forces converged to $2\cdot 10^{-3}$eV/\AA.
+We observed that other local optimization routines find the same ground-state structures.
+To ensure that the unit cell parameters were adjusted along with the atomic coordinates and that the crystal symmetry is not broken, we utlize a constraint and filter during the optimization process.
 \par
 
-The equation of state of state can be obtained by varying the cell volume and calculating the total system energy without allowing for atomic relaxations.
-From the obtained ground-state structures the unit cell volume is adjusted by applying an isotropic strain,
+\subsubsection{Equation of State}
+The EOS can be obtained by varying the cell volume and calculating the total system energy without allowing for atomic relaxations.
+Starting with the obtained ground-state structures the unit cell volume is adjusted by applying an isotropic strain,
 
@@
 Here $\epsilon$ is the fraction strain applied to the unit cell.
-To ensure that the strain is indeed isotropic we provide a quality check on the symmetry of the distorted structure using the spacegroup package \textit{spglib}.\cite{Togo2018} The isotropic strain is applied in compression and tension, from $-11\%$ to $11\%$, which gave a good range to asses the EOS curves and obtain fits.
-We fit the EOS to a third order inverse polynomial,
+To ensure that the strain is indeed isotropic we provide a quality check on the symmetry of the distorted structure using the spacegroup package \textit{spglib} \cite{Togo2018}.
+The isotropic strain is applied in compression and tension, from $-11\%$ to $11\%$, which gave a good range to characterize the EOS curves and obtain analytical fits.
+We fit the EOS using a third order inverse polynomial,
 
@@
 
-where the fit coefficients are used to determine equilibrium volume, system energy, and bulk modulus \cite{Alchagirov2003}.
-In order to compare respective equations of state amongst the different phases, we normalize the strained volume of each phase with respect to the obtained minimized ground-state volume.
+where the coefficients are used to determine equilibrium volume, system energy, and bulk modulus \cite{Alchagirov2003} at 0K.
+In order to compare respective EOS amongst the different phases, we normalize the strained volume of each phase with respect to the obtained minimized ground-state volume.
 \par
 
-Phonon calculations are done using the ASE phonons module, which uses the finite-difference method via small displacement of atoms in a reference cell.
-This requires the use of a supercell of the reference structure.
-In order to ensure convergence of this method, one needs to sweep different supercell sizes and how the displacement size impacts the phonon eigenvalues and vectors.
-This process was done initially for each interatomic potential and structure.
-Details on the settings are provided in the appendix.
+\subsubsection{Phonon Bandstructures}
+Phonon calculations are performend using finite-difference method via small displacement of atoms in a reference cell.\cite{Alfe2009} In the small displacement method, atoms in the primitive cell are slightly displaced to calculate forces on other atoms, constructing the force constant matrix, which, when used with the dynamical matrix, allows computing phonon frequencies at any q-vector in the Brillouin zone.
+This requires typically requires the use of a supercell of the original unit cell.
+In order to ensure convergence of this method, one needs to sweep different supercell sizes and atomic displacements.
+This process is done for each interatomic potential and structure.
+The supercells used in this work  have a repeat in any one direction ranging from $5-8$ and depends on the crystal system.
+The atomic displacement ranges from $0.03-0.26$\AA\,and was found to strongly depend on interatomic potential and crystal system.
+Details on the settings are provided in the appendix~\ref{sec:phonon_calc_config}.
 \par
 
-The ASE structures provide a routine for sampling the irreducible Brillouin zone for the specific structure, however, a significant amount of additional symmetry points are include in the band path.
-For this reason we only focus on high-symmetry points in the IBZ, given that the monoclinic and othorhombic phases are low symmetry structures.
-The exception is the cubic structure which we sample the full IBZ.
+In order to sample high-symmetry points in the irreducible Brillouin zone (IBZ) for specific structures, we sample from locations specified in the Bilbao crystallographic \textit{KVEC} utility\cite{Aroyo2014}, although the ASE structures object does provide a standardized routine for sampling the irreducible Brillouin zone for the specific structures.\cite{Setyawan2010}.
+We predominately focus on high-symmetry points in the IBZ that correspond to the corners and along directions known to exhibit phonon instabilities.
 The IBZ and points symmetry point locations can be seen in appendix \ref{sec:appx_ibz}.
@@
 
-To better understand dynamic instability in the various structures, same isotropic strain in eq.~\ref{eq:isotropic_strain} is used and respetive phonon bandstructures are calculated.
-These are then plotted to observed the onset or change of the instability of different branches.
+To better understand dynamic instability in the various structures, the same isotropic strain in eq.~\ref{eq:isotropic_strain} is used and the repsective phonon bandstructures are calculated.
+These are then plotted to observed the onset or change in stability or instability of phonon different branches.
 \par
 
-The elastic constants have been calculated using ....
+\subsubsection{Elastic Constants}
 
-\begin{itemize}
-\item Discuss EAM potentials Mutter and Zhong.
-\item GNN potentials M3GNet, CHGNet, MACE, ALIGNN
-\item EOS calcs
-\item Phonon calcs with strain.
+\subsubsection{Finite-Temperature Phase Transitions}
 
-  The calculation of the phonon dispersion is based on the dynamical matrix which in term is obtained from the fourier transform of the force-constant matrix.
-There are different numerical methods to obtain the force-constant matrix.
-Here we employ the finite-difference small displacement method (cite alfe), whereby atoms in a supercell are displaced by a small amount to ensure they remain within the harmonic (i.e., linear-response) regime.
-Then the force-constants are calculated.
-To obtain the dynamical matrix we sample reciprocal space on a given mesh within the irreducible Brillouin zone (BZ).
-Finally, we diagonalize the dynamical matrix along high-symmetry paths to obtain the phonon bands and eigenfrequencies.
-\par
+\subsection{Physics Based Potentials}
+\label{sec:physics_potentials}
 
-  Since the the finite-difference is a numerical approximation technique, one has to check convergence of the calculation for various models.
-In this work we have found that for most models and the avaible resources a supercell of $8\times8\times8$ is sufficient to converge, we also utlize the acoustic sum rule (add ref).
-The only stand-out for using that supercell is with use of the ALIGNN model, we found that supercells any larger than $5\times5\times5$ had considerable memory and we the required resources were unavilable.
-Despite this we find suitable convergence, althoug the ALIGNN model in general was not good at reproducing the NiTi phonon bandstructures.
-\par
+The EAM and MEAM potentials are
 
-  The displacement was choosen for all calculations to be $0.01\AA$ and the number of points along the BZ path in reciprocal space was 150.
-These values seemed to give the best trade-off of convergence for all the models.
-We also calculate the phonon does on a $30\times30\time30$ reciprocal space grid and use lorentz smoothing, the DOS can be found in the supplemental material.\par
-
-\item elastic constants.
-\end{itemize}
+\subsection{Graph Neural Network Potentials}
 
@@
 
-\bibliographystyle{naturemag}
+\bibliographystyle{elsarticle-num-names}
 \bibliography{references}
@@
 \appendix
+\subsection{Phonon Calculation Configurations}
+\label{sec:phonon_calc_config}
+
+The code used to generate the phonon configuration settings for each potential mode and structure are shown the table below. Settings are used for both NiTi and PtTi systems.
+
+\variable{output/Phonons_ConfigSettingsCode.tex}
+
 \subsection{Irreducible Brillouin Zones}

### 2024-01-02 src/scripts/WriteCodeToTex.py

@@ -1,0 +1,45 @@
+import paths
+import re
+import sys
+
+
+def extract_function(source_file, function_name):
+    with open(source_file, "r") as file:
+        code = file.read()
+
+    pattern = re.compile(rf'(def {function_name}\(.*?:\n(?:\s+.*\n)+)', re.MULTILINE)
+    match = pattern.search(code)
+
+    if match:
+        return match.group(1)
+    else:
+        raise ValueError(f"Function '{function_name}' not found in the file.")
+
+
+def write_to_tex(output_file, function_code):
+    """write extracted code in LaTeX listing environment"""
+    tex_content = (
+        f"\\begin{{lstlisting}}[language=Python]\n{function_code}\\end{{lstlisting}}\n"
+    )
+
+    with open(output_file, "w") as file:
+        file.write(tex_content)
+
+
+if __name__ == "__main__":
+    if len(sys.argv) != 4:
+        print(
+            "Usage: python script.py <source_python_file> <function_name> <output_tex_file>"
+        )
+        sys.exit(1)
+
+    source_python_file = paths.scripts / sys.argv[1]
+    function_name = sys.argv[2]
+    output_tex_file = paths.output / sys.argv[3]
+
+    try:
+        python_function_code = extract_function(source_python_file, function_name)
+        write_to_tex(output_tex_file, python_function_code)
+        print(f"Function {function_name} has been written to {output_tex_file}")
+    except ValueError as e:
+        print(e)

### 2024-01-04 src/tex/ms.tex
Changes to Calculators and Conda Env. Additions.

@@
 \label{sec:workflow}
+To promote reproducible research and align with findable, accessible, interoper-able and reusable (FAIR) principles\cite{Walsh2024}, we utilize \showyourwork python workflow\cite{Luger2021} and several open source python packages.
+Instructions for reproducing calculation data, figures, tables, and manuscript are provided in appendix~\ref{sec:appx_reproduce} and corresponding github repo\footnote{\url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP}}.
+\par
+
+\subsection{Calculation Details}
+\label{sec:calc_details}
 The ground-state structures for the NiTi phases \textit{B2}, \textit{B19}, \textit{B19'}, and \textit{BCO} are generated using the atomic simulation environment (ASE) package \cite{Larsen2017}.
@@
 To ensure that the unit cell parameters were adjusted along with the atomic coordinates and that the crystal symmetry is not broken, we utlize a constraint and filter during the optimization process.
+The optimized lattice parameters as well as basis atoms for all structures and models are provided in appendix \ref{sec:opt_structures}.
 \par
@@
 \subsubsection{Phonon Bandstructures}
-Phonon calculations are performend using finite-difference method via small displacement of atoms in a reference cell.\cite{Alfe2009} In the small displacement method, atoms in the primitive cell are slightly displaced to calculate forces on other atoms, constructing the force constant matrix, which, when used with the dynamical matrix, allows computing phonon frequencies at any q-vector in the Brillouin zone.
+Phonon calculations are performend using finite-difference method via small displacement (i.e., harmonic approximation) of atoms in a reference cell \cite{Alfe2009}.
+In the small displacement method, atoms in the primitive cell are slightly displaced to calculate forces on other atoms, constructing the force constant matrix, which, when used with the dynamical matrix, allows computing phonon frequencies at any q-vector in the Brillouin zone.
 This requires typically requires the use of a supercell of the original unit cell.
@@
 
+Similar to the phonon calculations the elastic constants within the harmonic approximation the elastic constants can be determined from linear-response of strain-stress calculations.
+Therefore, the calculation requires the use of crystal symmetry along with atomic displacements to determine respective elastic constant componenets.
+We use the \textit{Elastic}\cite{Jochym2018} python package which provides routines for generating the proper atomic displacement matrix given specific crystal spacegroup.
+The elastic constants are determined by solving a set of linear equations using singular value decomposition and given calculated strains and stress.
+The displacements occur with a min and max of 0.5\% and a total of 10 points are used.
+\par
+
 \subsubsection{Finite-Temperature Phase Transitions}
@@
 
+The M3GNet has been implemented in the MatGL framework and we utlize the \texttt{M3GNet-MP-2021.2.8-PES} pretrained model.
+
+For CHGNet, we have used the \texttt{v0.3} series pretrained model.
+
+%The MACE potential pretrained \texttt{2023-12-03-mace-128-L1_epoch-199.model} checkpoint file.
+
 \section{Results}
 \label{sec:results}
-
-\subsection{NiTi: Ground State Structures}
-\label{subsec:nitigs}
-
-%\comment{Display results in tables}
-
-\variable{output/Table_NiTi_Equilibrium_Structures.tex}
 
@@
 
+\subsubsection{Elastic Constants}
+
+\variable{output/Table_NiTi_Elastic_Constants.tex}
+
 \subsection{PtTi: Equation of States}
+\label{sec:ptti_eos}
 
@@
 \section*{Reproducibilty and Data Availability}
-This study was carried out using the reproducibility software \href{https://github.com/showyourwork/showyourwork}{\showyourwork}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}, create the figures, and compile the manuscript.
+This study was carried out using the reproducibility software \href{https://github.com/showyourwork/showyourwork}{\showyourwork}\cite{Luger2021}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}(if needed), create the figures, and compile the manuscript.
 Each figure caption contains two links: one to the dataset stored on zenodo used in the corresponding figure, and the other to the script used to make the figure (at the commit corresponding to the current build of the manuscript).
-The git repository associated to this study is publicly available at \url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP} and allows anyone to re-build the entire manuscript.
-The datasets generated by this paper are stored at \url{xyz}.
+The git repository associated to this study is publicly available at \url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP} and allows anyone to re-build the entire manuscript.\par
 
@@
 
+\newpage
 \appendix
+
+\subsection{Workflow Reproducibility}
+\label{sec:appx_reproduce}
+
+TODO
+
+\subsection{NiTi: Ground State Structures}
+\label{sec:opt_structures}
+
+\variable{output/Table_NiTi_Equilibrium_Structures.tex}
+
 \subsection{Phonon Calculation Configurations}

### 2024-01-04 src/scripts/Calculators.py

@@ -111,11 +111,18 @@ def get_ase_calculator(model="MutterASE"):
     elif model == "MACE":
         from mace.calculators import MACECalculator
 
+        #chkpoint = "2023-08-14-mace-universal.model"
+        chkpoint = "2023-12-03-mace-128-L1_epoch-199.model"
         asecalc = MACECalculator(
-            model_path=str(paths.static / "2023-08-14-mace-universal.model"),
+            model_paths=str(paths.static / chkpoint),
             device="cpu",
             default_dtype="float32",
         )
+
+        #Alternative API for newer MACE
+        #from mace.calculators import mace_mp
+        #asecalc = mace_mp() # return the default medium ASE calculator equivalent to mace_mp(model="medium")
+        #asecalc = mace_mp(model="large") # return a larger model
 
     elif model == "ALIGNN":
         from alignn.ff.ff import AlignnAtomwiseCalculator, default_path

### 2024-01-04 src/scripts/MinimizeStructure.py

@@ -4,7 +4,7 @@ from ase.constraints import UnitCellFilter
 from ase.spacegroup.symmetrize import FixSymmetry, check_symmetry
 
 
-def minimize_structure(structure, potential, write_db=None, fmax=2.0e-3, fstep=5000):
+def minimize_structure(structure, potential, write_db=None, fmax=5.0e-4, fstep=5000):
     """
     Minimize the ASE calculator using specific model and calculator.
 

### 2024-01-07 src/tex/ms.tex
Mods to rules, env, phonon calc, writing ms.

@@
 % Define document class
-\documentclass[preprint]{elsarticle}
+\documentclass[preprint,colorlinks=true,linkcolor=black,citecolor=black]{elsarticle}
 \usepackage{showyourwork}
@@
 \label{sec:intro}
-Shape memory alloys derive their unique displacive phase transformations from the dynamic instablity that occurs during the application of temperature, strain, or stress.
-The origin of this instability occurs due to softening or symmetry breaking of the phonon modes in the phase.
-Therefore, improved understanding of the phonon characteristics of the different phases that can occur for a given SMA is particular important.
-One of the challenges is in performing such calculations, which typically employ density functional theory in order to determine the atomic forces, which in turn are used to construct the dynamical matrix.
-However, the use of classical interatomic potentials can alleviate some of this computational burden.
+
+The design and development of shape memory alloys (SMA) via \textit{in-silico} methods would provide great value in advancing structural, medical, and other technologies.
+In order to enable such activity would require sampling from very large structural and compositional spaces and evaluating material properties indicative of displacive phase transformation phenomena.
+The use of density functional theory calculations to obtain properties such as equations of state, phonon disperisons, and elastic constants, can be done, however, such calculations are computationally demanding and are potentially prohibitive for smaller groups trying to develop new SMAs.
+To alleviate such restriction, one can resort to classical interatomic potentials which can greatly improve computational cost in  evaluating the energies and forces more.
+However, the immediate challenge is to whether a suitable interatomic potential exist given a chemical system, and moreover, how transferable is it to different compositions and phases.
 \par
@@
 
-The subsequent step once the featureization process is completed is to determine a suitable neural network archetcture which can learn the target property or properties given the feature inputs.
+The subsequent step once the featureization process is completed is to determine a suitable neural network (NN) architecture which can learn the target property or properties given the feature inputs.
 In the case of atomistic simulations the target is prediction of the potential energy surface and corresponding gradients (i.e., forces).
-Two of the most early attempts to do this were the works of \citet{Bartok2010} and \citet{Thompson2015}, whereby projection onto higher-dimensional basis, such as bispectrum componenets of local neighbor density, is learned in order to describe the atomic environments.
+Two of the most early attempts to do this were the works of \citet{Bartok2010} and \citet{Thompson2015}, whereby projection onto higher-dimensional basis, such as bispectrum components of local neighbor density, is learned in order to describe the atomic environments.
 The energy per atom is then determined by these learned features by employing linear regression or non-linear Gaussian process regression.
+We note that these are just two of the early examples but similar potentials have been developed in earlier, in parallel, or recently by various groups \cite{Behler2007,Wang2018,Deringer2019,Lee2019}.
 \par
 
-While these early attempts were successfully, they could only include structural details to learn new features from.
-Futhermore, the training process was typically restricted to a limited number of elemental constieunts and therefore for each chemical system and phases of interest one needs to provide a suitable training dataset.
+While these early attempts using ML models were successfully, they may only include limited representations of structural and/or chemical details to learn new features from which could limit transferability.
+Futhermore, the training process was typically restricted to a limited number of elemental constieunts of the target material system.
+As a consequence,  for each chemical system and phases of interest one needs to provide a suitable training dataset.
 To move beyond such limitations, groups began using graph neural network (GNN) architectures and message passing scheme to represent the atomic structures as nodes and edges \cite{Reiser2022}.
@@
 If succesfull, identifying and designing new potential SMA materials using GNN interatomic potentials could have tremendous impact on technology.
-Therefore, the primary goal is to provide a qualitative assement of predictions of the equations of state, phonon bandstructure, and elastic constants for various phases of NiTi and PtTi using three GNN potentials intended for general use: M3GNet, CHGNet, and MACE \cite{Chen2022,Deng2023,Batatia2022}.
+Therefore, the primary goal is to provide a qualitative assessment of predictions of the equations of state, phonon bandstructure, and elastic constants for various phases of NiTi and PtTi using three GNN potentials intended for general use: M3GNet, CHGNet, and MACE \cite{Chen2022,Deng2023,Batatia2022}.
 We note that the \citet{Riebesell2023} provides a summarized performance over these GNN using the materials project database, but does not specifically address SMA and phonon characteristics.
@@
 To promote reproducible research and align with findable, accessible, interoper-able and reusable (FAIR) principles\cite{Walsh2024}, we utilize \showyourwork python workflow\cite{Luger2021} and several open source python packages.
-Instructions for reproducing calculation data, figures, tables, and manuscript are provided in appendix~\ref{sec:appx_reproduce} and corresponding github repo\footnote{\url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP}}.
-\par
+Instructions for reproducing calculated data, figures, tables, and manuscript are provided in appendix~\ref{sec:appx_reproduce} and corresponding github repo\footnote{\url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP}}.\par
 
@@
 \subsubsection{Phonon Bandstructures}
-Phonon calculations are performend using finite-difference method via small displacement (i.e., harmonic approximation) of atoms in a reference cell \cite{Alfe2009}.
+
+Shape memory alloys derive their unique displacive phase transformations from the dynamic instability that occurs during the application of temperature, strain, or stress.
+The origin of this instability occurs due to softening or symmetry breaking of the phonon modes in the phase.
+Therefore, improved understanding of the phonon characteristics of the different phases that can occur for a given SMA is particular important.
+One of the challenges is in performing such calculations, which typically employ density functional theory (DFT) in order to determine the atomic forces, which in turn are used to construct the dynamical matrix.
+However, classical interatomic potentials can alleviate the computational burden.
+\par
+
+Phonon calculations are performed using finite-difference method via small displacement (i.e., harmonic approximation) of atoms in a reference cell \cite{Alfe2009}.
 In the small displacement method, atoms in the primitive cell are slightly displaced to calculate forces on other atoms, constructing the force constant matrix, which, when used with the dynamical matrix, allows computing phonon frequencies at any q-vector in the Brillouin zone.
@@
 
-The EAM and MEAM potentials are
+The EAM and MEAM potentials were developed to go beyond pair interatomic potentials to treat point and extended defects in metal and alloys systems.
+The motivation was taken from the physical results of  density functional theory and tight-binding methods, namely to add to the pair potentials a energy functional which depends on the local electron density which in turn is affected by the atomic environment.
+These type of interatomic potentials are broadly referred to as EAM or MEAM, with the later included direction (i.e.
+angle) dependent interactions.
+\par
+
+The functional form using during fitting proceed depends on the complexity of the material system but the total energy is typically written as,
+
+\begin{equation}
+  \label{eq:eam}
+  E = \sum_{i\neq j}^N V_{ij}(r_{ij}) + \sum_i^N F_i\left(\rho(r_i)\right)
+\end{equation}
+
+where $V_{ij}$ is the pair potential, $F_i$, the functional, and $\rho$ the density function that depends on the environment around $r_i$.
+The MEAM potential extends this by including three-body terms in the $\rho$ to capture features such as covalent bonding.
+The EAM and MEAM potentials have been particular successfully in predicting structure and properties of metals and alloys, including defect formation/energies, phase transitions, grain boundary stability/dynamics.
+One particular advantage of both EAM and MEAM, is their analytical or tabulated forms allow for quick numerical evaluation, with the primary reason being that the energies and forces on a atom are accurately determined within a relatively small interaction volume (i.e., 3-4 \AA interaction cutoff).
+\par
+
+For application of EAM to shape memory alloys, and in particular NiTi, \citet{Mutter2010} initially devised an Finnis-Sinclair style EAM potential to capture the structural phase transitions between \textit{B19'} and \textit{B2} phases.
+It was latter shown by \cite{Zhong2011} that some adjustments to the EAM parameters was needed to properly capture the finite-temperature displacement dynamics for the martensitic transition \cite{Gur2017}.
+More recently \citet{Ko2015} developed a second nearest neighbor(2NN) MEAM parameterization of NiTi that demonstrated very good predictive performance and improvements over ealier potentials.
 
@@
 
-The M3GNet has been implemented in the MatGL framework and we utlize the \texttt{M3GNet-MP-2021.2.8-PES} pretrained model.
+The use of ML potentials for SMA has been demonstrated by \citet{Tang2022}.
+In their work, they utilized the DeePMD-Kit\cite{Wang2018} neural network framework to parameterize a descriptor matrix  from atomic configurations that preserves permutations and  symmetries.
+The descriptors are then which feed into a multilayer perceptron to calculate energies and forces.
+The neural network interatomic potential demonstrated good accuracy of ground-state properties with respect to DFT results and showed agreement with experiments regarding temperature, stress, and defect induced martensitic transformation.
+\paragraphmark
+
+In the work of \citet{Tang2022} ground-state predictions for \textit{B2}, \textit{B19'}, \textit{B19}, and \textit{BCO}\footnote{Referred to as \textit{B33} prototype in \citet{Tang2022}} were provided and showed good agreement.
+However, we note that deviation from Ni-Ti 1:1 stoichiometry or addition of a third species was not investigated nor the inclusion of additional alloying species.
+\par
+
+The pre-trained GNN interatomic potentials explored in this work have been parameterized over a large fraction of the interatomic potential ($>$ 80 elements) and therefore are not restricted to specific chemistries or stoichiometries.
+They training data also includes various phases for a given chemical system.
+This permits great flexibility.
+We do not that fine-tuning, the process of adjusting GNN weights and biases using additional training data, can be performed with the GNN studied in this work.
+However, this is out of the scope for this manuscript, given resources and time needed to generate such training data for SMAs.
+\par
+
+The M3GNet is a GNN that utilizes a 3-body bond-update procedure to more accurately capture bond-order environment changes.
+The M3GNet has been implemented in the MatGL framework and we utilize the \texttt{M3GNet-MP-2021.2.8-PES} pretrained model.
 
@@
 
-The thermodynamic equation of state (EOS) is a standard assement of the phase stability of materials at a give condition.
-In the context of atomic calculations the equation of state can be determined at ground-state (i.e., 0K) to asses the phase stability of the different crystals.
-The equation of state can be represented as a pressure vs.
-volume or more commonly as the energy per atom  vs the ratio of the isotropically strained volume to that of the equilibrium volume.
-One criteria for a interatomic potential is that it accurately refelcts the EOS curves for different phases.
-This entails the construction of a convex hull which represents the minimum energy curve accross the structures for the volume strain conditions.
-\par
-
-The EOS curves for the \ce{NiTi} structures \textit{B2}, \textit{B19}, \textit{B19'}, \textit{B33}, and \textit{Pbcm}  for the different potentials studied in this work are shown in fig.~\ref{fig:eos}.
+The EOS curves for the \ce{NiTi} structures \textit{B2}, \textit{B19}, \textit{B19'}, \textit{BCO} for the different potentials studied in this work are shown in fig.~\ref{fig:eos}.
 The two EAM potentials, show that at 0 strain ($V/V_o = 1.0$) both the \textit{B19P} and \textit{B33} phases are stable.
@@
 
-\section*{Reproducibilty and Data Availability}
+\section*{Reproducibility and Data Availability}
 This study was carried out using the reproducibility software \href{https://github.com/showyourwork/showyourwork}{\showyourwork}\cite{Luger2021}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}(if needed), create the figures, and compile the manuscript.
@@
 \begin{figure}[!htp]
-    \script{AppendixBZ.py}                                                                           
-    \begin{centering}                                                                                    
-      \includegraphics[width=0.5\linewidth]{figures/B2_BrillouinZonePointsSampled.png} 
+  \centering
+  \begin{subfigure}[b]{0.45\linewidth}
+      \includegraphics[width=\linewidth]{figures/B2_BrillouinZonePointsSampled.png}
+      \caption{}
+      \label{fig:B2}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.45\linewidth}
+      \includegraphics[width=\linewidth]{figures/B19_BrillouinZonePointsSampled.png}
+      \caption{}
+      \label{fig:B19}
+  \end{subfigure}
   \vspace{1mm}
-      \includegraphics[width=0.5\linewidth]{figures/B19_BrillouinZonePointsSampled.png}
-      \vspace{1mm}
-      \includegraphics[width=0.5\linewidth]{figures/B19P_BrillouinZonePointsSampled.png}
-      \vspace{1mm}
-      \includegraphics[width=0.5\linewidth]{figures/BCO_BrillouinZonePointsSampled.png}
-      \vspace{1mm}                                                                                       
-      \caption{The irreducible Brillouin zones for \textit{B2}(\#221), \textit{B19}(\#51), \textit{B19'}(\#11), and \textit{BCO}(\#63) crystal structures}                                                               \label{fig:ibz}
-    \end{centering}                                                                                      
+  \begin{subfigure}[b]{0.45\linewidth}
+      \includegraphics[width=\linewidth]{figures/B19P_BrillouinZonePointsSampled.png}
+      \caption{}
+      \label{fig:B19P}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[b]{0.45\linewidth}
+      \includegraphics[width=\linewidth]{figures/BCO_BrillouinZonePointsSampled.png}
+      \caption{}
+      \label{fig:BCO}
+  \end{subfigure}
+  \caption{The irreducible Brillouin zones for \textit{B2}(\#221), \textit{B19}(\#51), \textit{B19'}(\#11), and \textit{BCO}(\#63) crystal structures}
+  \label{fig:ibz}
 \end{figure}
+
+\pagebreak
 

### 2024-01-07 src/scripts/CalculatePhonons.py

@@ -1,18 +1,13 @@
-import os
-import numpy as np
-import pickle
 import paths
 
 from ase.spacegroup import get_spacegroup
 from ase.spacegroup import crystal
 
-# from ase.constraints import StrainFilter, UnitCellFilter
-from ase.spacegroup.symmetrize import FixSymmetry, check_symmetry
 from ase.phonons import Phonons
 from ase.db import connect
 from ase.spectrum.band_structure import BandStructure
 
-from Structures import *
+from Structures import get_bandpath
 from Config import get_phonon_config
 from PlotPhonons import plot_default_phonons
 
@@ -45,16 +40,16 @@ def get_phonons(
         lorentz_width (float, optional): Lorentzian broadening width. Defaults to 4.5e-4.
 
     Returns:
-        tuple: Tuple containing band structure, density of states, and mode vectors.
+        tuple: Tuple containing band structure, density of states, and dynamical matrix (ndarray)
     """
 
     model = potential[0].upper()
     structure_name = structure.info["structure_name"]
-    phfolder = str(paths.data / model / structure_name) + "-phonon-calc-files"
+    phfolder = str(paths.data / model / structure_name)
 
     phonons = Phonons(
         structure,
-        name=phfolder,
+        name=phfolder + "_PhononCalcFiles",
         calc=potential[1],
         supercell=supercell,
         delta=displacement,
@@ -64,17 +59,18 @@ def get_phonons(
     phonons.run()
     phonons.read(acoustic=True)
     phonons.clean()
-    omegas, mode_vecs = phonons.band_structure(
-        bandpath.kpts, modes=True, born=False, verbose=False
+
+    dyn_mat = phonons.D_N
+
+    # Calculate the band structure for provide path
+    # NOTE: No longer need modes here, will use dyn_mat if needed
+    omegas = phonons.band_structure(
+        bandpath.kpts, modes=False, born=False, verbose=False
     )
     bs = BandStructure(bandpath, energies=omegas[None])
     dos = phonons.get_dos(kpts=doskpts).sample_grid(npts=dospts, width=lorentz_width)
 
-    # Pickle Phonons for later, i.e., mode analysis, different band path.
-    with open(os.path.join(phfolder, "PhononsObject.pkl"), "wb") as file:
-        pickle.dump(phonons, file)
-
-    return bs, dos, mode_vecs
+    return bs, dos, dyn_mat
 
 
 def calculate_phonons(
@@ -82,15 +78,11 @@ def calculate_phonons(
     structure_name,
     potential,
[…]

### 2024-01-07 src/scripts/Calculators.py

@@ -130,6 +130,12 @@ def get_ase_calculator(model="MutterASE"):
         alignn_params = default_path()
         asecalc = AlignnAtomwiseCalculator(path=alignn_params, device="cpu")
 
+    elif model == "DeepMD":
+        from deepmd.calculator import DP
+
+        model_params = str(paths.static / "NiTi_Tang.pb")
+        asecalc= DP(model=model_params)
+        
     return asecalc
 
 

### 2024-01-08 src/tex/ms.tex
Some adjustments

@@
 
-\title{Assessing Phonon Properties of Shape Memory Alloys using Graph Neural Network Potentials}
+\title{Qualitative Assessesment of Graph Neural Network Potentials for  Phonon Properties of Shape Memory Alloys using Graph Neural Network Potentials}
 
@@
 If succesfull, identifying and designing new potential SMA materials using GNN interatomic potentials could have tremendous impact on technology.
-Therefore, the primary goal is to provide a qualitative assessment of predictions of the equations of state, phonon bandstructure, and elastic constants for various phases of NiTi and PtTi using three GNN potentials intended for general use: M3GNet, CHGNet, and MACE \cite{Chen2022,Deng2023,Batatia2022}.
+Therefore, the primary goal is to provide a qualitative assessment of pre-trained GNN interatomic potentials that are promoted as being transferable across compositions and phases.
+We specfically focus on the predictions for equations of state, phonon bandstructure, and elastic constants for various phases of NiTi and PtTi using three GNN potentials: M3GNet, CHGNet, and MACE \cite{Chen2022,Deng2023,Batatia2022}.
 We note that the \citet{Riebesell2023} provides a summarized performance over these GNN using the materials project database, but does not specifically address SMA and phonon characteristics.
@@
 
+One way to understand the phase stability for a given composition is to compare the cohesive energy, which is defined as,
+
+\begin{equation}
+  \label{eq:cohesive_energy}
+  E_{c} = \frac{E_{\text{total}}}{N}
+\end{equation}
+
+With $E_{\text{total}}$ the total energy of the system and $N$ the number of atoms.
+In fig.~\ref{fig:ecoh} the bar chart for different phases and interatomic potentials models are show.
+We compare these to the predicted cohesive energy for The Materials Project (MP) entry \texttt{MP-571} which corresponds to the \textit{B2} phase \cite{MP--571}.
+In the figure we see that the EAM and MEAM potentials show significantly lower cohesive energy compared to the GNN potentials.
+The stark increase in binding energy for the GNN is surprising as the EAM and MEAM potentials are known to reproduce previous experimental and DFT data \cite{Haskins2016,Ko2015}.
+The GNNs predict a more strongly bound and stable \textit{B2} structure but all are consistent with the entry for The MP and calculation details can be found in ref.~\cite{MP--571}.
+\par
+
+\begin{figure}[ht!]
+    \script{PlotCohesiveEnergy.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/NiTi_CohesiveEnergyPlot.png}
+        \caption{
+          Comparison of predicted cohesive energy for respective ground-state structure and model. As a reference we use The Materials Project entry for the \ce{NiTi} \textit{B2} entry \cite{MP--571}.
+        }
+        \label{fig:ecoh}
+    \end{centering}
+\end{figure}
+
 \subsection{NiTi: Equation of States}
@@
 At compressive strain beyond $\approx 0.94$ the \textit{B2} phase is stable.
-
-\begin{figure}[ht!]
-    \script{PlotCohesiveEnergy.py}
-    \begin{centering}
-        \includegraphics[width=\linewidth]{figures/NiTi_CohesiveEnergyPlot.png}
-        \caption{
-          Comparison of predicted cohesive energy for respective ground-state structure and model.
-        }
-        \label{fig:ecoh}
-    \end{centering}
-\end{figure}
 

### 2024-01-08 src/scripts/PlotEOS.py

@@ -1,11 +1,10 @@
 import sys
 import numpy as np
-import os
+import warnings
 import paths
 
 import matplotlib.pyplot as plt
 import matplotlib.ticker as ticker
-from matplotlib.ticker import MultipleLocator, AutoLocator
 from cycler import cycler
 import seaborn as sns  # For color palette
 from ase.db import connect
@@ -52,7 +51,8 @@ def plot_eos(
     with open(label_file_path, "w") as label_file:
         for i, calculator_model in enumerate(calculator_models):
             if calculator_model not in SUBPLOT_ORDER.values():
-                raise ValueError(f"{calculator_model} model is not expected")
+                warnings.warn(f"{calculator_model} model results not used/expected", UserWarning)
+
 
             ax = axs[i] 
             alphabetic_label = f"({chr(97 + i)})"

### 2024-01-09 src/tex/ms.tex
Add info and reworking plotting.

@@
         \caption{
-          Comparison of predicted cohesive energy for respective ground-state structure and model. As a reference we use The Materials Project entry for the \ce{NiTi} \textit{B2} entry \cite{MP--571}.
+          Comparison of predicted \ce{NiTi} cohesive energy for respective ground-state structure and model.
+As a reference we use The Materials Project entry for the \ce{NiTi} \textit{B2} entry \cite{MP--571}.
         }
-        \label{fig:ecoh}
+        \label{fig:ecoh_niti}
+    \end{centering}
+\end{figure}
+
+\begin{figure}[ht!]
+    \script{PlotCohesiveEnergy.py}
+    \begin{centering}
+        \includegraphics[width=\linewidth]{figures/PtTi_CohesiveEnergyPlot.png}
+        \caption{
+          Comparison of predicted \ce{PtTi} cohesive energy for respective ground-state structure and model.
+        }
+        \label{fig:ecoh_ptti}
     \end{centering}

### 2024-01-09 src/scripts/CalculateElastic.py

@@ -43,6 +43,7 @@ def serial_calculate(systems, calculator):
 
 def calculate_elastic_constants(
     dbname,
+    chemsys,
     structure_name,
     potential,
     npoints=10,
@@ -55,6 +56,7 @@ def calculate_elastic_constants(
 
     Args:
         dbname (string): The path of the ASE database file.
+        chemsys (string): Chemical system, e.g., NiTi, PtTi, NiAl3
         structure_name (string): The name of the structure in the ASE database file.
         potential (string): The name of the potential style.
         npoints (int,10):  Optional, sets the number of displacement points.
@@ -72,28 +74,29 @@ def calculate_elastic_constants(
     potname = potential[0]
     calculator = potential[1]
 
-    for entry in db.select(structure_name=structure_name, model_name=potname):
-        spg = entry.spacegroup
-        structure = crystal(entry.toatoms(), spacegroup=spg)
-        # calculator.clean = potential[1].reset
-        structure.calc = calculator
+    entry = db.get(
+        chemsys=chemsys, structure_name=structure_name, model_name=potname, relaxed=True
+    )
+    
+    spg = entry.spacegroup
+    structure = crystal(entry.toatoms(), spacegroup=spg)
+    # calculator.clean = potential[1].reset
+    structure.calc = calculator
 
-        systems = get_elementary_deformations(structure, n=npoints, d=displacement)
+    systems = get_elementary_deformations(structure, n=npoints, d=displacement)
 
-     
-        result = serial_calculate(systems, calculator)
+    result = serial_calculate(systems, calculator)
 
-        ordering = get_cij_order(structure)
-        cij, _ = get_elastic_tensor(structure, systems=result)
-        cij /= units.GPa
+    ordering = get_cij_order(structure)
+    cij, _ = get_elastic_tensor(structure, systems=result)
+    cij /= units.GPa
 
-        C = dict(zip(ordering, cij))
+    C = dict(zip(ordering, cij))
 
-
-        if update:
-            db.update(entry.id, data={"elastic_constants": C})
-        else:
-            return C
+    if update:
+        db.update(entry.id, data={"elastic_constants": C})
+    else:
+        return C
 
     return None
 
@@ -102,12 +105,22 @@ if __name__ == "__main__":
     # Testing data
     print("!!! TESTING ElasticCalculation.py !!!")
     from Calculators import *
+
     calculator = get_ase_calculator("M3GNet")
[…]

### 2024-01-09 src/scripts/CalculatePhonons.py

@@ -75,6 +75,7 @@ def get_phonons(
 
 def calculate_phonons(
     dbname,
+    chemsys,
     structure_name,
     potential,
     strain=[
@@ -91,6 +92,7 @@ def calculate_phonons(
 
     Args:
         dbname (str): Name of the database for storing results.
+        chemsys (str): The chemical system, e.g. NiTi, PtTi, NiAl3
         structure_name (str): Name of the structure to analyze.
         potential (tuple): Tuple containing potential model name and calculator object.
         strain (list of float, optional): List of fractional strain values to apply. Defaults to pre-defined list.
@@ -117,7 +119,10 @@ def calculate_phonons(
 
     supercell, displacement, bspts = get_phonon_config(structure_name, potential[0])
 
-    entry = db.get(model_name=potname, structure_name=structure_name)
+    # Extract relaxed-structure from database
+    entry = db.get(
+        chemsys=chemsys, model_name=potname, structure_name=structure_name, relaxed=True
+    )
     spg = entry.spacegroup
     cell_relaxed = entry.cell
 
@@ -161,9 +166,7 @@ def calculate_phonons(
                 bandstructure,
                 dos,
                 {
-                    "chemsys": "".join(
-                        sorted(set(mod_structure.get_chemical_symbols()))
-                    ),
+                    "chemsys": chemsys,
                     "vol": volume,
                     "strain": e * 100.0,
                     "structure": structure.info["structure_name"],

### 2024-01-09 src/scripts/Calculators.py

@@ -44,9 +44,12 @@ def get_ase_calculator(model="MutterASE"):
         - Zhong, Y.; Gall, K.; Zhu, T. Atomistic Study of Nanotwins in NiTi Shape Memory Alloys. Journal of Applied Physics 2011, 110 (3), 033532. https://doi.org/10.1063/1.3621429.
         - Kim, J.-S.; et al.  Calphad 2017, 59, 131–141. https://doi.org/10.1016/j.calphad.2017.09.005.
         - Ko, W.-S.; Grabowski, B.; Neugebauer, J. Phys. Rev. B 2015, 92 (13), 134107. https://doi.org/10.1103/PhysRevB.92.134107.
+        - Kavousi, S.; et al. Modelling Simul. Mater. Sci. Eng. 2019, 28 (1), 015006. https://doi.org/10.1088/1361-651X/ab580c.
         - Chen, C.; Ong, S. P. Nat Comput Sci 2022, 2 (11), 718–728. https://doi.org/10.1038/s43588-022-00349-3.
         - Deng, B.; et al. Nat Mach Intell 2023, 5 (9), 1031–1041. https://doi.org/10.1038/s42256-023-00716-3.
         - Batatia, I.; et al. MACE: Higher Order Equivariant Message Passing Neural Networks for Fast and Accurate Force Fields; 2022.
+        - Tang, H.; et al. Acta Materialia 2022, 238, 118217. https://doi.org/10.1016/j.actamat.2022.118217.
+
     """
     if model == "Mutter":
         from ase.calculators.lammpslib import LAMMPSlib
@@ -77,6 +80,18 @@ def get_ase_calculator(model="MutterASE"):
 
         library_file = str(paths.static / "NiTi_Ko.meam.library")
         potential_file = str(paths.static / "NiTi_Ko.meam.potential")
+        cmds = [
+            "pair_style meam",
+            f"pair_coeff * *  {library_file} Ni Ti {potential_file} Ni Ti",
+        ]
+        amds = ["thermo_style custom etotal lx ly lz vol pxx pyy pzz pxy pxz pyz press"]
+        asecalc = LAMMPSlib(lmpcmds=cmds, amendments=amds, keep_alive=True)
+
+    elif model == "Kavousi":
+        from ase.calculators.lammpslib import LAMMPSlib
+
+        library_file = str(paths.static / "NiTi_Kavousi.meam.library")
+        potential_file = str(paths.static / "NiTi_Kavousi.meam.potential")
         cmds = [
[…]

### 2024-01-09 src/scripts/Config.py

@@ -1,27 +1,3 @@
-ORDER = {
-    "Mutter": 1,
-    "Zhong": 2,
-    "Ko" : 3,
-    "M3GNet": 4,
-    "CHGNet": 5,
-    "MACE": 6,
-}
-
-SUBPLOT_ORDER = {
-    "a": "Mutter",
-    "b": "Zhong",
-    "c": "Ko",
-    "d": "M3GNet",
-    "e": "CHGNet",
-    "f": "MACE",
-}
-
-
-SPACEGROUP_MAP = {"B2" : 221,
-                  "B19P" : 11,
-                  "B19" : 51,
-                  "BCO" : 63,
-                  }
 
 def get_phonon_config(structure_name, potential):
     """Settings for the phonon calculations
@@ -38,6 +14,7 @@ def get_phonon_config(structure_name, potential):
             "Mutter": ((8, 8, 8), 0.0935, 200),
             "Zhong": ((8, 8, 8), 0.0935, 200),
             "Ko": ((8, 8, 8), 0.01, 200),
+            "Kavousi" : ((8, 8, 8), 0.01, 200),
             "Kim": ((8, 8, 8), 0.01, 200),
             "M3GNet": ((8, 8, 8), 0.260, 200),
             "CHGNet": ((8, 8, 8), 0.030, 200),
@@ -49,6 +26,7 @@ def get_phonon_config(structure_name, potential):
             "Mutter": ((7, 9, 7), 0.05, 200),
             "Zhong": ((7, 9, 7), 0.05, 200),
             "Ko": ((7, 9, 7), 0.03, 200),
+            "Kavousi" : ((7, 9, 7), 0.03, 200),
             "Kim": ((7, 9, 7), 0.03, 200),
             "M3GNet": ((7, 9, 7), 0.05, 200),
             "CHGNet": ((7, 9, 7), 0.05, 200),
@@ -60,6 +38,7 @@ def get_phonon_config(structure_name, potential):
             "Mutter": ((8, 6, 4), 0.05, 200),
             "Zhong": ((8, 6, 4), 0.05, 200),
             "Ko": ((8, 6, 4), 0.01, 200),
+            "Kavousi" : ((8, 6, 4), 0.01, 200),
             "Kim": ((8, 6, 4), 0.01, 200),
             "M3GNet": ((8, 6, 4), 0.05, 200),
             "CHGNet": ((8, 6, 4), 0.05, 200),
[…]

### 2024-01-09 src/scripts/PlotPhonons.py

@@ -7,6 +7,7 @@ import argparse
 
 from ase.db import connect
 
+from PlotConfigs import SUBPLOT_ORDER
 
 def format_xticklabels(labels, locs):
     # Dictionary to keep track of formatted labels with their locations
@@ -97,7 +98,7 @@ def plot_default_phonons(bs, dos, info, ymaxlim=9.0, of="."):
     )
     fig.clear()
     plt.close(fig)
-    
+
     return None
 
 
@@ -169,7 +170,7 @@ def plot_individual_phonon_bandstructure(dbname, chemsys, model_name, structure_
                 )
                 plt.savefig(paths.data / fname)
                 plt.close(fig)
-                
+
     return None
 
 
@@ -192,7 +193,9 @@ def plot_all_strains_phonons(
     fig, ax = plt.subplots()
 
     with connect(dbname) as db:
-        first_row = db.get(structure_name=structure_name, model_name=model_name)
+        first_row = db.get(
+            chemsys=chemsys, structure_name=structure_name, model_name=model_name
+        )
         if chemsys != "".join(sorted(set(first_row.symbols))):
             return None
 
@@ -238,18 +241,18 @@ def plot_all_strains_phonons(
         )
 
         # Calculate midpoints for q_vec_labels
-        midpoints = 0.5 * (norm_locs[f_strain_key][:-1] + norm_locs[f_strain_key][1:])
-        q_vec_labels = f_pdata["bandstructure"]["q_vec_labels"]
-        for midpoint, vec_label in zip(midpoints, q_vec_labels):
-            ax.text(
-                midpoint,
-                ax.get_ylim()[1],
-                vec_label,
-                ha="center",
-                va="bottom",
-                transform=ax.transData,
-                size=6,
-            )
+        #midpoints = 0.5 * (norm_locs[f_strain_key][:-1] + norm_locs[f_strain_key][1:])
+        #q_vec_labels = f_pdata["bandstructure"]["q_vec_labels"]
+        #for midpoint, vec_label in zip(midpoints, q_vec_labels):
+        #    ax.text(
+        #        midpoint,
+        #        ax.get_ylim()[1],
+        #        vec_label,
+        #        ha="center",
+        #        va="bottom",
+        #        transform=ax.transData,
+        #        size=6,
+        #    )
 
         ax.set_xlim((0, norm_locs[f_strain_key][-1]))
 
[…]

### 2024-01-09 src/scripts/Structures.py

@@ -112,7 +112,7 @@ def NiTi_B19_Structure(a=4.09, b=2.86, c=4.59, alpha=90.0, beta=90.0, gamma=90.0
     return B19
 
 
-def PbTi_B19_Structure(a=4.61, b=2.76, c=4.86, alpha=90.0, beta=90.0, gamma=90.0):
+def PtTi_B19_Structure(a=4.61, b=2.76, c=4.86, alpha=90.0, beta=90.0, gamma=90.0):
     """
     Pmma (spacegroup 51)
     Ref. Kadkhodaei, S.; van de Walle, Acta Materialia 2018, 147, 296–303.

### 2024-01-09 src/scripts/TableConfig.py

@@ -1,0 +1,22 @@
+SPACEGROUP_MAP = {"B2" : 221,
+                  "B19P" : 11,
+                  "B19" : 51,
+                  "BCO" : 63,
+                  }
+
+ORDER = {
+    "NiTi": {
+        "Mutter": 1,
+        "Zhong": 2,
+        "Ko": 3,
+        "M3GNet": 4,
+        "CHGNet": 5,
+        "MACE": 6,
+    },
+    "PtTi": {
+        "Kim": 1,
+        "M3GNet": 2,
+        "CHGNet": 3,
+        "MACE": 4,
+    },
+}

### 2024-01-12 src/tex/ms.tex
Updates.

@@
 \usepackage{longtable}
+\usepackage{tabularx}
 \usepackage{subcaption}
@@
         \caption{
-          Initial unit cells of used in this work for NiTi, (a) \textit{B2}(221), (b) \textit{B19}(11), (c) \textit{B19'}(51), and (d) \textit{BCO}(61).
-The spacegroup numbers are given in the parentheses.
-Also additional atoms (i.e.
-non-basis sites) are shown based on periodic boundary conditions.
-For PtTi (a-b) structures are used with different unit cell parameters and basis coordinates.
+          Initial \ce{NiTi} structures (i.e., unrelaxed) used for (a) \textit{B2}(221), (b) \textit{B19}(11), (c) \textit{B19'}(51), and (d) \textit{BCO}(61) with spacegroup numbers are given in the parentheses.
+The green atoms are \ce{Ni} and \ce{Ti} as gray and unit cells are highlighted with parametes shown on axis.
+For all unit cells the $\alpha$ and $\gamma$ angles are 90.0$^\circ$.
+For PtTi, (a-b) structures are used with different unit cell parameters and basis coordinates and then optimized accordingly.
         }
@@
 To ensure that the unit cell parameters were adjusted along with the atomic coordinates and that the crystal symmetry is not broken, we utlize a constraint and filter during the optimization process.
+\par
+
+In fig.~\ref{fig:opt_density} the density of the structures for different potentials is shown.
+We observe that in general the densities and lattice parameters are within agreement with previously reported values.
 The optimized lattice parameters as well as basis atoms for all structures and models are provided in appendix \ref{sec:opt_structures}.
 \par
+
+\begin{figure}[ht!]
+  \label{fig:opt_density}
+  \centering
+  \caption{}
+\end{figure}
 
@@
 \appendix
-
+\section{Appendix}
 \subsection{Workflow Reproducibility}

### 2024-01-12 src/scripts/CalculateElastic.py

@@ -47,7 +47,7 @@ def calculate_elastic_constants(
     structure_name,
     potential,
     npoints=10,
-    displacement=0.5,
+    displacement=0.25,
     update=True,
 ):
     """

### 2024-01-12 src/scripts/CalculationRunner.py

@@ -45,6 +45,7 @@ def min_eos_calc(dbname, structure, potential, strain_info):
             """
         )
         entry_existed = True
+
     except:
         print(
             f"""
@@ -63,6 +64,7 @@ def min_eos_calc(dbname, structure, potential, strain_info):
             ------------------------------------------------------------
             """
         )
+
     return entry_existed
 
 
@@ -232,7 +234,6 @@ def main():
         )
     elif args.calc_type == "elastic":
         elastic_calc(db_path_file, structure, (args.model, asecalc))
-
     return None
 
 

### 2024-01-12 src/scripts/Config.py

@@ -16,7 +16,7 @@ def get_phonon_config(structure_name, potential):
             "Ko": ((8, 8, 8), 0.01, 200),
             "Kavousi" : ((8, 8, 8), 0.01, 200),
             "Kim": ((8, 8, 8), 0.01, 200),
-            "M3GNet": ((8, 8, 8), 0.260, 200),
+            "M3GNet": ((8, 8, 8), 0.010, 200),
             "CHGNet": ((8, 8, 8), 0.030, 200),
             "MACE": ((8, 8, 8), 0.010, 200),
             "ALIGNN": ((6, 6, 6), 0.050, 200),

### 2024-01-12 src/scripts/GenerateElasticTable.py

@@ -2,13 +2,21 @@ import sys
 from ase.db import connect
 import paths
 
+from TableConfig import SPACEGROUP_MAP, ORDER
+
 
 def generate_elastic_constants_table(dbname, chemsys, output_filename):
     db = connect(dbname)
+    
+    sorder = ORDER[chemsys]["structure"]
+    default_order = max(sorder.values())
+    unique_structure_types = set(entry.structure_name for entry in db.select())
+    structure_order = sorted(unique_structure_types, key=lambda x: sorder.get(x, default_order))
+
 
     with open(output_filename, "w") as file:
         # Top-level table with two main columns
-        file.write("\\begin{longtable}{|l|c|}\n")
+        file.write("\\begin{longtable}{|l|l|}\n")
         file.write(
             "\\caption{Elastic constants for "
             + chemsys
@@ -17,7 +25,7 @@ def generate_elastic_constants_table(dbname, chemsys, output_filename):
             + "}\\\\\n"
         )
         file.write("\\hline\n")
-        file.write("Structure & Properties \\\\\n")
+        file.write("Structure (\#) & Properties \\\\\n")
         file.write("\\hline\n")
         file.write("\\endfirsthead\n")
         file.write(
@@ -29,26 +37,30 @@ def generate_elastic_constants_table(dbname, chemsys, output_filename):
         file.write("\\hline\n")
         file.write("\\endfoot\n")
 
-        # Group entries by structure type
-        grouped_entries = {}
-        for entry in db.select(chemsys=chemsys):
-            structure_name = entry.structure_name.replace("_", "-")
-            if structure_name not in grouped_entries:
-                grouped_entries[structure_name] = []
-            grouped_entries[structure_name].append(entry)
+        for structure_type in structure_order:
+            structure_name = structure_type.replace("_", "-")
+            spg_num = SPACEGROUP_MAP[structure_name]
+            file.write(f"{structure_name} ({spg_num}) & ")
 
-        # Process each group
-        for structure_name, entries in grouped_entries.items():
-            file.write(f"{structure_name} & ")
+            entries = db.select(chemsys=chemsys, structure_name=structure_type)
 
-            # Subtable for model and elastic constants
-            file.write("\\begin{tabular}{c|c|c|c|c|c|c|c|c|c|c|c|c|c}\n")
+            # Filter and sort based on models defined in ORDER
+            sentries = sorted(
+                [
+                    entry
[…]

### 2024-01-12 src/scripts/GenerateEquilibriumTable.py

@@ -67,7 +67,12 @@ def get_element_basis_list(structure, spacegroup, tol=1e-3):
 def generate_structure_table(dbname, chemsys, output_filename):
     db = connect(dbname)
 
+    sorder = ORDER[chemsys]["structure"]
+    default_order = max(sorder.values())
     unique_structure_types = set(entry.structure_name for entry in db.select())
+    structure_order = sorted(unique_structure_types, key=lambda x: sorder.get(x, default_order))
+
+    
 
     with open(output_filename, "w") as file:
         file.write("\\begin{longtable}{|l|c|}\n")
@@ -85,7 +90,7 @@ def generate_structure_table(dbname, chemsys, output_filename):
         file.write("\\hline\n")
         file.write("\\endfoot\n")
 
-        for structure_type in unique_structure_types:
+        for structure_type in structure_order:
             structure_name = structure_type.replace("_", "-")
             spg_num = SPACEGROUP_MAP[structure_name]
             file.write(f"{structure_name}\n({spg_num}) & ")
@@ -98,8 +103,8 @@ def generate_structure_table(dbname, chemsys, output_filename):
 
             # Filter and sort based on models defined in ORDER
             sentries = sorted(
-                [entry for entry in entries if entry.model_name in ORDER[chemsys]],
-                key=lambda entry: ORDER[chemsys][entry.model_name]
+                [entry for entry in entries if entry.model_name in ORDER[chemsys]["model"]],
+                key=lambda entry: ORDER[chemsys]["model"][entry.model_name]
             )
 
             for entry in sentries:

### 2024-01-12 src/scripts/GruneisenParameters.py

@@ -226,8 +226,8 @@ def generate_mode_gruneisen_table(
 
             # Filter and sort based on models defined in ORDER
             sentries = sorted(
-                [entry for entry in entries if entry.model_name in ORDER[chemsys]],
-                key=lambda entry: ORDER[chemsys][entry.model_name]
+                [entry for entry in entries if entry.model_name in ORDER[chemsys]['model']],
+                key=lambda entry: ORDER[chemsys]['model'][entry.model_name]
             )
 
             for entry in sentries:

### 2024-01-12 src/scripts/PlotConfigs.py

@@ -7,14 +7,12 @@ ORDER = {
         "M3GNet": 4,
         "CHGNet": 5,
         "MACE": 6,
-        "ALIGNN":7
     },
     "PtTi": {
         "Kim": 1,
         "M3GNet": 2,
         "CHGNet": 3,
         "MACE": 4,
-        "ALIGNN":5
     },
 }
 
@@ -26,14 +24,12 @@ SUBPLOT_ORDER = {
         "d": "M3GNet",
         "e": "CHGNet",
         "f": "MACE",
-        "g": "ALIGNN",
     },
     "PtTi": {
         "a": "Kim",
         "b": "M3GNet",
         "c": "CHGNet",
         "d": "MACE",
-        "e": "ALIGNN",
     },
 }
 

### 2024-01-12 src/scripts/Structures.py

@@ -120,7 +120,7 @@ def PtTi_B19_Structure(a=4.61, b=2.76, c=4.86, alpha=90.0, beta=90.0, gamma=90.0
     Suppl. Mater. Appendix A
     """
     B19 = crystal(
-        ["Pb", "Ti"],
+        ["Pt", "Ti"],
         [(0.25, 0.0, 0.6874), (0.25, 0.50, 0.1958)],
         cellpar=[a, b, c, alpha, beta, gamma],
         pbc=True,
@@ -128,7 +128,7 @@ def PtTi_B19_Structure(a=4.61, b=2.76, c=4.86, alpha=90.0, beta=90.0, gamma=90.0
     )
 
     B19.info["structure_name"] = "B19"
-    B19.info["chemsys"] = "PbTi"
+    B19.info["chemsys"] = "PtTi"
     return B19
 
 
@@ -259,7 +259,7 @@ def get_structure(chemsys, structure_name):
     elif chemsys == "PtTi":
         structure_functions = {
             "B2": PtTi_B2_Structure,
-            "B19": PbTi_B19_Structure,
+            "B19": PtTi_B19_Structure,
         }
 
     if structure_name not in structure_functions:
@@ -362,7 +362,7 @@ def spacegroup51_bandpath(
     npoints=200,
 ):
     """
-    Need to verify special points  because c-axis is taken as largest in NiTi/PbTi B19.
+    Need to verify special points  because c-axis is taken as largest in NiTi/PtTi B19.
     """
 
     cell = structure.cell

### 2024-01-18 src/tex/ms.tex
Updates to plot scripts

@@
 %% Packages
-\usepackage{orcidlink}
-\usepackage[utf8]{inputenc}
-\usepackage{fontspec}
-\usepackage{fontawesome5}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{siunitx}
-\usepackage[version=4]{mhchem}
-\usepackage{pdflscape}
+\usepackage{orcidlink} \usepackage[utf8]{inputenc}
+\usepackage{fontspec} \usepackage{fontawesome5} \usepackage{amssymb}
+\usepackage{amsmath} \usepackage{siunitx}
+\usepackage[version=4]{mhchem} \usepackage{pdflscape}
 %% Handles multipage table. Not sure allowed by pub.
-\usepackage{longtable}
-\usepackage{tabularx}
-\usepackage{subcaption}
-\usepackage{listings}
-\lstset{
-  basicstyle=\ttfamily,
-  columns=fullflexible,
-  frame=single,
-  breaklines=true,
-  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},
-}
+\usepackage{longtable} \usepackage{tabularx} \usepackage{subcaption}
+\usepackage{listings} \lstset{ basicstyle=\ttfamily,
+  columns=fullflexible, frame=single, breaklines=true,
+  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}, }
 
@@
 
-\title{Qualitative Assessesment of Graph Neural Network Potentials for  Phonon Properties of Shape Memory Alloys using Graph Neural Network Potentials}
+\title{Qualitative Assessesment of Graph Neural Network Potentials for
+  Phonon Properties of Shape Memory Alloys using Graph Neural Network
+  Potentials}
 
 % Author list
-\author[1]{Stefan Bringuier\corref{cor1}\orcidlink{0000-0001-6753-1437}}
-\ead{stefanbringuier@gmail.com}
-\cortext[cor1]{Corresponding author}
-\affiliation[1]{organization={Indpendent Researcher},
-  city={San Diego, CA},
-  country={United States},
-}
+\author[1]{Stefan
+  Bringuier\corref{cor1}\orcidlink{0000-0001-6753-1437}}
+\ead{stefanbringuier@gmail.com} \cortext[cor1]{Corresponding author}
+\affiliation[1]{organization={Indpendent Researcher}, city={San Diego,
+    CA}, country={United States}, }
 
@@
 \begin{abstract}
-Shape Memory Alloys (SMAs) such as NiTi are pivotal in vibrational dampening,robotic arms, prothestic hands, and mechanical vales as a result of their unique displacive phase transformation properties.
-However, assessing their phonon characteristics, which provide insight into the displacive phase transformation, by using conventional Density Functional Theory (DFT) methods is computationally challenging.
-In this paper we assess the use of recent Graph Neural Network (GNN) potentials as a viable alternative.
-These GNN potentials have leveraged large DFT-computed databases for training and provide a good approximation to the potential energy surfaces which enable inference on energies and forces in lattice and atomic dynamic calculations.
-We compare GNN potentials with the Embedded Atom Method (EAM) and Modified EAM style potentials in predicting phonon properties of NiTi and PtTi.
-We find that in general the EAM/MEAM potentials are slightly more performant than the GNN in that they appear to more accurately describe equation of state, phonon dispersions, and elastic constants.
-However, the MACE potential shows significant promise as an exploritory potential for designing and characterizing more complex SMAs.
-Our findings suggest GNN potentials as a promising initial avenue for advancing SMA research, in particular ternary and quaternary systems, upon fine-tuning of pre-trained GNN models with more relevant DFT data (i.e., force-constants) to permit better outcomes.
+  Shape Memory Alloys (SMAs) such as NiTi are pivotal in vibrational
+  dampening,robotic arms, prothestic hands, and mechanical vales as a
+  result of their unique displacive phase transformation
+  properties. However, assessing their phonon characteristics, which
+  provide insight into the displacive phase transformation, by using
+  conventional Density Functional Theory (DFT) methods is
+  computationally challenging. In this paper we assess the use of
+  recent Graph Neural Network (GNN) potentials as a viable
+  alternative. These GNN potentials have leveraged large DFT-computed
+  databases for training and provide a good approximation to the
+  potential energy surfaces which enable inference on energies and
+  forces in lattice and atomic dynamic calculations. We compare GNN
+  potentials with the Embedded Atom Method (EAM) and Modified EAM
+  style potentials in predicting phonon properties of NiTi and
+  PtTi. We find that in general the EAM/MEAM potentials are slightly
+  more performant than the GNN in that they appear to more accurately
+  describe equation of state, phonon dispersions, and elastic
+  constants. However, the MACE potential shows significant promise as
+  an exploritory potential for designing and characterizing more
+  complex SMAs. Our findings suggest GNN potentials as a promising
+  initial avenue for advancing SMA research, in particular ternary and
+  quaternary systems, upon fine-tuning of pre-trained GNN models with
+  more relevant DFT data (i.e., force-constants) to permit better
+  outcomes.
 \end{abstract}
@@
 
-The design and development of shape memory alloys (SMA) via \textit{in-silico} methods would provide great value in advancing structural, medical, and other technologies.
-In order to enable such activity would require sampling from very large structural and compositional spaces and evaluating material properties indicative of displacive phase transformation phenomena.
-The use of density functional theory calculations to obtain properties such as equations of state, phonon disperisons, and elastic constants, can be done, however, such calculations are computationally demanding and are potentially prohibitive for smaller groups trying to develop new SMAs.
-To alleviate such restriction, one can resort to classical interatomic potentials which can greatly improve computational cost in  evaluating the energies and forces more.
-However, the immediate challenge is to whether a suitable interatomic potential exist given a chemical system, and moreover, how transferable is it to different compositions and phases.
-\par
+The design and development of shape memory alloys (SMA) via
+\textit{in-silico} methods would provide great value in advancing
+structural, medical, and other technologies. In order to enable such
+activity would require sampling from very large structural and
+compositional spaces and evaluating material properties indicative of
+displacive phase transformation phenomena. The use of density
+functional theory calculations to obtain properties such as equations
+of state, phonon disperisons, and elastic constants, can be done,
+however, such calculations are computationally demanding and are
+potentially prohibitive for smaller groups trying to develop new
+SMAs. To alleviate such restriction, one can resort to classical
+interatomic potentials which can greatly improve computational cost in
+evaluating the energies and forces more. However, the immediate
+challenge is to whether a suitable interatomic potential exist given a
+chemical system, and moreover, how transferable is it to different
+compositions and phases. \par
 
-Recently, there as been a growing effort in the development of machine learning potentials.
-These deviate from traditional physics approaches to interatomic potentials, which have historically been inspired from more accurate methods like tight-binding method.
-In machine learning potentials, one describes the atoms and structure using descriptors of features, these are typically learned or prescribed.
-The features can be prescribed in isolation or combination of  structural, compositional, and property characteristics of the material data point.
-Futhermore, one may choose to fined new representations via unsurpervised learning.
-\par
+Recently, there as been a growing effort in the development of machine
+learning potentials. These deviate from traditional physics approaches
+to interatomic potentials, which have historically been inspired from
+more accurate methods like tight-binding method. In machine learning
+potentials, one describes the atoms and structure using descriptors of
+features, these are typically learned or prescribed. The features can
+be prescribed in isolation or combination of structural,
+compositional, and property characteristics of the material data
+point. Futhermore, one may choose to fined new representations via
+unsurpervised learning. \par
 
-The subsequent step once the featureization process is completed is to determine a suitable neural network (NN) architecture which can learn the target property or properties given the feature inputs.
-In the case of atomistic simulations the target is prediction of the potential energy surface and corresponding gradients (i.e., forces).
-Two of the most early attempts to do this were the works of \citet{Bartok2010} and \citet{Thompson2015}, whereby projection onto higher-dimensional basis, such as bispectrum components of local neighbor density, is learned in order to describe the atomic environments.
-The energy per atom is then determined by these learned features by employing linear regression or non-linear Gaussian process regression.
-We note that these are just two of the early examples but similar potentials have been developed in earlier, in parallel, or recently by various groups \cite{Behler2007,Wang2018,Deringer2019,Lee2019}.
-\par
+The subsequent step once the featureization process is completed is to
+determine a suitable neural network (NN) architecture which can learn
+the target property or properties given the feature inputs. In the
+case of atomistic simulations the target is prediction of the
+potential energy surface and corresponding gradients (i.e.,
+forces). Two of the most early attempts to do this were the works of
+\citet{Bartok2010} and \citet{Thompson2015}, whereby projection onto
+higher-dimensional basis, such as bispectrum components of local
+neighbor density, is learned in order to describe the atomic
+environments. The energy per atom is then determined by these learned
+features by employing linear regression or non-linear Gaussian process
+regression. We note that these are just two of the early examples but
+similar potentials have been developed in earlier, in parallel, or
+recently by various groups
+\cite{Behler2007,Wang2018,Deringer2019,Lee2019}. \par
 
-While these early attempts using ML models were successfully, they may only include limited representations of structural and/or chemical details to learn new features from which could limit transferability.
-Futhermore, the training process was typically restricted to a limited number of elemental constieunts of the target material system.
-As a consequence,  for each chemical system and phases of interest one needs to provide a suitable training dataset.
-To move beyond such limitations, groups began using graph neural network (GNN) architectures and message passing scheme to represent the atomic structures as nodes and edges \cite{Reiser2022}.
-This enabled two distinctions, first being that the nodes and edges corresponding to the atoms and bonds could include features, and secondly the learning could be done over a large structural and compositional space enabling a generalization\footnote{Typically these are referred to as Universal interatomic potentials or forcefields since they in theory generalize well to any chemical system or phase.}.
-The use of GNNs is efficient in representing structure-composition-property inputs to additional NN layers for predicting potential energy surfaces and gradients and was initially pursued in the CGCNN, MEGNet, and ALIGNN potentials \cite{Choudhary2021,Chen2019,Xie2018}.
-The differences between the approaches are predominately in atomic and bond features, nearest-neighbor graph, and specific NN archetcture.
-\par
+While these early attempts using ML models were successfully, they may
+only include limited representations of structural and/or chemical
+details to learn new features from which could limit
+transferability. Futhermore, the training process was typically
+restricted to a limited number of elemental constieunts of the target
+material system. As a consequence, for each chemical system and phases
+of interest one needs to provide a suitable training dataset. To move
+beyond such limitations, groups began using graph neural network (GNN)
+architectures and message passing scheme to represent the atomic
+structures as nodes and edges \cite{Reiser2022}. This enabled two
+distinctions, first being that the nodes and edges corresponding to
+the atoms and bonds could include features, and secondly the learning
+could be done over a large structural and compositional space enabling
+a generalization\footnote{Typically these are referred to as Universal
+  interatomic potentials or forcefields since they in theory
+  generalize well to any chemical system or phase.}. The use of GNNs
+is efficient in representing structure-composition-property inputs to
+additional NN layers for predicting potential energy surfaces and
+gradients and was initially pursued in the CGCNN, MEGNet, and ALIGNN
+potentials \cite{Choudhary2021,Chen2019,Xie2018}. The differences
+between the approaches are predominately in atomic and bond features,
+nearest-neighbor graph, and specific NN archetcture. \par
 
-The training of these early GNN utlized large density functional theory calculation repositories and mostly tuned the learning cost function to ground-state structure and energy.
-This resulted in mean-absolute-errors and other measures of performance that could be improved upon \cite{Riebesell2023}.
-More recently there have been significant improvements in both training data used and the GNN architecture that show promise for truly approaching a general interatomic potential suitable for static and dynamic atomistic calculations of multicomponents and multiphases.
-If succesfull, identifying and designing new potential SMA materials using GNN interatomic potentials could have tremendous impact on technology.
-Therefore, the primary goal is to provide a qualitative assessment of pre-trained GNN interatomic potentials that are promoted as being transferable across compositions and phases.
-We specfically focus on the predictions for equations of state, phonon bandstructure, and elastic constants for various phases of NiTi and PtTi using three GNN potentials: M3GNet, CHGNet, and MACE \cite{Chen2022,Deng2023,Batatia2022}.
-We note that the \citet{Riebesell2023} provides a summarized performance over these GNN using the materials project database, but does not specifically address SMA and phonon characteristics.
-\par
+The training of these early GNN utlized large density functional
+theory calculation repositories and mostly tuned the learning cost
+function to ground-state structure and energy. This resulted in
+mean-absolute-errors and other measures of performance that could be
+improved upon \cite{Riebesell2023}. More recently there have been
+significant improvements in both training data used and the GNN
+architecture that show promise for truly approaching a general
+interatomic potential suitable for static and dynamic atomistic
+calculations of multicomponents and multiphases. If succesfull,
+identifying and designing new potential SMA materials using GNN
+interatomic potentials could have tremendous impact on
+technology. Therefore, the primary goal is to provide a qualitative
+assessment of pre-trained GNN interatomic potentials that are promoted
+as being transferable across compositions and phases. We specfically
+focus on the predictions for equations of state, phonon bandstructure,
+and elastic constants for various phases of NiTi and PtTi using three
+GNN potentials: M3GNet, CHGNet, and MACE
+\cite{Chen2022,Deng2023,Batatia2022}. We note that the
+\citet{Riebesell2023} provides a summarized performance over these GNN
+using the materials project database, but does not specifically
+address SMA and phonon characteristics. \par
 
@@
 \label{sec:workflow}
-To promote reproducible research and align with findable, accessible, interoper-able and reusable (FAIR) principles\cite{Walsh2024}, we utilize \showyourwork python workflow\cite{Luger2021} and several open source python packages.
-Instructions for reproducing calculated data, figures, tables, and manuscript are provided in appendix~\ref{sec:appx_reproduce} and corresponding github repo\footnote{\url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP}}.\par
+To promote reproducible research and align with findable, accessible,
+interoper-able and reusable (FAIR) principles\cite{Walsh2024}, we
+utilize \showyourwork python workflow\cite{Luger2021} and several open
+source python packages. Instructions for reproducing calculated data,
+figures, tables, and manuscript are provided in
+appendix~\ref{sec:appx_reproduce} and corresponding github
+repo\footnote{\url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP}}.\par
 
@@
 \label{sec:calc_details}
-The ground-state structures for the NiTi phases \textit{B2}, \textit{B19}, \textit{B19'}, and \textit{BCO} are generated using the atomic simulation environment (ASE) package \cite{Larsen2017}.
-This requires that the unit cell parameters, basis atoms, and spacegroup number be provided.
-The initial unit cell parameters and basis corrdinates for these structures have been taken from the materials project \cite{Jain2013} and several references \cite{Haskins2016,Kadkhodaei2018}.
-The initial structures for NiTi are shown in fig~\ref{fig:niti_structures} and are adapted for PtTi as well given same spacegroup symmetries.
-The energy and forces on the atoms are determined using the ASE calculator interfaces for EAM, MEAM, M3GNet, CHGNet, and MACE.\cite{Mutter2010,Zhong2011,Ko2015,Kim2017,Chen2022,Deng2023,Batatia2022} The EAM and MEAM potential interfaces are achieved through the use of LAMMPS python library \cite{Thompson2022} and ASE.
-The ALIGNN ASE calculator interface was also implemented, however, we found that the predictions for NiTi and PtTi to be unreliable are not included in this study.\par
+The ground-state structures for the NiTi phases \textit{B2},
+\textit{B19}, \textit{B19'}, and \textit{BCO} are generated using the
+atomic simulation environment (ASE) package \cite{Larsen2017}. This
+requires that the unit cell parameters, basis atoms, and spacegroup
+number be provided. The initial unit cell parameters and basis
+corrdinates for these structures have been taken from the materials
+project \cite{Jain2013} and several references
+\cite{Haskins2016,Kadkhodaei2018}. The initial structures for NiTi are
+shown in fig~\ref{fig:niti_structures} and are adapted for PtTi as
+well given same spacegroup symmetries. The energy and forces on the
+atoms are determined using the ASE calculator interfaces for EAM,
+MEAM, M3GNet, CHGNet, and
+MACE.\cite{Mutter2010,Zhong2011,Ko2015,Kim2017,Chen2022,Deng2023,Batatia2022}
+The EAM and MEAM potential interfaces are achieved through the use of
+LAMMPS python library \cite{Thompson2022} and ASE. The ALIGNN ASE
+calculator interface was also implemented, however, we found that the
+predictions for NiTi and PtTi to be unreliable are not included in
+this study.\par
 
-For each combination of potential/model, chemical system, and phase we perform calcualtions for the structure minimization, equation of state (EOS), phonons, and elastic constants.
-To determine the finite-temperature phase transitions we utlize an isothermal-isobaric ensemble via molecular dynamics simulations.
-The thermostating and barostating is done with Nose-Hoover and Parrinello-Rahman dynamics and the tempertaure is adjusted in a quasi-static manner which entails equlibrating the simulation for 10 ns while decrementing the temperature by 10 Kelvin; the pressure is kept constant at 1 bar.
-We note that in order to achieve the different computing environments required by the GNN potentials we leverage the ability to specify the python computing environments using \href{https://github.com/showyourwork/showyourwork}{\showyourwork} package.\par
+For each combination of potential/model, chemical system, and phase we
+perform calcualtions for the structure minimization, equation of state
+(EOS), phonons, and elastic constants. To determine the
+finite-temperature phase transitions we utlize an isothermal-isobaric
+ensemble via molecular dynamics simulations. The thermostating and
+barostating is done with Nose-Hoover and Parrinello-Rahman dynamics
+and the tempertaure is adjusted in a quasi-static manner which entails
+equlibrating the simulation for 10 ns while decrementing the
+temperature by 10 Kelvin; the pressure is kept constant at 1 bar. We
+note that in order to achieve the different computing environments
+required by the GNN potentials we leverage the ability to specify the
+python computing environments using
+\href{https://github.com/showyourwork/showyourwork}{\showyourwork}
+package.\par
 
@@
     \includegraphics[width=\linewidth]{figures/NiTi_VisualizedStructures.png}
-        \caption{
-          Initial \ce{NiTi} structures (i.e., unrelaxed) used for (a) \textit{B2}(221), (b) \textit{B19}(11), (c) \textit{B19'}(51), and (d) \textit{BCO}(61) with spacegroup numbers are given in the parentheses.
-The green atoms are \ce{Ni} and \ce{Ti} as gray and unit cells are highlighted with parametes shown on axis.
-For all unit cells the $\alpha$ and $\gamma$ angles are 90.0$^\circ$.
-For PtTi, (a-b) structures are used with different unit cell parameters and basis coordinates and then optimized accordingly.
-        }
+    \caption{ Initial \ce{NiTi} structures (i.e., unrelaxed) used for
+      (a) \textit{B2}(221), (b) \textit{B19}(11), (c)
+      \textit{B19'}(51), and (d) \textit{BCO}(61) with spacegroup
+      numbers are given in the parentheses. The green atoms are
+      \ce{Ni} and \ce{Ti} as gray and unit cells are highlighted with
+      parametes shown on axis. For all unit cells the $\alpha$ and
+      $\gamma$ angles are 90.0$^\circ$. For PtTi, (a-b) structures are
+      used with different unit cell parameters and basis coordinates
+      and then optimized accordingly.  }
     \label{fig:niti_structures}
@@
 
-Prior to any property calculations or dynamics the structures are optimized using the FIRE\cite{Bitzek2006} local optimizer with the forces converged to $2\cdot 10^{-3}$eV/\AA.
-We observed that other local optimization routines find the same ground-state structures.
-To ensure that the unit cell parameters were adjusted along with the atomic coordinates and that the crystal symmetry is not broken, we utlize a constraint and filter during the optimization process.
-\par
+Prior to any property calculations or dynamics the structures are
+optimized using the FIRE\cite{Bitzek2006} local optimizer with the
+forces converged to $2\cdot 10^{-3}$eV/\AA. We observed that other
+local optimization routines find the same ground-state structures. To
+ensure that the unit cell parameters were adjusted along with the
+atomic coordinates and that the crystal symmetry is not broken, we
+utlize a constraint and filter during the optimization process. \par
 
-In fig.~\ref{fig:opt_density} the density of the structures for different potentials is shown.
-We observe that in general the densities and lattice parameters are within agreement with previously reported values.
-The optimized lattice parameters as well as basis atoms for all structures and models are provided in appendix \ref{sec:opt_structures}.
-\par
+In fig.~\ref{fig:opt_density} the density of the structures for
+different potentials is shown. We observe that in general the
+densities and lattice parameters are within agreement with previously
+reported values. The optimized lattice parameters as well as basis
+atoms for all structures and models are provided in appendix
+\ref{sec:opt_structures}. \par
 
@@
 \subsubsection{Equation of State}
-The EOS can be obtained by varying the cell volume and calculating the total system energy without allowing for atomic relaxations.
-Starting with the obtained ground-state structures the unit cell volume is adjusted by applying an isotropic strain,
+The EOS can be obtained by varying the cell volume and calculating the
+total system energy without allowing for atomic relaxations. Starting
+with the obtained ground-state structures the unit cell volume is
+adjusted by applying an isotropic strain,
 
@@
 
-to the cell and scaling the fractional atomic coordinates.
-Here $\epsilon$ is the fraction strain applied to the unit cell.
-To ensure that the strain is indeed isotropic we provide a quality check on the symmetry of the distorted structure using the spacegroup package \textit{spglib} \cite{Togo2018}.
-The isotropic strain is applied in compression and tension, from $-11\%$ to $11\%$, which gave a good range to characterize the EOS curves and obtain analytical fits.
-We fit the EOS using a third order inverse polynomial,
+to the cell and scaling the fractional atomic coordinates. Here
+$\epsilon$ is the fraction strain applied to the unit cell. To ensure
+that the strain is indeed isotropic we provide a quality check on the
+symmetry of the distorted structure using the spacegroup package
+\textit{spglib} \cite{Togo2018}. The isotropic strain is applied in
+compression and tension, from $-11\%$ to $11\%$, which gave a good
+range to characterize the EOS curves and obtain analytical fits. We
+fit the EOS using a third order inverse polynomial,
 
@@
 
-where the coefficients are used to determine equilibrium volume, system energy, and bulk modulus \cite{Alchagirov2003} at 0K.
-In order to compare respective EOS amongst the different phases, we normalize the strained volume of each phase with respect to the obtained minimized ground-state volume.
-\par
+where the coefficients are used to determine equilibrium volume,
+system energy, and bulk modulus \cite{Alchagirov2003} at 0K. In order
+to compare respective EOS amongst the different phases, we normalize
+the strained volume of each phase with respect to the obtained
+minimized ground-state volume. \par
 
@@
 
-Shape memory alloys derive their unique displacive phase transformations from the dynamic instability that occurs during the application of temperature, strain, or stress.
-The origin of this instability occurs due to softening or symmetry breaking of the phonon modes in the phase.
-Therefore, improved understanding of the phonon characteristics of the different phases that can occur for a given SMA is particular important.
-One of the challenges is in performing such calculations, which typically employ density functional theory (DFT) in order to determine the atomic forces, which in turn are used to construct the dynamical matrix.
-However, classical interatomic potentials can alleviate the computational burden.
-\par
+Shape memory alloys derive their unique displacive phase
+transformations from the dynamic instability that occurs during the
+application of temperature, strain, or stress. The origin of this
+instability occurs due to softening or symmetry breaking of the phonon
+modes in the phase. Therefore, improved understanding of the phonon
+characteristics of the different phases that can occur for a given SMA
+is particular important. One of the challenges is in performing such
+calculations, which typically employ density functional theory (DFT)
+in order to determine the atomic forces, which in turn are used to
+construct the dynamical matrix. However, classical interatomic
+potentials can alleviate the computational burden. \par
 
-Phonon calculations are performed using finite-difference method via small displacement (i.e., harmonic approximation) of atoms in a reference cell \cite{Alfe2009}.
-In the small displacement method, atoms in the primitive cell are slightly displaced to calculate forces on other atoms, constructing the force constant matrix, which, when used with the dynamical matrix, allows computing phonon frequencies at any q-vector in the Brillouin zone.
-This requires typically requires the use of a supercell of the original unit cell.
-In order to ensure convergence of this method, one needs to sweep different supercell sizes and atomic displacements.
-This process is done for each interatomic potential and structure.
-The supercells used in this work  have a repeat in any one direction ranging from $5-8$ and depends on the crystal system.
-The atomic displacement ranges from $0.03-0.26$\AA\,and was found to strongly depend on interatomic potential and crystal system.
-Details on the settings are provided in the appendix~\ref{sec:phonon_calc_config}.
-\par
+Phonon calculations are performed using finite-difference method via
+small displacement (i.e., harmonic approximation) of atoms in a
+reference cell \cite{Alfe2009}. In the small displacement method,
+atoms in the primitive cell are slightly displaced to calculate forces
+on other atoms, constructing the force constant matrix, which, when
+used with the dynamical matrix, allows computing phonon frequencies at
+any q-vector in the Brillouin zone. This requires typically requires
+the use of a supercell of the original unit cell. In order to ensure
+convergence of this method, one needs to sweep different supercell
+sizes and atomic displacements. This process is done for each
+interatomic potential and structure. The supercells used in this work
+have a repeat in any one direction ranging from $5-8$ and depends on
+the crystal system. The atomic displacement ranges from
+$0.03-0.26$\AA\,and was found to strongly depend on interatomic
+potential and crystal system. Details on the settings are provided in
+the appendix~\ref{sec:phonon_calc_config}. \par
 
-In order to sample high-symmetry points in the irreducible Brillouin zone (IBZ) for specific structures, we sample from locations specified in the Bilbao crystallographic \textit{KVEC} utility\cite{Aroyo2014}, although the ASE structures object does provide a standardized routine for sampling the irreducible Brillouin zone for the specific structures.\cite{Setyawan2010}.
-We predominately focus on high-symmetry points in the IBZ that correspond to the corners and along directions known to exhibit phonon instabilities.
-The IBZ and points symmetry point locations can be seen in appendix \ref{sec:appx_ibz}.
-The final phonon dispersion diagrams are plot along these points.
-\par
+In order to sample high-symmetry points in the irreducible Brillouin
+zone (IBZ) for specific structures, we sample from locations specified
+in the Bilbao crystallographic \textit{KVEC} utility\cite{Aroyo2014},
+although the ASE structures object does provide a standardized routine
+for sampling the irreducible Brillouin zone for the specific
+structures.\cite{Setyawan2010}. We predominately focus on
+high-symmetry points in the IBZ that correspond to the corners and
+along directions known to exhibit phonon instabilities. The IBZ and
+points symmetry point locations can be seen in appendix
+\ref{sec:appx_ibz}. The final phonon dispersion diagrams are plot
+along these points. \par
 
-To better understand dynamic instability in the various structures, the same isotropic strain in eq.~\ref{eq:isotropic_strain} is used and the repsective phonon bandstructures are calculated.
-These are then plotted to observed the onset or change in stability or instability of phonon different branches.
-\par
+To better understand dynamic instability in the various structures,
+the same isotropic strain in eq.~\ref{eq:isotropic_strain} is used and
+the repsective phonon bandstructures are calculated. These are then
+plotted to observed the onset or change in stability or instability of
+phonon different branches. \par
 
@@
 
-Similar to the phonon calculations the elastic constants within the harmonic approximation the elastic constants can be determined from linear-response of strain-stress calculations.
-Therefore, the calculation requires the use of crystal symmetry along with atomic displacements to determine respective elastic constant componenets.
-We use the \textit{Elastic}\cite{Jochym2018} python package which provides routines for generating the proper atomic displacement matrix given specific crystal spacegroup.
-The elastic constants are determined by solving a set of linear equations using singular value decomposition and given calculated strains and stress.
-The displacements occur with a min and max of 0.5\% and a total of 10 points are used.
-\par
+Similar to the phonon calculations the elastic constants within the
+harmonic approximation the elastic constants can be determined from
+linear-response of strain-stress calculations. Therefore, the
+calculation requires the use of crystal symmetry along with atomic
+displacements to determine respective elastic constant componenets. We
+use the \textit{Elastic}\cite{Jochym2018} python package which
+provides routines for generating the proper atomic displacement matrix
+given specific crystal spacegroup. The elastic constants are
+determined by solving a set of linear equations using singular value
+decomposition and given calculated strains and stress. The
+displacements occur with a min and max of 0.5\% and a total of 10
+points are used. \par
 
@@
 
-The EAM and MEAM potentials were developed to go beyond pair interatomic potentials to treat point and extended defects in metal and alloys systems.
-The motivation was taken from the physical results of  density functional theory and tight-binding methods, namely to add to the pair potentials a energy functional which depends on the local electron density which in turn is affected by the atomic environment.
-These type of interatomic potentials are broadly referred to as EAM or MEAM, with the later included direction (i.e.
-angle) dependent interactions.
-\par
+The EAM and MEAM potentials were developed to go beyond pair
+interatomic potentials to treat point and extended defects in metal
+and alloys systems. The motivation was taken from the physical results
+of density functional theory and tight-binding methods, namely to add
+to the pair potentials a energy functional which depends on the local
+electron density which in turn is affected by the atomic
+environment. These type of interatomic potentials are broadly referred
+to as EAM or MEAM, with the later included direction (i.e. angle)
+dependent interactions. \par
 
-The functional form using during fitting proceed depends on the complexity of the material system but the total energy is typically written as,
+The functional form using during fitting proceed depends on the
+complexity of the material system but the total energy is typically
+written as,
 
@@
 
-where $V_{ij}$ is the pair potential, $F_i$, the functional, and $\rho$ the density function that depends on the environment around $r_i$.
-The MEAM potential extends this by including three-body terms in the $\rho$ to capture features such as covalent bonding.
-The EAM and MEAM potentials have been particular successfully in predicting structure and properties of metals and alloys, including defect formation/energies, phase transitions, grain boundary stability/dynamics.
-One particular advantage of both EAM and MEAM, is their analytical or tabulated forms allow for quick numerical evaluation, with the primary reason being that the energies and forces on a atom are accurately determined within a relatively small interaction volume (i.e., 3-4 \AA interaction cutoff).
-\par
+where $V_{ij}$ is the pair potential, $F_i$, the functional, and
+$\rho$ the density function that depends on the environment around
+$r_i$. The MEAM potential extends this by including three-body terms
+in the $\rho$ to capture features such as covalent bonding. The EAM
+and MEAM potentials have been particular successfully in predicting
+structure and properties of metals and alloys, including defect
+formation/energies, phase transitions, grain boundary
+stability/dynamics. One particular advantage of both EAM and MEAM, is
+their analytical or tabulated forms allow for quick numerical
+evaluation, with the primary reason being that the energies and forces
+on a atom are accurately determined within a relatively small
+interaction volume (i.e., 3-4 \AA interaction cutoff).  \par
 
-For application of EAM to shape memory alloys, and in particular NiTi, \citet{Mutter2010} initially devised an Finnis-Sinclair style EAM potential to capture the structural phase transitions between \textit{B19'} and \textit{B2} phases.
-It was latter shown by \cite{Zhong2011} that some adjustments to the EAM parameters was needed to properly capture the finite-temperature displacement dynamics for the martensitic transition \cite{Gur2017}.
-More recently \citet{Ko2015} developed a second nearest neighbor(2NN) MEAM parameterization of NiTi that demonstrated very good predictive performance and improvements over ealier potentials.
+For application of EAM to shape memory alloys, and in particular NiTi,
+\citet{Mutter2010} initially devised an Finnis-Sinclair style EAM
+potential to capture the structural phase transitions between
+\textit{B19'} and \textit{B2} phases. It was latter shown by
+\cite{Zhong2011} that some adjustments to the EAM parameters was
+needed to properly capture the finite-temperature displacement
+dynamics for the martensitic transition \cite{Gur2017}. More recently
+\citet{Ko2015} developed a second nearest neighbor(2NN) MEAM
+parameterization of NiTi that demonstrated very good predictive
+performance and improvements over ealier potentials.
 
@@
 
-The use of ML potentials for SMA has been demonstrated by \citet{Tang2022}.
-In their work, they utilized the DeePMD-Kit\cite{Wang2018} neural network framework to parameterize a descriptor matrix  from atomic configurations that preserves permutations and  symmetries.
-The descriptors are then which feed into a multilayer perceptron to calculate energies and forces.
-The neural network interatomic potential demonstrated good accuracy of ground-state properties with respect to DFT results and showed agreement with experiments regarding temperature, stress, and defect induced martensitic transformation.
-\paragraphmark
+The use of ML potentials for SMA has been demonstrated by
+\citet{Tang2022}. In their work, they utilized the
+DeePMD-Kit\cite{Wang2018} neural network framework to parameterize a
+descriptor matrix from atomic configurations that preserves
+permutations and symmetries. The descriptors are then which feed into
+a multilayer perceptron to calculate energies and forces. The neural
+network interatomic potential demonstrated good accuracy of
+ground-state properties with respect to DFT results and showed
+agreement with experiments regarding temperature, stress, and defect
+induced martensitic transformation. \paragraphmark
 
-In the work of \citet{Tang2022} ground-state predictions for \textit{B2}, \textit{B19'}, \textit{B19}, and \textit{BCO}\footnote{Referred to as \textit{B33} prototype in \citet{Tang2022}} were provided and showed good agreement.
-However, we note that deviation from Ni-Ti 1:1 stoichiometry or addition of a third species was not investigated nor the inclusion of additional alloying species.
-\par
+In the work of \citet{Tang2022} ground-state predictions for
+\textit{B2}, \textit{B19'}, \textit{B19}, and
+\textit{BCO}\footnote{Referred to as \textit{B33} prototype in
+  \citet{Tang2022}} were provided and showed good agreement. However,
+we note that deviation from Ni-Ti 1:1 stoichiometry or addition of a
+third species was not investigated nor the inclusion of additional
+alloying species. \par
 
-The pre-trained GNN interatomic potentials explored in this work have been parameterized over a large fraction of the interatomic potential ($>$ 80 elements) and therefore are not restricted to specific chemistries or stoichiometries.
-They training data also includes various phases for a given chemical system.
-This permits great flexibility.
-We do not that fine-tuning, the process of adjusting GNN weights and biases using additional training data, can be performed with the GNN studied in this work.
-However, this is out of the scope for this manuscript, given resources and time needed to generate such training data for SMAs.
-\par
+The pre-trained GNN interatomic potentials explored in this work have
+been parameterized over a large fraction of the interatomic potential
+($>$ 80 elements) and therefore are not restricted to specific
+chemistries or stoichiometries. They training data also includes
+various phases for a given chemical system. This permits great
+flexibility. We do not that fine-tuning, the process of adjusting GNN
+weights and biases using additional training data, can be performed
+with the GNN studied in this work. However, this is out of the scope
+for this manuscript, given resources and time needed to generate such
+training data for SMAs. \par
 
-The M3GNet is a GNN that utilizes a 3-body bond-update procedure to more accurately capture bond-order environment changes.
-The M3GNet has been implemented in the MatGL framework and we utilize the \texttt{M3GNet-MP-2021.2.8-PES} pretrained model.
+The M3GNet is a GNN that utilizes a 3-body bond-update procedure to
+more accurately capture bond-order environment changes. The M3GNet has
+been implemented in the MatGL framework and we utilize the
+\texttt{M3GNet-MP-2021.2.8-PES} pretrained model.
 
@@
 
-%The MACE potential pretrained \texttt{2023-12-03-mace-128-L1_epoch-199.model} checkpoint file.
+% The MACE potential pretrained
+% \texttt{2023-12-03-mace-128-L1_epoch-199.model} checkpoint file.
 
@@
 
-One way to understand the phase stability for a given composition is to compare the cohesive energy, which is defined as,
+One way to understand the phase stability for a given composition is
+to compare the cohesive energy, which is defined as,
 
@@
   \label{eq:cohesive_energy}
-  E_{c} = \frac{E_{\text{total}}}{N}
+  E_{c} = \frac{E_{\text{bulk}}-\sum_i^N n_i E^_{\text{atom_i}}}{N}
 \end{equation}
 
-With $E_{\text{total}}$ the total energy of the system and $N$ the number of atoms.
-In fig.~\ref{fig:ecoh} the bar chart for different phases and interatomic potentials models are show.
-We compare these to the predicted cohesive energy for The Materials Project (MP) entry \texttt{MP-571} which corresponds to the \textit{B2} phase \cite{MP--571}.
-In the figure we see that the EAM and MEAM potentials show significantly lower cohesive energy compared to the GNN potentials.
-The stark increase in binding energy for the GNN is surprising as the EAM and MEAM potentials are known to reproduce previous experimental and DFT data \cite{Haskins2016,Ko2015}.
-The GNNs predict a more strongly bound and stable \textit{B2} structure but all are consistent with the entry for The MP and calculation details can be found in ref.~\cite{MP--571}.
-\par
+With $E_{\text{bulk}}$ the total energy of the system,
+$\sum_i^N n_i E_{\text{atom_i}}$ is the sum over isolated atom
+energies for $n_i$ species, and $N$ the number of atoms in bulk
+reference. Classical interatomic potentials are offset with respect to
+the isolated atom energies, so these terms are zero and neglected,
+giving $E_{c} = \frac{E_{\text{bulk}}}{N}$. In fig.~\ref{fig:ecoh} the
+bar chart for different phases and interatomic potentials models are
+show. We compare these to dervided experimental values for
+\textit{B2}\cite{Vandermause2024} and the value from The Materials
+Project (MP) entry \texttt{MP-571} which corresponds to the
+\textit{B2} phase \cite{MP--571}. In the figure we see that the EAM
+and MEAM potentials show significantly lower cohesive energy compared
+to the GNN potentials and are in better agreement with the
+experimental value. Previous studies using the EAM and MEAM potentials
+have been shown to reproduce previous experimental and DFT data
+\cite{Haskins2016,Ko2015} such as cohesive energy. The GNNs predict a
+more strongly bound and stable \textit{B2} structure but all are
+consistent with the entry for The MP and calculation details can be
+found in ref.~\cite{MP--571}. This consistency is resonable
+considering all pretrained GNN models used in this work have been
+developed with MP data. \par
 
@@
     \includegraphics[width=\linewidth]{figures/NiTi_CohesiveEnergyPlot.png}
-        \caption{
-          Comparison of predicted \ce{NiTi} cohesive energy for respective ground-state structure and model.
-As a reference we use The Materials Project entry for the \ce{NiTi} \textit{B2} entry \cite{MP--571}.
-        }
+    \caption{ Comparison of predicted \ce{NiTi} cohesive energy for
+      respective ground-state structure and model. The experimental
+      data point has been extracted from ref.~\cite{Vandermause2024}
+      and the Materials Project entry for the \ce{NiTi} \textit{B2}
+      structure \cite{MP--571} has also been included.  }
     \label{fig:ecoh_niti}
@@
 \end{figure}
+
+For the \ce{PtTi} system we only consider the \textit{B2} and
+\textit{B19'} structures. The MEAM potential shows a lower cohesive
+energy than all three GNN. Interestingly the difference between the
+MEAM and GNN is roughly the same order, $\approx 2.1$ eV, which maybe
+appears to be related to the MP dataset. To our knowledge most the
+pretrained GNN models are using the MP dataset which defaults to DFT
+calculation results that employ the meta-GGA R2SCAN functional. This
+could be the reason for the difference as most EAM and MEAM potentials
+are either parameterized from experimental or DFT data using the PBE
+functional\cite{Mutter2010,Ko2015}. \par
 
@@
     \includegraphics[width=\linewidth]{figures/PtTi_CohesiveEnergyPlot.png}
-        \caption{
-          Comparison of predicted \ce{PtTi} cohesive energy for respective ground-state structure and model.
-        }
+    \caption{ Comparison of predicted \ce{PtTi} cohesive energy for
+      respective ground-state structure and model.  }
     \label{fig:ecoh_ptti}
@@
 
-The EOS curves for the \ce{NiTi} structures \textit{B2}, \textit{B19}, \textit{B19'}, \textit{BCO} for the different potentials studied in this work are shown in fig.~\ref{fig:eos}.
-The two EAM potentials, show that at 0 strain ($V/V_o = 1.0$) both the \textit{B19P} and \textit{B33} phases are stable.
-I believe this is the first reporting of the \textit{B33} phase for these two fittings of the EAM potential for \ce{NiTi}.
-Its important to note that these potentials have been developed to capture the displacive phase transformation between \textit{B19P} to \textit{B2}.
-This is shown in fig.~\ref{fig:eos}(a,b) with the stability indicated upon compressive isotropic strain.
-A mininum appears around $\approx 0.9$.
-At tensile isotropic strains the \textit{B19P} phase is stable up till around $1.18$ where the \textit{Pbcm} phase shows stability.
-An interesting note is that these two fittings of the EAM potential show a double well curve for the \textit{B2} phase.
-\par
+The EOS curves for the \ce{NiTi} structures \textit{B2}, \textit{B19},
+\textit{B19'}, \textit{BCO} for the different potentials studied in
+this work are shown in fig.~\ref{fig:eos}. The two EAM potentials,
+show that at 0 strain ($V/V_o = 1.0$) both the \textit{B19P} and
+\textit{B33} phases are stable. I believe this is the first reporting
+of the \textit{B33} phase for these two fittings of the EAM potential
+for \ce{NiTi}. Its important to note that these potentials have been
+developed to capture the displacive phase transformation between
+\textit{B19P} to \textit{B2}. This is shown in fig.~\ref{fig:eos}(a,b)
+with the stability indicated upon compressive isotropic strain. A
+mininum appears around $\approx 0.9$. At tensile isotropic strains the
+\textit{B19P} phase is stable up till around $1.18$ where the
+\textit{Pbcm} phase shows stability. An interesting note is that these
+two fittings of the EAM potential show a double well curve for the
+\textit{B2} phase. \par
 
-In the UIP-GNN potentials M3GNet, CHGNet, and MACE, the equation of states shown in fig.~\ref{fig:eos}(c-e) look fairly different from thos in fig.~\ref{fig:eos}(a,b).
-For the M3GNet,  at no strain, the \textit{B33} appears to be the predicted stable phase and the \textit{B2} phase becomes stable around $0.9$ (see fig.~\ref{fig:eos}(c).
-The stability for \textit{B19} and \textit{B19P} exit, in trade-off, between approximately $1.025$ and $1.1$ tensile isotropic strain.
-Interestingly beyond a strain of $1.1$ the \textit{Pbcm} is dominate.
-The double-well curve shown for the EAM potentials is not present for the B2-phase.
-The CHGNet potential results shown in fig.~\ref{fig:eos}(d) are similar to the M3GNet, but the stability window of the \textit{B2} occurs at a lower compressive isotropic strain.
-\par
+In the UIP-GNN potentials M3GNet, CHGNet, and MACE, the equation of
+states shown in fig.~\ref{fig:eos}(c-e) look fairly different from
+thos in fig.~\ref{fig:eos}(a,b). For the M3GNet, at no strain, the
+\textit{B33} appears to be the predicted stable phase and the
+\textit{B2} phase becomes stable around $0.9$ (see
+fig.~\ref{fig:eos}(c). The stability for \textit{B19} and
+\textit{B19P} exit, in trade-off, between approximately $1.025$ and
+$1.1$ tensile isotropic strain. Interestingly beyond a strain of $1.1$
+the \textit{Pbcm} is dominate. The double-well curve shown for the EAM
+potentials is not present for the B2-phase. The CHGNet potential
+results shown in fig.~\ref{fig:eos}(d) are similar to the M3GNet, but
+the stability window of the \textit{B2} occurs at a lower compressive
+isotropic strain. \par
 
-The MACE potential shown in fig.~\ref{fig:eos}(e) has a very different set of EOS curves compard to the M3GNet and CHGNet.
-Here many of the phases exhibit a stablity window, except the \textit{B33} phase, which was stable in M3GNet and CHGNet.
-Interestingly, the \text{B19} phase, not the \text{B19P} phase is shown to be the stable zero strain phase.
-One unique perspective of the the UIP-GNN models is that they provide a different displacive phase transformation pathway then that shown by the EAM potentials.
-However, this does not include the meta-stable R-phase or other potential meta-stable phases which can also offer such transformation mechanisms.
-\par
+The MACE potential shown in fig.~\ref{fig:eos}(e) has a very different
+set of EOS curves compard to the M3GNet and CHGNet. Here many of the
+phases exhibit a stablity window, except the \textit{B33} phase, which
+was stable in M3GNet and CHGNet. Interestingly, the \text{B19} phase,
+not the \text{B19P} phase is shown to be the stable zero strain
+phase. One unique perspective of the the UIP-GNN models is that they
+provide a different displacive phase transformation pathway then that
+shown by the EAM potentials. However, this does not include the
+meta-stable R-phase or other potential meta-stable phases which can
+also offer such transformation mechanisms. \par
 
-Finally, the ALIGNN potential EOS was generated, but a few issues arose during the calculations.
-The most prominant was the difficulty in convergence of the structural optimzation to the specified tolerance (see \ref{sec:methods}.
-Secondly the calculated cohesive energies differs from the UIP-GNN by a factor $\approx 2$ and by $1$ eV for the EAM potentials.
-The \textit{B19P} phase is the stable at zero strain and for a considerable portion of the compressive, but all of the tensile.
-At compressive strain beyond $\approx 0.94$ the \textit{B2} phase is stable.
+Finally, the ALIGNN potential EOS was generated, but a few issues
+arose during the calculations. The most prominant was the difficulty
+in convergence of the structural optimzation to the specified
+tolerance (see \ref{sec:methods}. Secondly the calculated cohesive
+energies differs from the UIP-GNN by a factor $\approx 2$ and by $1$
+eV for the EAM potentials. The \textit{B19P} phase is the stable at
+zero strain and for a considerable portion of the compressive, but all
+of the tensile. At compressive strain beyond $\approx 0.94$ the
+\textit{B2} phase is stable.
 
@@
     \includegraphics[width=\linewidth]{figures/NiTi_EquationOfStates.png}
-        \caption{
-            The equation of state curves for different \ce{NiTi} structures for (a) Mutter (EAM) (b) Zhong (EAM) (c) Ko (2NN-MEAM) (d) M3GNet (e) CHGNet (f) MACE interatomic potentials.
-The x-axis is the ratio of the isotropic strained unit cell volume, $V$,  with respect to the ground-state unit cell volume, $V_o$.
-        }
+    \caption{ The equation of state curves for different \ce{NiTi}
+      structures for (a) Mutter (EAM) (b) Zhong (EAM) (c) Ko
+      (2NN-MEAM) (d) M3GNet (e) CHGNet (f) MACE interatomic
+      potentials. The x-axis is the ratio of the isotropic strained
+      unit cell volume, $V$, with respect to the ground-state unit
+      cell volume, $V_o$.  }
     \label{fig:eos}
@@
 \subsection{NiTi: Phonon Instability}
-\label{subsec:niphonons}
-
-\subsubsection{B2 Phase}
-\label{subsubsec:b2}
+\label{subsec:nitiphonons}
 
@@
     \includegraphics[width=\linewidth]{figures/NiTi_B2_ModelsPhononBandstructures.png}
-        \caption{
-          Ground-state, no isotropic strain, Phonon bandstructure of B2 phase for all models.
-The Mutter and Zhong EAM potentials differ in that Mutter predicts no dynamic instability of the B2 phase at 0K.
-        }
+    \caption{ Ground-state, no isotropic strain, Phonon bandstructure
+      of B2 phase for all models. The Mutter and Zhong EAM potentials
+      differ in that Mutter predicts no dynamic instability of the B2
+      phase at 0K.  }
     \label{fig:allmodels_b2}
@@
     \includegraphics[width=\linewidth]{figures/NiTi_B19_ModelsPhononBandstructures.png}
-        \caption{
-          Ground-state phonon bandstructure of B19 phase for all models.
-        }
+    \caption{ Ground-state phonon bandstructure of B19 phase for all
+      models.  }
     \label{fig:allmodels_B19P}
@@
     \includegraphics[width=\linewidth]{figures/NiTi_B19P_ModelsPhononBandstructures.png}
-        \caption{
-          Ground-state phonon bandstructure of \textit{B19}$^\prime$ phase for all models.
-        }
+    \caption{ Ground-state phonon bandstructure of
+      \textit{B19}$^\prime$ phase for all models.  }
     \label{fig:allmodels_B19P}
@@
     \includegraphics[width=\linewidth]{figures/NiTi_BCO_ModelsPhononBandstructures.png}
-        \caption{
-          Ground-state phonon bandstructure of BCO phase for all models.
-        }
+    \caption{ Ground-state phonon bandstructure of BCO phase for all
+      models.  }
     \label{fig:allmodels_B19P}
@@
 \end{figure}
+
+\subsubsection{Mode Grun\"{e}isen Parameter}
+
+The B2 phase of NiTi exhibits notable phonon instabilities,
+particularly at the \textit{M} point in the Brillouin zone. These
+instabilities are characterized by imaginary modes, represented as
+negative phonon frequencies, indicating a dynamical instability in
+this crystal structure. The \textit{M} point, located at the edge of
+the Brillouin zone, corresponds to a wave vector of
+$\left[\frac{1}{2}, \frac{1}{2}, 0\right]$ in units of the reciprocal
+lattice vectors. The phonon dispersion curves for \ce{NiTi} in the B2
+phase reveal a softening of the acoustic modes at the \textit{M}
+point, suggesting a propensity for a structural phase transition. \par
+
+The mode Gruneisen parameter, $\gamma$, plays a pivotal role in
+understanding the stabilization of the NiTi B2 phase under external
+strain, particularly in the context of phonon instabilities at the M
+point. The mode Gruneisen parameter is defined as,
+
+\begin{equation}
+  \label{eq:modegruneisen}
+  \gamma = -\frac{V}{\omega}\left(\frac{\partial \omega}{\partial V}\right)
+\end{equation},
+
+where $V$ is the volume, usually taken at equilibrium, and $\omega$ is
+the phonon frequency. For the \textit{B2} phase of \ce{NiTi}, the
+negative phonon frequencies at the \textit{M} point indicate an
+instability that can be modulated by applying strain. The strain
+perturbs the lattice volume, thereby altering the phonon
+frequencies. The effect of this strain can be quantitatively assessed
+using the mode Gr\"{u}neisen parameter, which describes how the phonon
+frequencies change with volume. A positive $\gamma$ suggests that the
+phonon frequency increases as the volume decreases under strain,
+potentially stabilizing the lattice. In the case of NiTi, applying
+compressive strain may increase the frequencies of the soft modes at
+the \textit{M} point, thus stabilizing the \textit{B2} phase and
+delaying or suppressing the martensitic transformation. This interplay
+between strain and phonon behavior, quantified by $\gamma$, is crucial
+in tailoring the mechanical properties of shape memory and
+superelastic materials.
+
+In fig.~\ref{fig:modegruneisen} the Gr\"{u}neisen parameter for
+\textit{M}-mode in \textit{B2} is shown for different interatomic
+potentials for \ce{NiTi}. The instable modes (i.e., negative
+eigenfrequencies) for the MEAM, M3GNet, and MACE potentials have
+Gr\"{u}neisen parameter which are negative. The negative values
+suggest that compressive isotropic strains will provide stabalizing
+force against displacive phase transformation. At higher frequencies
+the Gr\"{u}neisen parameter is positive, indicating this modes
+experience a decrease in frequency with respect to volume
+increase. The optical modes are less sensitive as indicated by smaller
+values. \par
+
+\begin{figure}[!htp]
+  \script{GruneisenParameters.py} \centering
+  \includegraphics[width=\columnwidth]{figures/Plot_NiTi_M_ModeGruneisen.png}
+  \label{fig:modegruneisen}
+  \caption{The sign-log of the \textit{M}-mode Gr\"{u}neisen
+    parameter, $\gamma$, as a function of eigenfrequency for different
+    interatomic potentials.}
+\end{figure}
+
+The Mutter EAM potential shows negative Gr\"{u}''neisen parameters for
+all eigenfrequencies indicating compressive forces increase
+frequencies, however, none of the \textit{M}-point modes have negative
+eigenfrequencies. \par
+
+In contrast the the MEAM, M3GNet, and MACE potentials, the Zhong EAM
+and CHGNet potentials show positive Gr\"{u}niesen parameters for
+negatives eignefrequencies, suggesting that isotropic tensile strain
+acts to stabalize the mode. \par
+
+To better coberate the results in fig.~\ref{fig:modegruneisen}, the
+phonon bandstructures as a result of isoptroic strain applied in
+compression and tension are show in fig
 
@@
     \includegraphics[width=0.5\linewidth]{figures/NiTi_Ko_B2_StrainsPhononBandstructures.png}
-        \caption{
-           Phonon dispersion of B2 structure for the  (top) EAM Mutter, (middle) EAM Zhong, and (bottom) 2NN-MEAM Ko potentials at different isotropic strains.
-The optical branches have been made more transparent to emphasis the behavior in the TA$_1$ and TA$_2$ branches.
-The color indicates direction from compressive to tensile strain.
-Note that the strain percent shown is with respect to isotropic changes (i.e.
-volume).
-        }
-        \label{fig:mutter_zhong_phonon_b2}
+    \caption{ Phonon dispersion of B2 structure for the (top) EAM
+      Mutter, (middle) EAM Zhong, and (bottom) 2NN-MEAM Ko potentials
+      at different isotropic strains. The optical branches have been
+      made more transparent to emphasis the behavior in the TA$_1$ and
+      TA$_2$ branches. The color indicates direction from compressive
+      to tensile strain. Note that the strain percent shown is with
+      respect to isotropic changes (i.e. volume).  }
+    \label{fig:mutter_zhong_ko_phonon_b2}
   \end{centering}
@@
 
-%ADD Plot of zero strain EAM and first-pricinpals results
+% ADD Plot of zero strain EAM and first-pricinpals results
 
@@
     \includegraphics[width=0.5\linewidth]{figures/NiTi_MACE_B2_StrainsPhononBandstructures.png}
-      \caption{
-        Phonon dispersion of B2 structure for the (top) M3GNet, (middle) CHGNet, and (bottom) MACE potentials for different isotropic strains.
-Similar transcparency is used as done in \ref{fig:mutter_zhong_phonon_b2}.
-      }
+    \caption{ Phonon dispersion of B2 structure for the (top) M3GNet,
+      (middle) CHGNet, and (bottom) MACE potentials for different
+      isotropic strains. Similar transcparency is used as done in
+      \ref{fig:mutter_zhong_phonon_b2}.  }
     \label{fig:gnn_phonon_b2}
@@
 
-The B2 phase of NiTi exhibits notable phonon instabilities, particularly at the \textit{M} point in the Brillouin zone.
-These instabilities are characterized by imaginary modes, represented as negative phonon frequencies, indicating a dynamical instability in this crystal structure.
-The \textit{M} point, located at the edge of the Brillouin zone, corresponds to a wave vector of $\left[\frac{1}{2}, \frac{1}{2}, 0\right]$ in units of the reciprocal lattice vectors.
-The phonon dispersion curves for NiTi in the B2 phase reveal a softening of the acoustic modes at the \textit{M} point, suggesting a propensity for a structural phase transition.
-\par
-
-The mode Gruneisen parameter, $\gamma$, plays a pivotal role in understanding the stabilization of the NiTi B2 phase under external strain, particularly in the context of phonon instabilities at the M point.
-The mode Gruneisen parameter is defined as,
-
-\begin{equation}
-  \label{eq:modegruneisen}
-  \gamma = -\frac{V}{\omega}\left(\frac{\partial \omega}{\partial V}\right)
-\end{equation},
-
-where $V$ is the volume, usually taken at equilibrium, and $\omega$ is the phonon frequency.
-For the B2 phase of NiTi, the negative phonon frequencies at the M point indicate an instability that can be modulated by applying strain.
-The strain perturbs the lattice volume, thereby altering the phonon frequencies.
-The effect of this strain can be quantitatively assessed using the mode Gruneisen parameter, which describes how the phonon frequencies change with volume.
-A positive $\gamma$ suggests that the phonon frequency increases as the volume decreases under strain, potentially stabilizing the lattice.
-In the case of NiTi, applying compressive strain may increase the frequencies of the soft modes at the M point, thus stabilizing the B2 phase and delaying or suppressing the martensitic transformation.
-This interplay between strain and phonon behavior, quantified by $\gamma$, is crucial in tailoring the mechanical properties of shape memory and superelastic materials.
-
-\variable{output/Table_NiTi_M_ModeGruneisen.tex}
+% \variable{output/Table_NiTi_M_ModeGruneisen.tex}
 
@@
 \label{sec:conclusion}
-This word aimed to provide an assement of the of GNN univeral interatomic potentials on the ability to capture the stable ground-state phases of binary and ternary SMA.
-In addition, the dynamic stability of these SMA is crucially evaluated to understand the suitability in observing the onset of displacive phase transformations via the softening of TA branch modes.
-In the case of NiTi we find that all the GNN potentials predict either the Monoclinic or BCO phases as stable over the B2 structure.
-However, not all agree with the recent works by (cite) which have indicated that it is the BCO phase which is the ground-state structure.
+This word aimed to provide an assement of the of GNN univeral
+interatomic potentials on the ability to capture the stable
+ground-state phases of binary and ternary SMA. In addition, the
+dynamic stability of these SMA is crucially evaluated to understand
+the suitability in observing the onset of displacive phase
+transformations via the softening of TA branch modes. In the case of
+NiTi we find that all the GNN potentials predict either the Monoclinic
+or BCO phases as stable over the B2 structure. However, not all agree
+with the recent works by (cite) which have indicated that it is the
+BCO phase which is the ground-state structure.
 
-We observe that the M3GNet and CHGnet capture the M-point instability of the B2 phase, however, no stablizing isotropic strain is observed.
-With regard to the MACE, very good agrement is found with the phonon bandstructure and that from ab-initio methods.
-The ALIGNN potential was found not to be suitable of NiTi.
+We observe that the M3GNet and CHGnet capture the M-point instability
+of the B2 phase, however, no stablizing isotropic strain is
+observed. With regard to the MACE, very good agrement is found with
+the phonon bandstructure and that from ab-initio methods. The ALIGNN
+potential was found not to be suitable of NiTi.
 
-\bibliographystyle{elsarticle-num-names}
-\bibliography{references}
+\bibliographystyle{elsarticle-num-names} \bibliography{references}
 
@@
 
-S.B: Conceptualization, Formal Analysis, Investigation, Methodology, Software, Visualisation, Writing---original draft.
+S.B: Conceptualization, Formal Analysis, Investigation, Methodology,
+Software, Visualisation, Writing---original draft.
 
 \section*{Reproducibility and Data Availability}
-This study was carried out using the reproducibility software \href{https://github.com/showyourwork/showyourwork}{\showyourwork}\cite{Luger2021}, which leverages continuous integration to programmatically download the data from \href{https://zenodo.org/}{zenodo.org}(if needed), create the figures, and compile the manuscript.
-Each figure caption contains two links: one to the dataset stored on zenodo used in the corresponding figure, and the other to the script used to make the figure (at the commit corresponding to the current build of the manuscript).
-The git repository associated to this study is publicly available at \url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP} and allows anyone to re-build the entire manuscript.\par
+This study was carried out using the reproducibility software
+\href{https://github.com/showyourwork/showyourwork}{\showyourwork}\cite{Luger2021},
+which leverages continuous integration to programmatically download
+the data from \href{https://zenodo.org/}{zenodo.org}(if needed),
+create the figures, and compile the manuscript. Each figure caption
+contains two links: one to the dataset stored on zenodo used in the
+corresponding figure, and the other to the script used to make the
+figure (at the commit corresponding to the current build of the
+manuscript). The git repository associated to this study is publicly
+available at
+\url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP} and allows
+anyone to re-build the entire manuscript.\par
 
@@
 
-The code used to generate the phonon configuration settings for each potential mode and structure are shown the table below. Settings are used for both NiTi and PtTi systems.
+The code used to generate the phonon configuration settings for each
+potential mode and structure are shown the table below. Settings are
+used for both NiTi and PtTi systems.
 
@@
 
-The irreducible Brillouin zones for the the different crystals investigated in this study are provided below.
-The high-symmetry points are indicated.
-Additionally the coordinates of the high-symmetry points in the IBZ are shown in the tables below.
-In Ref.
-xxx the  specific implementation of the band path sampling and definitions are provided.
+The irreducible Brillouin zones for the the different crystals
+investigated in this study are provided below. The high-symmetry
+points are indicated. Additionally the coordinates of the
+high-symmetry points in the IBZ are shown in the tables below. In
+Ref. xxx the specific implementation of the band path sampling and
+definitions are provided.
 
@@
   \end{subfigure}
-  \caption{The irreducible Brillouin zones for \textit{B2}(\#221), \textit{B19}(\#51), \textit{B19'}(\#11), and \textit{BCO}(\#63) crystal structures}
+  \caption{The irreducible Brillouin zones for \textit{B2}(\#221),
+    \textit{B19}(\#51), \textit{B19'}(\#11), and \textit{BCO}(\#63)
+    crystal structures}
   \label{fig:ibz}

### 2024-01-18 src/scripts/CalculateEOS.py

@@ -1,15 +1,14 @@
 import os
-import paths
 
-import numpy as np
 import matplotlib.pyplot as plt
-
-from ase.db import connect
+import numpy as np
+import paths
 from ase.constraints import StrainFilter
-from ase.units import kJ
+from ase.db import connect
 from ase.eos import EquationOfState
-
+from ase.units import kJ
 from MinimizeStructure import minimize_structure
+
 
 def to_scalar(value):
     if isinstance(value, np.ndarray) and value.size == 1:
@@ -17,12 +16,14 @@ def to_scalar(value):
     else:
         return value
 
-def calculate_eos(structure,
-                  potential,
-                  strain,
-                  eos_eq="sj",
-                  plot=True,
-                  ):
+
+def calculate_eos(
+    structure,
+    potential,
+    strain,
+    eos_eq="sj",
+    plot=True,
+):
     """Calculates the equation of state (EOS) for a given structure under strain.
 
     This function applies isotropic strain to the given structure, computes the energy
@@ -47,49 +48,49 @@ def calculate_eos(structure,
             - energies (List[float]): List of energies corresponding to the strains.
     """
 
-    
     structure.set_calculator(potential[1])
 
     sf = StrainFilter(structure)
     energies = []
     volumes = []
     for e in strain:
-        isotropic_strain = np.array([e,e,e, 0, 0, 0])
+        isotropic_strain = np.array([e, e, e, 0, 0, 0])
         sf.set_positions(isotropic_strain)
         energy = to_scalar(sf.get_potential_energy())
         volume = sf.get_cell().volume
         energies.append(energy)
         volumes.append(volume)
-        
-    eos = EquationOfState(volumes, energies,eos=eos_eq)
-    v0, e0, B = eos.fit()
 
+    eos = EquationOfState(volumes, energies, eos=eos_eq)
+    v0, e0, B = eos.fit()
 
     if plot:
         species = "".join(set(structure.get_chemical_symbols()))
         name = structure.info["structure_name"]
         model = potential[0]
-        plot_eos(eos,model,species,name)
-        
+        plot_eos(eos, model, species, name)
 
     return (eos_eq, v0, e0, B, volumes, energies)
 
-def plot_eos(eos,potential_name,specie_names,structure_name):
+
+def plot_eos(eos, potential_name, specie_names, structure_name):
     # Plot and save EOS
[…]

### 2024-01-18 src/scripts/CalculateElastic.py

@@ -1,14 +1,12 @@
-import paths
 from copy import deepcopy
 
-from ase.db import connect
 import ase.units as units
-from ase.spacegroup import crystal
+import paths
 from ase.calculators.lammpslib import LAMMPSlib
-
-from elastic import get_elementary_deformations
+from ase.db import connect
+from ase.spacegroup import crystal
+from elastic import get_elastic_tensor, get_elementary_deformations
 from elastic.elastic import get_cij_order
-from elastic import get_elastic_tensor
 
 
 def serial_calculate(systems, calculator):
@@ -77,7 +75,7 @@ def calculate_elastic_constants(
     entry = db.get(
         chemsys=chemsys, structure_name=structure_name, model_name=potname, relaxed=True
     )
-    
+
     spg = entry.spacegroup
     structure = crystal(entry.toatoms(), spacegroup=spg)
     # calculator.clean = potential[1].reset

### 2024-01-18 src/scripts/CalculatePhonons.py

@@ -1,15 +1,11 @@
 import paths
-
-from ase.spacegroup import get_spacegroup
-from ase.spacegroup import crystal
-
-from ase.phonons import Phonons
 from ase.db import connect
+from ase.phonons import Phonons
+from ase.spacegroup import crystal, get_spacegroup
 from ase.spectrum.band_structure import BandStructure
-
-from Structures import get_bandpath
 from Config import get_phonon_config
 from PlotPhonons import plot_default_phonons
+from Structures import get_bandpath
 
 EV_to_THz = 241.799050402293e0
 
@@ -204,6 +200,7 @@ def calculate_phonons(
             "dyn_mat": dyn_mat,
             "volume": volume,
             "stress": stress,
+            "supercell": supercell,
         }
 
     db.update(entry.id, data={"strain_phonons": strain_phonons})

### 2024-01-18 src/scripts/Calculators.py

@@ -1,5 +1,5 @@
-import paths
 import numpy as np
+import paths
 from ase.calculators.eam import EAM
 
 # TODO: For M3GNet and CHGNet is stress_weight=1.0*GPa or stress_weight=1.0?
@@ -126,7 +126,7 @@ def get_ase_calculator(model="MutterASE"):
     elif model == "MACE":
         from mace.calculators import MACECalculator
 
-        #chkpoint = "2023-08-14-mace-universal.model"
+        # chkpoint = "2023-08-14-mace-universal.model"
         chkpoint = "2023-12-03-mace-128-L1_epoch-199.model"
         asecalc = MACECalculator(
             model_paths=str(paths.static / chkpoint),
@@ -134,24 +134,53 @@ def get_ase_calculator(model="MutterASE"):
             default_dtype="float32",
         )
 
-        #Alternative API for newer MACE
-        #from mace.calculators import mace_mp
-        #asecalc = mace_mp() # return the default medium ASE calculator equivalent to mace_mp(model="medium")
-        #asecalc = mace_mp(model="large") # return a larger model
+        # Alternative API for newer MACE
+        # from mace.calculators import mace_mp
+        # asecalc = mace_mp() # return the default medium ASE calculator equivalent to mace_mp(model="medium")
+        # asecalc = mace_mp(model="large") # return a larger model
 
     elif model == "ALIGNN":
         from alignn.ff.ff import AlignnAtomwiseCalculator, default_path
 
         alignn_params = default_path()
-        asecalc = AlignnAtomwiseCalculator(path=alignn_params, device="cpu",stress_wt = 1.0 * GPa)
+        asecalc = AlignnAtomwiseCalculator(
+            path=alignn_params, device="cpu", stress_wt=1.0 * GPa
+        )
 
     elif model == "DeepMD":
         from deepmd.calculator import DP
 
         model_params = str(paths.static / "NiTi_DeepMD.pb")
         asecalc = DP(model=model_params)
-        
+
     return asecalc
+
+
+def get_gpaw_calculator(
+    xc="PBE",
+    cutoff=680,
+    kpts=(8, 8, 8),
+):
+    from gpaw import GPAW, PW, FermiDirac
+
+    # import gpaw.mpi as mpi
+    # world_size = mpi.world.size  # Total number of MPI processes
+    # all_ranks = list(range(world_size))  # List of all ranks
[…]

### 2024-01-18 src/scripts/Config.py

@@ -1,4 +1,3 @@
-
 def get_phonon_config(structure_name, potential):
     """Settings for the phonon calculations
 
@@ -14,19 +13,20 @@ def get_phonon_config(structure_name, potential):
             "Mutter": ((8, 8, 8), 0.0935, 200),
             "Zhong": ((8, 8, 8), 0.0935, 200),
             "Ko": ((8, 8, 8), 0.01, 200),
-            "Kavousi" : ((8, 8, 8), 0.01, 200),
+            "Kavousi": ((8, 8, 8), 0.01, 200),
             "Kim": ((8, 8, 8), 0.01, 200),
             "M3GNet": ((8, 8, 8), 0.010, 200),
             "CHGNet": ((8, 8, 8), 0.030, 200),
             "MACE": ((8, 8, 8), 0.010, 200),
             "ALIGNN": ((6, 6, 6), 0.050, 200),
-            "DeepMD": ((8, 8, 8),0.03, 200),
+            "DeepMD": ((8, 8, 8), 0.03, 200),
+            "GPAW": ((2, 2, 2), 0.03, 200),
         },
         "B19": {
             "Mutter": ((7, 9, 7), 0.05, 200),
             "Zhong": ((7, 9, 7), 0.05, 200),
             "Ko": ((7, 9, 7), 0.03, 200),
-            "Kavousi" : ((7, 9, 7), 0.03, 200),
+            "Kavousi": ((7, 9, 7), 0.03, 200),
             "Kim": ((7, 9, 7), 0.03, 200),
             "M3GNet": ((7, 9, 7), 0.05, 200),
             "CHGNet": ((7, 9, 7), 0.05, 200),
@@ -38,7 +38,7 @@ def get_phonon_config(structure_name, potential):
             "Mutter": ((8, 6, 4), 0.05, 200),
             "Zhong": ((8, 6, 4), 0.05, 200),
             "Ko": ((8, 6, 4), 0.01, 200),
-            "Kavousi" : ((8, 6, 4), 0.01, 200),
+            "Kavousi": ((8, 6, 4), 0.01, 200),
             "Kim": ((8, 6, 4), 0.01, 200),
             "M3GNet": ((8, 6, 4), 0.05, 200),
             "CHGNet": ((8, 6, 4), 0.05, 200),
@@ -50,7 +50,7 @@ def get_phonon_config(structure_name, potential):
[…]

### 2024-01-18 src/scripts/GPAW_Runner.py

@@ -1,25 +1,35 @@
 import argparse
 import os
+import sys
+
+sys.path.append(os.path.dirname(os.path.abspath(__file__)))
+
+# Surpress warning from dependencies for which I can do nothing about.
 import warnings
 
 import numpy as np
 import paths
+from ase import Atoms
 from ase.db import connect
 from CalculateElastic import calculate_elastic_constants
 from CalculateEOS import get_min_structure_and_calculate_eos
 from CalculatePhonons import calculate_phonons
-from Calculators import get_ase_calculator
+from Calculators import get_gpaw_calculator
+from gpaw import mpi
 from Structures import get_structure
 
-# Surpress warning from dependencies for which I can do nothing about.
 warnings.filterwarnings("ignore", category=UserWarning, module="ase.spacegroup")
-warnings.filterwarnings(
-    "ignore", category=UserWarning, module="dgl.backend.pytorch.tensor"
-)
-warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_cuda'.*")
-warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_cudnn'.*")
-warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_mps'.*")
-warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_mkldnn'.*")
+
+
+def isolated_atom(element, a=10.0):
+    atom = Atoms(
+        element,
+        scaled_positions=[[0.5, 0.5, 0.5]],
+        cell=[a, a, a],
+        pbc=[True, True, True],
+    )
+
+    return atom
 
 
 def min_eos_calc(dbname, structure, potential, strain_info):
@@ -80,13 +90,14 @@ def phonon_calc(dbname, structure, potential, strain_info):
     -----------------------------------------------------------------
     """
     )
-    strain = np.linspace(*strain_info)
+
     calculate_phonons(
         dbname,
         structure.info["chemsys"],
         structure.info["structure_name"],
         potential,
-        strain=strain,
+        strain=strain_info,
+        plotting=True,
     )
     print(
         f"""
@@ -126,50 +137,22 @@ def simulation():
     return ValueError("Simulation routine not implemented!")
 
 
-def main():
-    """
-    Provides top level run script. Be mindful that correct python environment is
-    activated for selected model (i.e., ASE Calculator) otherwise routines will fail to run.
-    """
+if __name__ == "__main__":
     parser = argparse.ArgumentParser(
[…]

### 2024-01-18 src/scripts/GenerateEquilibriumTable.py

@@ -1,17 +1,13 @@
-import sys
 import re
-import numpy as np
-from math import gcd
+import sys
 from functools import reduce
+from math import gcd
+
+import numpy as np
 import paths
-
-
 from ase.db import connect
-from ase.spacegroup import get_basis
-from ase.spacegroup import Spacegroup, get_spacegroup
-
-
-from TableConfig import SPACEGROUP_MAP, ORDER
+from ase.spacegroup import Spacegroup, get_basis, get_spacegroup
+from TableConfig import ORDER, SPACEGROUP_MAP
 
 
 def get_element_basis_list(structure, spacegroup, tol=1e-3):
@@ -70,9 +66,9 @@ def generate_structure_table(dbname, chemsys, output_filename):
     sorder = ORDER[chemsys]["structure"]
     default_order = max(sorder.values())
     unique_structure_types = set(entry.structure_name for entry in db.select())
-    structure_order = sorted(unique_structure_types, key=lambda x: sorder.get(x, default_order))
-
-    
+    structure_order = sorted(
+        unique_structure_types, key=lambda x: sorder.get(x, default_order)
+    )
 
     with open(output_filename, "w") as file:
         file.write("\\begin{longtable}{|l|c|}\n")
@@ -103,8 +99,12 @@ def generate_structure_table(dbname, chemsys, output_filename):
 
             # Filter and sort based on models defined in ORDER
             sentries = sorted(
-                [entry for entry in entries if entry.model_name in ORDER[chemsys]["model"]],
-                key=lambda entry: ORDER[chemsys]["model"][entry.model_name]
+                [
+                    entry
+                    for entry in entries
+                    if entry.model_name in ORDER[chemsys]["model"]
+                ],
+                key=lambda entry: ORDER[chemsys]["model"][entry.model_name],
             )
 
             for entry in sentries:

### 2024-01-18 src/scripts/GruneisenParameters.py

@@ -1,30 +1,22 @@
-import paths
 import sys
-import numpy as np
-import numpy.polynomial.polynomial as poly
 
 # TODO: Move to ConfigPlots and use for all.
 import matplotlib.pyplot as plt
-import matplotlib.style as style
-# Tufte-inspired style
-#style.use('seaborn-whitegrid')
-plt.rcParams['font.family'] = 'serif'
-plt.rcParams['font.serif'] = 'Times New Roman'
-plt.rcParams['axes.labelsize'] = 10
-plt.rcParams['axes.titlesize'] = 12
-plt.rcParams['xtick.labelsize'] = 8
-plt.rcParams['ytick.labelsize'] = 8
-plt.rcParams['legend.fontsize'] = 8
-plt.rcParams['figure.dpi'] = 300
+import numpy as np
+import numpy.polynomial.polynomial as poly
+import paths
+import scienceplots
 
+plt.style.use(["science", "scatter"])
 
 from ase.db import connect
-
 from TableConfig import ORDER
+
 
 # Function to apply the signed log transformation
 def log_transform(val):
     return np.sign(val) * np.log10(abs(val) + 1)
+
 
 def find_index(data, label):
     """
@@ -238,7 +230,7 @@ def generate_mode_gruneisen_table(
             # Guarenteed to be correct! The proper way is to identify what the
             # cross product of the mode eigenvector  w.r.t the q-vector. This needs
             # to be done in the CalculatePhonons.py
-            file.write("Model & TA$_1$ & TA$_2$ & LA & TO$_1$ & TO$_2$ & LO \\\\\n")
+            file.write("Model & TA$_1$ & TA$_2$ & LA & LO & TO$_1$ & TO$_2$ \\\\\n")
             file.write("\\hline\n")
 
             entries = db.select(chemsys=chemsys, structure_name=name)
@@ -290,64 +282,30 @@ def generate_mode_gruneisen_plot(
         ],
         key=lambda entry: ORDER[chemsys]["model"][entry.model_name],
     )
-    # Figure size for one-column width (inches)
-    plt.figure(figsize=(3.5, 2.5))
 
+    plt.figure()
     for entry in sentries:
         model_name = entry.get("model_name", "NA")
         omega0, gammas = calculate_mode_gruneisen(entry)
-        plt.plot(omega0, log_transform(gammas), label=f"{model_name}", marker='o', markersize=4, linestyle='-')
+        print(omega0)
[…]

### 2024-01-18 src/scripts/PlotEOS.py

@@ -1,14 +1,17 @@
 import sys
-import numpy as np
 import warnings
-import paths
 
 import matplotlib.pyplot as plt
 import matplotlib.ticker as ticker
-from matplotlib.ticker import MultipleLocator
-from cycler import cycler
+import numpy as np
+import paths
+import scienceplots
 import seaborn as sns  # For color palette
 from ase.db import connect
+from cycler import cycler
+from matplotlib.ticker import MultipleLocator
+
+plt.style.use(["science", "scatter", "no-latex"])
 
 from PlotConfigs import SUBPLOT_ORDER
 
@@ -25,11 +28,11 @@ def plot_eos(
     n_colors=10,
     wspace=0.3,
     hspace=0.3,
-    linewidth=1,
+    linewidth=0.75,
     linestyle="-",
-    fontsize=8,
-    tick_fontsize=6,
-    legend_fontsize=8,
+    fontsize=5,
+    tick_fontsize=5,
+    legend_fontsize=6,
     legend_ncol=5,
     label_notes="EOS_Labels.txt",
 ):
@@ -49,7 +52,7 @@ def plot_eos(
     structure_names = set(row.structure_name for row in db.select(chemsys=chemsys))
 
     # Create a figure with a 2x3 grid of subplots
-    fig, axs = plt.subplots(2, 3, figsize=figsize)
+    fig, axs = plt.subplots(2, 3)  # , figsize=figsize)
     axs = axs.ravel()  # Flatten the array of axes for easier indexing
 
     # Create a color cycle iterator using a slightly modified color palette
@@ -71,10 +74,10 @@ def plot_eos(
             ax.set_prop_cycle(cycler(color=color_cycle))
             ax.set_xlim(xlim)
             ax.xaxis.set_major_locator(MultipleLocator(0.1))
-            ax.yaxis.set_major_locator(ticker.MaxNLocator(nbins=6))
+            ax.yaxis.set_major_locator(ticker.MaxNLocator(nbins=4))
 
-            ax.tick_params(axis="both", direction="in", length=2)
-            ax.tick_params(labelsize=tick_fontsize)
+            # ax.tick_params(axis="both", direction="in", length=2)
+            # ax.tick_params(labelsize=tick_fontsize)
 
             ax.axvline(x=1.0, color="lightgray", linestyle=":", linewidth=linewidth)
 
@@ -82,6 +85,7 @@ def plot_eos(
                 if structure_name in ignore_list:
[…]

### 2024-01-18 src/scripts/SimulatePhaseChange.py

@@ -1,14 +1,13 @@
 import sys
-import paths
-import numpy as np
-from ase.db import connect
-from ase.io import Trajectory
-from ase.md import MDLogger
-from ase.units import fs, bar, kg, m, GPa
-from ase.md.velocitydistribution import MaxwellBoltzmannDistribution
-from ase.md.npt import NPT
-from ase.io import read, write
 
+import numpy as np
+import paths
+from ase.db import connect
+from ase.io import Trajectory, read, write
+from ase.md import MDLogger
+from ase.md.npt import NPT
+from ase.md.velocitydistribution import MaxwellBoltzmannDistribution
+from ase.units import GPa, bar, fs, kg, m
 from Calculators import get_ase_calculator
 
 
@@ -36,8 +35,8 @@ class CustomMDLogger(MDLogger):
     def __init__(self, *args, **kwargs):
         super().__init__(*args, **kwargs)
         self.cumulative_time = 0.0 / (1000 * fs)
-        self.density = (0.0 * 1.0e-3 * kg) / ((1.0e-2 * m)**3)
-        
+        self.density = (0.0 * 1.0e-3 * kg) / ((1.0e-2 * m) ** 3)
+
     def __call__(self):
         # Obtain potential and kinetic energy, and temperature
         epot = self.atoms.get_potential_energy()
@@ -57,23 +56,22 @@ class CustomMDLogger(MDLogger):
                 factor = (self.cumulative_time + current_time) % self.cumulative_time
             except ZeroDivisionError:
                 factor = 0.0e0
-            
+
             if not factor != 0.0e0:
                 display_time = self.cumulative_time + current_time
             else:
                 display_time = self.cumulative_time
 
-
         dat = (display_time, epot + ekin, epot, ekin, temp)
-        
+
         # Include stress in the log if requested
         if self.stress:
             stress_data = tuple(self.atoms.get_stress(include_ideal_gas=True) / GPa)
             dat += stress_data
 
-        #self.add_density()
+        # self.add_density()
 
-        #dat += (self.density)
+        # dat += (self.density)
 
         # Write data to the logfile
         self.logfile.write(self.fmt % dat)
@@ -88,12 +86,13 @@ class CustomMDLogger(MDLogger):
         mass = self.atoms.get_masses().sum() * grams
         volume = self.atoms.cell.volume * cm**3
         if volume > 0:  # Check to avoid division by zero
[…]

### 2024-01-18 src/scripts/WriteCodeToTex.py

@@ -1,13 +1,14 @@
-import paths
 import re
 import sys
+
+import paths
 
 
 def extract_function(source_file, function_name):
     with open(source_file, "r") as file:
         code = file.read()
 
-    pattern = re.compile(rf'(def {function_name}\(.*?:\n(?:\s+.*\n)+)', re.MULTILINE)
+    pattern = re.compile(rf"(def {function_name}\(.*?:\n(?:\s+.*\n)+)", re.MULTILINE)
     match = pattern.search(code)
 
     if match:

### 2024-01-30 src/tex/ms.tex
Plot density. MS edits.

@@
 appendix~\ref{sec:appx_reproduce} and corresponding github
-repo\footnote{\url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP}}.\par
+repo\footnote{\url{https://github.com/stefanbringuier/SMA_Phonons_GNNIP}}. We
+note that in order to achieve the different computing environments
+required by the interatomic potentials we leverage the ability to specify the
+python computing environments using \href{https://github.com/showyourwork/showyourwork}{\showyourwork}
+package.\par
 
@@
 shown in fig~\ref{fig:niti_structures} and are adapted for PtTi as
-well given same spacegroup symmetries. The energy and forces on the
+well given same space group symmetries. The energy and forces on the
 atoms are determined using the ASE calculator interfaces for EAM,
-MEAM, M3GNet, CHGNet, and
-MACE.\cite{Mutter2010,Zhong2011,Ko2015,Kim2017,Chen2022,Deng2023,Batatia2022}
+MEAM, M3GNet, CHGNet, and MACE.\cite{Mutter2010,Zhong2011,Ko2015,Kim2017,Chen2022,Deng2023,Batatia2022}
 The EAM and MEAM potential interfaces are achieved through the use of
-LAMMPS python library \cite{Thompson2022} and ASE. The ALIGNN ASE
-calculator interface was also implemented, however, we found that the
-predictions for NiTi and PtTi to be unreliable are not included in
-this study.\par
+LAMMPS python library \cite{Thompson2022} and ASE.
+We have also implemented calculators for potentials by \cite{Kavousi2019}, \cite{Wang2018}, and \cite{Choudhary2021} but are not shown in the main manuscript of this work.\par
 
 For each combination of potential/model, chemical system, and phase we
-perform calcualtions for the structure minimization, equation of state
+perform calculations for the structure minimization, equation of state
 (EOS), phonons, and elastic constants. To determine the
-finite-temperature phase transitions we utlize an isothermal-isobaric
+finite-temperature phase transitions we utilize an isothermal-isobaric
 ensemble via molecular dynamics simulations. The thermostating and
 barostating is done with Nose-Hoover and Parrinello-Rahman dynamics
-and the tempertaure is adjusted in a quasi-static manner which entails
-equlibrating the simulation for 10 ns while decrementing the
-temperature by 10 Kelvin; the pressure is kept constant at 1 bar. We
-note that in order to achieve the different computing environments
-required by the GNN potentials we leverage the ability to specify the
-python computing environments using
-\href{https://github.com/showyourwork/showyourwork}{\showyourwork}
-package.\par
+and the ramping down or up of temperature is adjusted in a dynamical manner
+which entails adjusting the target thermostat temperature every few timesteps
+based on the ramp-rate. The pressure is kept constant at 1 bar.
 
@@
 			\ce{Ni} and \ce{Ti} as gray and unit cells are highlighted with
-      parametes shown on axis. For all unit cells the $\alpha$ and
+			parameters shown on axis. For all unit cells the $\alpha$ and
 			$\gamma$ angles are 90.0$^\circ$. For PtTi, (a-b) structures are
@@
 optimized using the FIRE\cite{Bitzek2006} local optimizer with the
-forces converged to $2\cdot 10^{-3}$eV/\AA. We observed that other
+forces converged to $2\cdot 10^{-3}$eV/\AA\@. We observed that other
 local optimization routines find the same ground-state structures. To
@@
 utlize a constraint and filter during the optimization process. \par
-
-In fig.~\ref{fig:opt_density} the density of the structures for
-different potentials is shown. We observe that in general the
-densities and lattice parameters are within agreement with previously
-reported values. The optimized lattice parameters as well as basis
-atoms for all structures and models are provided in appendix
-\ref{sec:opt_structures}. \par
-
-\begin{figure}[ht!]
-  \label{fig:opt_density}
-  \centering
-  \caption{}
-\end{figure}
 
@@
 
-\subsubsection{Phonon Bandstructures}
+\subsubsection{Phonon band structure}
 
@@
 the same isotropic strain in eq.~\ref{eq:isotropic_strain} is used and
-the repsective phonon bandstructures are calculated. These are then
+the respective phonon band structure are calculated. These are then
 plotted to observed the onset or change in stability or instability of
@@
 calculation requires the use of crystal symmetry along with atomic
-displacements to determine respective elastic constant componenets. We
+displacements to determine respective elastic constant components. We
 use the \textit{Elastic}\cite{Jochym2018} python package which
 provides routines for generating the proper atomic displacement matrix
-given specific crystal spacegroup. The elastic constants are
+given specific crystal space group. The elastic constants are
 determined by solving a set of linear equations using singular value
@@
 \textit{B19'} and \textit{B2} phases. It was latter shown by
-\cite{Zhong2011} that some adjustments to the EAM parameters was
+\cite{Zhong2011} that some adjustments to the EAM parameters were
 needed to properly capture the finite-temperature displacement
@@
 parameterization of NiTi that demonstrated very good predictive
-performance and improvements over ealier potentials.
+performance and improvements over earlier potentials.
 
@@
 descriptor matrix from atomic configurations that preserves
-permutations and symmetries. The descriptors are then which feed into
+permutations and symmetries. The descriptors are then feed into
 a multilayer perceptron to calculate energies and forces. The neural
@@
 
-In the work of \citet{Tang2022} ground-state predictions for
+The ground-state predictions for
 \textit{B2}, \textit{B19'}, \textit{B19}, and
 \textit{BCO}\footnote{Referred to as \textit{B33} prototype in
-  \citet{Tang2022}} were provided and showed good agreement. However,
+	\citet{Tang2022}} in \citet{Tang2022} showed good agreement. However,
 we note that deviation from Ni-Ti 1:1 stoichiometry or addition of a
 third species was not investigated nor the inclusion of additional
-alloying species. \par
+alloying species was explored. \par
 
@@
 ($>$ 80 elements) and therefore are not restricted to specific
-chemistries or stoichiometries. They training data also includes
-various phases for a given chemical system. This permits great
-flexibility. We do not that fine-tuning, the process of adjusting GNN
-weights and biases using additional training data, can be performed
-with the GNN studied in this work. However, this is out of the scope
+chemistries or stoichiometries. The training data also includes
+various phases for a given chemical system which permits great
+flexibility, although broad transferability still requires verification.
+This allows for investigation of both non-stoichiometric and multi-species
+SMA which previous physics and ML potentials were not developed for. We do
+note that fine-tuning, the process of adjusting GNN weights and biases
+using additional training data, can be performed
+with the GNNs studied in this work. However, this is out of the scope
 for this manuscript, given resources and time needed to generate such
@@
 
-One way to understand the phase stability for a given composition is
+In fig.~\ref{fig:opt_density} the density of the structures for
+different potentials is shown. We observe that in general the
+densities and lattice parameters are within agreement with previously
+reported values. The optimized lattice parameters as well as basis
+atoms for all structures and models are provided in appendix~\ref{sec:opt_structures}. \par
+
+\begin{figure}[ht!]
+	\label{fig:opt_density}
+	\centering
+	\includegraphics[options]{figures/NiTi_Equilibrium_Density.png}
+	\caption{Equilibrium Density for \ce{NiTi} structures with different models.}
+\end{figure}
+
+To asses the relative phase stability for a given composition is
 to compare the cohesive energy, which is defined as,
@@
 	\label{eq:cohesive_energy}
-  E_{c} = \frac{E_{\text{bulk}}-\sum_i^N n_i E^_{\text{atom_i}}}{N}
+	E_{c} = \frac{E_{\text{bulk}} - \sum_{i}^{N} n_i E_{\text{atom}_i}}{N}
 \end{equation}
@@
 With $E_{\text{bulk}}$ the total energy of the system,
-$\sum_i^N n_i E_{\text{atom_i}}$ is the sum over isolated atom
+$\sum_i^N n_i E_{\text{atom}_i}$ is the sum over isolated atom
 energies for $n_i$ species, and $N$ the number of atoms in bulk
@@
 	\begin{centering}
-    \includegraphics[width=\linewidth]{figures/NiTi_B2_ModelsPhononBandstructures.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B2_ModelsPhononband structure.png}
 		\caption{ Ground-state, no isotropic strain, Phonon bandstructure
@@
 	\begin{centering}
-    \includegraphics[width=\linewidth]{figures/NiTi_B19_ModelsPhononBandstructures.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B19_ModelsPhononband structure.png}
 		\caption{ Ground-state phonon bandstructure of B19 phase for all
@@
 	\begin{centering}
-    \includegraphics[width=\linewidth]{figures/NiTi_B19P_ModelsPhononBandstructures.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B19P_ModelsPhononband structure.png}
 		\caption{ Ground-state phonon bandstructure of
@@
 	\begin{centering}
-    \includegraphics[width=\linewidth]{figures/NiTi_BCO_ModelsPhononBandstructures.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_BCO_ModelsPhononband structure.png}
 		\caption{ Ground-state phonon bandstructure of BCO phase for all
@@
 To better coberate the results in fig.~\ref{fig:modegruneisen}, the
-phonon bandstructures as a result of isoptroic strain applied in
+phonon band structure as a result of isoptroic strain applied in
 compression and tension are show in fig
@@
 	\begin{centering}
-    \includegraphics[width=0.5\linewidth]{figures/NiTi_Mutter_B2_StrainsPhononBandstructures.png}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_Mutter_B2_StrainsPhononband structure.png}
 		\vspace{1mm}
-    \includegraphics[width=0.5\linewidth]{figures/NiTi_Zhong_B2_StrainsPhononBandstructures.png}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_Zhong_B2_StrainsPhononband structure.png}
 		\vspace{1mm}
-    \includegraphics[width=0.5\linewidth]{figures/NiTi_Ko_B2_StrainsPhononBandstructures.png}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_Ko_B2_StrainsPhononband structure.png}
 		\caption{ Phonon dispersion of B2 structure for the (top) EAM
@@
 	\begin{centering}
-    \includegraphics[width=0.5\linewidth]{figures/NiTi_M3GNet_B2_StrainsPhononBandstructures.png}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_M3GNet_B2_StrainsPhononband structure.png}
 		\vspace{1mm}
-    \includegraphics[width=0.5\linewidth]{figures/NiTi_CHGNet_B2_StrainsPhononBandstructures.png}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_CHGNet_B2_StrainsPhononband structure.png}
 		\vspace{1mm}
-    \includegraphics[width=0.5\linewidth]{figures/NiTi_MACE_B2_StrainsPhononBandstructures.png}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_MACE_B2_StrainsPhononband structure.png}
 		\caption{ Phonon dispersion of B2 structure for the (top) M3GNet,

### 2024-01-30 src/scripts/PlotDensity.py

@@ -1,0 +1,96 @@
+import sys
+
+import matplotlib.pyplot as plt
+import paths
+import scienceplots
+from matplotlib.ticker import MaxNLocator
+
+plt.style.use(["science", "scatter", "no-latex"])
+
+from ase.db import connect
+from PlotConfigs import ORDER, STRUCTURE_ORDER
+
+
+def plot_density(dbname, chemsys, output_filename):
+    db = connect(dbname)
+
+    sorder = STRUCTURE_ORDER[chemsys]
+    default_order = max(sorder.values())
+    unique_structure_types = set(entry.structure_name for entry in db.select())
+    structure_order = sorted(
+        unique_structure_types, key=lambda x: sorder.get(x, default_order)
+    )
+
+    # Data containers
+    structure_names = []
+    model_names = set()
+    density_data = {}
+
+    # Collecting data
+    for name in structure_order:
+        entries = db.select(chemsys=chemsys, structure_name=name)
+        structure_names.append(name.replace("_", "-"))
+        density_data[name] = {}
+
+        for entry in entries:
+            model_name = entry.get("model_name", "NA")
+            density = entry.get("density", 0)  # Default to 0 if not available
+            if model_name in ORDER[chemsys]:
+                density_data[name][model_name] = density
+                model_names.add(model_name)
+
+    # structure_names = sorted(structure_names)
+    model_names = sorted(model_names, key=lambda x: ORDER[chemsys].get(x, float("inf")))
+
+    # Fancy colors and patterns
+    colors = ["red", "green", "blue", "cyan", "magenta", "yellow", "black"]
+    patterns = ["//", "+", "*", "x", "o", "-", "O", "||", "++"]
+
+    # Plotting
+    fig, ax = plt.subplots()
+    bar_width = 0.125
+    space_between_bars = 0.075  # Adjust this value as needed
+    indices = list(range(len(structure_names)))
+    num_models = len(model_names)
+    group_centers = [
+        1.65 * x + (num_models / 2) * (bar_width + space_between_bars) for x in indices
+    ]
+
+    for i, model_name in enumerate(model_names):
+        offset = [(1.5 * x + i * (bar_width + space_between_bars)) for x in indices]
+        density_values = [
+            density_data[struct].get(model_name, 0) for struct in structure_order
+        ]
+        # Assign color and pattern
[…]

### 2024-01-31 src/tex/ms.tex
Auctex formatting

@@
 note that in order to achieve the different computing environments
-required by the interatomic potentials we leverage the ability to specify the
-python computing environments using \href{https://github.com/showyourwork/showyourwork}{\showyourwork}
+required by the interatomic potentials we leverage the ability to
+specify the python computing environments using
+\href{https://github.com/showyourwork/showyourwork}{\showyourwork}
 package.\par
@@
 atoms are determined using the ASE calculator interfaces for EAM,
-MEAM, M3GNet, CHGNet, and MACE.\cite{Mutter2010,Zhong2011,Ko2015,Kim2017,Chen2022,Deng2023,Batatia2022}
+MEAM, M3GNet, CHGNet, and
+MACE.\cite{Mutter2010,Zhong2011,Ko2015,Kim2017,Chen2022,Deng2023,Batatia2022}
 The EAM and MEAM potential interfaces are achieved through the use of
-LAMMPS python library \cite{Thompson2022} and ASE.
-We have also implemented calculators for potentials by \cite{Kavousi2019}, \cite{Wang2018}, and \cite{Choudhary2021} but are not shown in the main manuscript of this work.\par
+LAMMPS python library \cite{Thompson2022} and ASE. We have also
+implemented calculators for potentials by \cite{Kavousi2019},
+\cite{Wang2018}, and \cite{Choudhary2021} but are not shown in the
+main manuscript of this work.\par
 
@@
 barostating is done with Nose-Hoover and Parrinello-Rahman dynamics
-and the ramping down or up of temperature is adjusted in a dynamical manner
-which entails adjusting the target thermostat temperature every few timesteps
-based on the ramp-rate. The pressure is kept constant at 1 bar.
+and the ramping down or up of temperature is adjusted in a dynamical
+manner which entails adjusting the target thermostat temperature every
+few timesteps based on the ramp-rate. The pressure is kept constant at
+1 bar.
 
@@
 descriptor matrix from atomic configurations that preserves
-permutations and symmetries. The descriptors are then feed into
-a multilayer perceptron to calculate energies and forces. The neural
+permutations and symmetries. The descriptors are then feed into a
+multilayer perceptron to calculate energies and forces. The neural
 network interatomic potential demonstrated good accuracy of
@@
 
-The ground-state predictions for
-\textit{B2}, \textit{B19'}, \textit{B19}, and
-\textit{BCO}\footnote{Referred to as \textit{B33} prototype in
-	\citet{Tang2022}} in \citet{Tang2022} showed good agreement. However,
-we note that deviation from Ni-Ti 1:1 stoichiometry or addition of a
-third species was not investigated nor the inclusion of additional
-alloying species was explored. \par
+The ground-state predictions for \textit{B2}, \textit{B19'},
+\textit{B19}, and \textit{BCO}\footnote{Referred to as \textit{B33}
+	prototype in \citet{Tang2022}} in \citet{Tang2022} showed good
+agreement. However, we note that deviation from Ni-Ti 1:1
+stoichiometry or addition of a third species was not investigated nor
+the inclusion of additional alloying species was explored. \par
 
@@
 various phases for a given chemical system which permits great
-flexibility, although broad transferability still requires verification.
-This allows for investigation of both non-stoichiometric and multi-species
-SMA which previous physics and ML potentials were not developed for. We do
-note that fine-tuning, the process of adjusting GNN weights and biases
-using additional training data, can be performed
-with the GNNs studied in this work. However, this is out of the scope
-for this manuscript, given resources and time needed to generate such
-training data for SMAs. \par
+flexibility, although broad transferability still requires
+verification.  This allows for investigation of both
+non-stoichiometric and multi-species SMA which previous physics and ML
+potentials were not developed for. We do note that fine-tuning, the
+process of adjusting GNN weights and biases using additional training
+data, can be performed with the GNNs studied in this work. However,
+this is out of the scope for this manuscript, given resources and time
+needed to generate such training data for SMAs. \par
 
@@
 reported values. The optimized lattice parameters as well as basis
-atoms for all structures and models are provided in appendix~\ref{sec:opt_structures}. \par
+atoms for all structures and models are provided in
+appendix~\ref{sec:opt_structures}. \par
 
@@
 	\includegraphics[options]{figures/NiTi_Equilibrium_Density.png}
-	\caption{Equilibrium Density for \ce{NiTi} structures with different models.}
+	\caption{Equilibrium Density for \ce{NiTi} structures with different
+		models.}
 \end{figure}
 
-To asses the relative phase stability for a given composition is
-to compare the cohesive energy, which is defined as,
+To asses the relative phase stability for a given composition is to
+compare the cohesive energy, which is defined as,
 
@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_B2_ModelsPhononband structure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B2_ModelsPhononband
+			structure.png}
 		\caption{ Ground-state, no isotropic strain, Phonon bandstructure
@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_B19_ModelsPhononband structure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B19_ModelsPhononband
+			structure.png}
 		\caption{ Ground-state phonon bandstructure of B19 phase for all
@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_B19P_ModelsPhononband structure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B19P_ModelsPhononband
+			structure.png}
 		\caption{ Ground-state phonon bandstructure of
@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_BCO_ModelsPhononband structure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_BCO_ModelsPhononband
+			structure.png}
 		\caption{ Ground-state phonon bandstructure of BCO phase for all
@@
 	\begin{centering}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_Mutter_B2_StrainsPhononband structure.png}
-		\vspace{1mm}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_Zhong_B2_StrainsPhononband structure.png}
-		\vspace{1mm}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_Ko_B2_StrainsPhononband structure.png}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_Mutter_B2_StrainsPhononband
+			structure.png} \vspace{1mm}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_Zhong_B2_StrainsPhononband
+			structure.png} \vspace{1mm}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_Ko_B2_StrainsPhononband
+			structure.png}
 		\caption{ Phonon dispersion of B2 structure for the (top) EAM
@@
 	\begin{centering}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_M3GNet_B2_StrainsPhononband structure.png}
-		\vspace{1mm}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_CHGNet_B2_StrainsPhononband structure.png}
-		\vspace{1mm}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_MACE_B2_StrainsPhononband structure.png}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_M3GNet_B2_StrainsPhononband
+			structure.png} \vspace{1mm}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_CHGNet_B2_StrainsPhononband
+			structure.png} \vspace{1mm}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_MACE_B2_StrainsPhononband
+			structure.png}
 		\caption{ Phonon dispersion of B2 structure for the (top) M3GNet,

### 2024-02-04 src/tex/ms.tex
Auctex in emacs was formatting w/ line spits, causing figure reference issue.

@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_B2_ModelsPhononband
-			structure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B2_ModelsPhononbandstructure.png}
 		\caption{ Ground-state, no isotropic strain, Phonon bandstructure
@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_B19_ModelsPhononband
-			structure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B19_ModelsPhononbandstructure.png}
 		\caption{ Ground-state phonon bandstructure of B19 phase for all
@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_B19P_ModelsPhononband
-			structure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B19P_ModelsPhononbandstructure.png}
 		\caption{ Ground-state phonon bandstructure of
@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_BCO_ModelsPhononband
-			structure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_BCO_ModelsPhononbandstructure.png}
 		\caption{ Ground-state phonon bandstructure of BCO phase for all
@@
 	\begin{centering}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_Mutter_B2_StrainsPhononband
-			structure.png} \vspace{1mm}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_Zhong_B2_StrainsPhononband
-			structure.png} \vspace{1mm}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_Ko_B2_StrainsPhononband
-			structure.png}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_Mutter_B2_StrainsPhononbandstructure.png} \vspace{1mm}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_Zhong_B2_StrainsPhononbandstructure.png} \vspace{1mm}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_Ko_B2_StrainsPhononbandstructure.png}
 		\caption{ Phonon dispersion of B2 structure for the (top) EAM

### 2024-02-04 src/tex/ms.tex
Dummy pay attention to file names. Does showyourwork make this easier to track.

@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_B2_ModelsPhononbandstructure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B2_ModelsPhononBandstructure.png}
 		\caption{ Ground-state, no isotropic strain, Phonon bandstructure
@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_B19_ModelsPhononbandstructure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B19_ModelsPhononBandstructure.png}
 		\caption{ Ground-state phonon bandstructure of B19 phase for all
@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_B19P_ModelsPhononbandstructure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B19P_ModelsPhononBandstructure.png}
 		\caption{ Ground-state phonon bandstructure of
@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_BCO_ModelsPhononbandstructure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_BCO_ModelsPhononBandstructure.png}
 		\caption{ Ground-state phonon bandstructure of BCO phase for all
@@
 		\includegraphics[width=0.5\linewidth]{figures/NiTi_Mutter_B2_StrainsPhononbandstructure.png} \vspace{1mm}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_Zhong_B2_StrainsPhononbandstructure.png} \vspace{1mm}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_Ko_B2_StrainsPhononbandstructure.png}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_Zhong_B2_StrainsPhononBandstructure.png} \vspace{1mm}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_Ko_B2_StrainsPhononBandstructure.png}
 		\caption{ Phonon dispersion of B2 structure for the (top) EAM
@@
 	\begin{centering}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_M3GNet_B2_StrainsPhononband
-			structure.png} \vspace{1mm}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_CHGNet_B2_StrainsPhononband
-			structure.png} \vspace{1mm}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_MACE_B2_StrainsPhononband
-			structure.png}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_M3GNet_B2_StrainsPhononBandstructure.png} \vspace{1mm}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_CHGNet_B2_StrainsPhononBandstructure.png} \vspace{1mm}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_MACE_B2_StrainsPhononBandstructure.png}
 		\caption{ Phonon dispersion of B2 structure for the (top) M3GNet,

### 2024-02-04 src/tex/ms.tex
So annoying that the ms.tex, Snakefile, and showyourwork file need output fig name?.
Ahh I'm so impatient with naming, make sure to check figure namesgit add NiTi.Snakefile src/tex/ms.tex !

@@
 	\begin{centering}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_Mutter_B2_StrainsPhononbandstructure.png} \vspace{1mm}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_Zhong_B2_StrainsPhononBandstructure.png} \vspace{1mm}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_Ko_B2_StrainsPhononBandstructure.png}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_Mutter_B2_StrainsPhononBandstructures.png} \vspace{1mm}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_Zhong_B2_StrainsPhononBandstructures.png} \vspace{1mm}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_Ko_B2_StrainsPhononBandstructures.png}
 		\caption{ Phonon dispersion of B2 structure for the (top) EAM
@@
 	\begin{centering}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_M3GNet_B2_StrainsPhononBandstructure.png} \vspace{1mm}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_CHGNet_B2_StrainsPhononBandstructure.png} \vspace{1mm}
-		\includegraphics[width=0.5\linewidth]{figures/NiTi_MACE_B2_StrainsPhononBandstructure.png}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_M3GNet_B2_StrainsPhononBandstructures.png} \vspace{1mm}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_CHGNet_B2_StrainsPhononBandstructures.png} \vspace{1mm}
+		\includegraphics[width=0.5\linewidth]{figures/NiTi_MACE_B2_StrainsPhononBandstructures.png}
 		\caption{ Phonon dispersion of B2 structure for the (top) M3GNet,

### 2024-04-08 src/tex/ms.tex
Fix name of figure again arrgggit add src/tex/ms.tex
Fix name again :(.
Update plotting. Some new text.

@@
 
-\title{Qualitative Assessesment of Graph Neural Network Potentials for
-	Phonon Properties of Shape Memory Alloys using Graph Neural Network
-	Potentials}
+\title{Suitability of Graph Neural Network for Shape Memory Alloys}
 
@@
 	\centering
-	\includegraphics[options]{figures/NiTi_Equilibrium_Density.png}
+	\includegraphics[width=\linewidth]{figures/NiTi_Equilibrium_Density.png}
 	\caption{Equilibrium Density for \ce{NiTi} structures with different
@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_B2_ModelsPhononBandstructure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B2_ModelsPhononBandstructures.png}
 		\caption{ Ground-state, no isotropic strain, Phonon bandstructure
@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_B19_ModelsPhononBandstructure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B19_ModelsPhononBandstructures.png}
 		\caption{ Ground-state phonon bandstructure of B19 phase for all
@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_B19P_ModelsPhononBandstructure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_B19P_ModelsPhononBandstructures.png}
 		\caption{ Ground-state phonon bandstructure of
@@
 	\begin{centering}
-		\includegraphics[width=\linewidth]{figures/NiTi_BCO_ModelsPhononBandstructure.png}
+		\includegraphics[width=\linewidth]{figures/NiTi_BCO_ModelsPhononBandstructures.png}
 		\caption{ Ground-state phonon bandstructure of BCO phase for all

### 2024-04-08 src/scripts/PlotCohesiveEnergy.py

@@ -54,7 +54,9 @@ def plot_cohesive_energy(dbname, chemsys, output_filename):
 
     sorder = STRUCTURE_ORDER[chemsys]
     default_order = max(sorder.values())
-    unique_structure_types = set(entry.structure_name for entry in db.select())
+    unique_structure_types = set(
+        entry.structure_name for entry in db.select(chemsys=chemsys)
+    )
     structure_order = sorted(
         unique_structure_types, key=lambda x: sorder.get(x, default_order)
     )
@@ -62,6 +64,7 @@ def plot_cohesive_energy(dbname, chemsys, output_filename):
     # unique_structure_types = set(
     #     entry.structure_name for entry in db.select(chemsys=chemsys)
     # )
+    print(structure_order)
 
     # Data containers
     structure_names = []

### 2024-04-08 src/scripts/PlotEOS.py

@@ -23,7 +23,7 @@ def plot_eos(
     figsize=(7.24, 7.24),
     dpi=600,
     xlim=(0.75, 1.25),
-    ignore_list=["B32", "R_Phase"],
+    ignore_list=["B32", "L21P", "R_Phase"],
     color_palette="muted",
     n_colors=10,
     wspace=0.3,

### 2024-05-22 src/tex/ms.tex
First Major Refactor of Draft Preprint (#1)
Not the build failed by intention as I cannot run all the simulations on Github actions.

## Refactors
* Rename the EAM potential files.

* Refactor minimize, EOS, elastic, etc.

* Refactor function names. Core should work!

* Refactored Gruneisen code and coh. energy plot

* MEAM by Ko et al. potential files setup, not validated but working.

* Calculation scripts good. Some plotting good. Added draft MD simulation script. New Calculator.

* Fairly complete refactor. Working Process.

* New scripts, potential, rework Snakemake, move folders, added m

@@
   postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}, }
+
+\usepackage{pdflscape}
+
+  % Customize footer to show only "preprint"
+\makeatletter
+\def\ps@pprintTitle{%
+     \let\@oddhead\@empty
+     \let\@evenhead\@empty
+     \def\@oddfoot{\hfill\footnotesize\itshape Preprint, Incomplete Draft\hfill}%
+     \let\@evenfoot\@oddfoot}
+\makeatother
 
@@
 	Shape Memory Alloys (SMAs) such as NiTi are pivotal in vibrational
-	dampening,robotic arms, prothestic hands, and mechanical vales as a
+	dampening,robotic arms, prosthetic hands, and mechanical vales as a
 	result of their unique displacive phase transformation
@@
 	constants. However, the MACE potential shows significant promise as
-	an exploritory potential for designing and characterizing more
+	an exploratory potential for designing and characterizing more
 	complex SMAs. Our findings suggest GNN potentials as a promising
@@
 point. Futhermore, one may choose to fined new representations via
-unsurpervised learning. \par
+unsupervised learning. \par
 
-The subsequent step once the featureization process is completed is to
+The subsequent step once the featurization process is completed is to
 determine a suitable neural network (NN) architecture which can learn
@@
 transferability. Futhermore, the training process was typically
-restricted to a limited number of elemental constieunts of the target
+restricted to a limited number of elemental constituents of the target
 material system. As a consequence, for each chemical system and phases
@@
 between the approaches are predominately in atomic and bond features,
-nearest-neighbor graph, and specific NN archetcture. \par
+nearest-neighbor graph, and specific NN architecture. \par
 
-The training of these early GNN utlized large density functional
+The training of these early GNN utilized large density functional
 theory calculation repositories and mostly tuned the learning cost
@@
 interatomic potential suitable for static and dynamic atomistic
-calculations of multicomponents and multiphases. If succesfull,
+calculations of multi-components and multi-phases. If successful,
 identifying and designing new potential SMA materials using GNN
@@
 assessment of pre-trained GNN interatomic potentials that are promoted
-as being transferable across compositions and phases. We specfically
+as being transferable across compositions and phases. We specifically
 focus on the predictions for equations of state, phonon bandstructure,
@@
 To promote reproducible research and align with findable, accessible,
-interoper-able and reusable (FAIR) principles\cite{Walsh2024}, we
+interoperable and reusable (FAIR) principles\cite{Walsh2024}, we
 utilize \showyourwork python workflow\cite{Luger2021} and several open
@@
 number be provided. The initial unit cell parameters and basis
-corrdinates for these structures have been taken from the materials
+coordinates for these structures have been taken from the materials
 project \cite{Jain2013} and several references
@@
 \begin{figure}[ht!]
-	\label{fig:opt_density}
+	\script{PlotDensity.py}
 	\centering
@@
 		models.}
+	\label{fig:opt_density}
 \end{figure}
@@
 the isolated atom energies, so these terms are zero and neglected,
-giving $E_{c} = \frac{E_{\text{bulk}}}{N}$. In fig.~\ref{fig:ecoh} the
+giving $E_{c} = \frac{E_{\text{bulk}}}{N}$. In fig.~\ref{fig:ecoh_niti} the
 bar chart for different phases and interatomic potentials models are
@@
 consistent with the entry for The MP and calculation details can be
-found in ref.~\cite{MP--571}. This consistency is resonable
+found in ref.~\cite{MP--571}. This consistency is reasonable
 considering all pretrained GNN models used in this work have been
@@
 \begin{figure}[!htp]
-	\script{GruneisenParameters.py} \centering
+	\script{GruneisenParameters.py}
+	\centering
 	\includegraphics[width=\columnwidth]{figures/Plot_NiTi_M_ModeGruneisen.png}
-	\label{fig:modegruneisen}
 	\caption{The sign-log of the \textit{M}-mode Gr\"{u}neisen
@@
 	interatomic potentials.}
+	\label{fig:modegruneisen}
 \end{figure}
@@
 			isotropic strains. Similar transcparency is used as done in
-			\ref{fig:mutter_zhong_phonon_b2}.  }
+			\ref{fig:mutter_zhong_ko_phonon_b2}.  }
 		\label{fig:gnn_phonon_b2}
@@
 \label{sec:appx_reproduce}
-
-TODO
 

### 2024-05-22 src/scripts/AppendixBZ.py

@@ -1,0 +1,42 @@
+import matplotlib.pyplot as plt
+import paths
+from Structures import *
+
+structure_names = ["B2", "B19", "B19P", "BCO"]
+fmap = {
+    "B2": NiTi_B2_Structure,
+    "B19": NiTi_B19_Structure,
+    "B19P": NiTi_B19P_Structure,
+    "BCO": NiTi_BCO_Structure,
+}
+
+for name in structure_names:
+    structure = fmap[name]()
+    structure.info["name"] = name
+    spg = structure.info["spacegroup"].no
+    path, _ = get_bandpath(structure, npoints=1)
+    bz_diagram = f"{name}_BrillouinZonePointsSampled.png"
+    bz_plot = path.plot(vectors=True)
+    bz_plot.figure.savefig(paths.figures / bz_diagram)
+    special_points = path.special_points
+
+    # Write table of special points
+    latex_table = "\\begin{table}[h]\n"
+    latex_table += "\\centering\n"
+    latex_table += "\\begin{tabular}{|c c|}\n"
+    latex_table += "\\hline\n"
+    latex_table += "Point & Coordinates \\\\"
+    latex_table += "\\hline\n"
+
+    for point, coords in special_points.items():
+        coords_str = ", ".join(map(str, coords))
+        latex_table += f"{point} & {coords_str} \\\\\n"
+
+    latex_table += "\\hline\n"
+    latex_table += "\\end{tabular}\n"
+    latex_table += f"\\caption{{Special points sampled in the irreducible Brillouin Zone for space group {spg}}}\n"
+    latex_table += "\\end{table}"
+
+    sp_table = f"{name}_SpecialSymmetryPointsBZ.tex"
+    with open(paths.output / sp_table, "w") as file:
+        file.write(latex_table)

### 2024-05-22 src/scripts/AppendixPlotBZ.py

@@ -1,45 +1,0 @@
-import paths
-
-import matplotlib.pyplot as plt
-
-from NiTiStructures import *
-
-
-structure_names = ["B2", "B19", "B19P", "BCO"]
-fmap = {
-    "B2": NiTi_B2_Structure,
-    "B19": NiTi_B19_Structure,
-    "B19P": NiTi_B19P_Structure,
-    "BCO": NiTi_BCO_Structure,
-}
-
-for name in structure_names:
-    structure = fmap[name]()
-    structure.info["name"] = name
-    spg = structure.info["spacegroup"].no
-    path, _ = get_bandpath(structure,npoints=1)
-    bz_diagram = f"{name}_BrillouinZonePointsSampled.png"
-    bz_plot = path.plot(vectors=True)
-    bz_plot.figure.savefig(paths.figures / bz_diagram)
-    special_points = path.special_points
-
-    # Write table of special points
-    latex_table = "\\begin{table}[h]\n"
-    latex_table += "\\centering\n"
-    latex_table += "\\begin{tabular}{|c c|}\n"
-    latex_table += "\\hline\n"
-    latex_table += "Point & Coordinates \\\\"
-    latex_table += "\\hline\n"
-
-    for point, coords in special_points.items():
-        coords_str = ", ".join(map(str, coords))
-        latex_table += f"{point} & {coords_str} \\\\\n"
-
-    latex_table += "\\hline\n"
-    latex_table += "\\end{tabular}\n"
-    latex_table += f"\\caption{{Special points sampled in the irreducible Brillouin Zone for space group {spg}}}\n"
-    latex_table += "\\end{table}"
-
-    sp_table = f"{name}_SpecialSymmetryPointsBZ.tex"
-    with open(paths.output / sp_table, "w") as file:
-        file.write(latex_table)

### 2024-05-22 src/scripts/CalculateElastic.py

@@ -1,91 +1,133 @@
-import paths
-
-from ase.db import connect
-import ase.units as units
-from ase.spacegroup import crystal
-
-
-#from parcalc import ParCalculate
-from elastic import get_elementary_deformations
-from elastic.elastic import get_cij_order
-from elastic import get_elastic_tensor
-
-from ase.calculators.lammpslib import LAMMPSlib
-
-from queue import Empty
-
-from multiprocessing import Process, Queue
-
-import time
-import os
-import tempfile
-import shutil
+import csv
 from copy import deepcopy
-from subprocess import check_output
-from contextlib import contextmanager
+
+import ase.units as units
+import paths
+from ase.db import connect
+from ase.spacegroup import crystal
+from elastic import get_elastic_tensor, get_elementary_deformations
+from elastic.elastic import get_cij_order
 
 
-def SerialCalculate(systems,calc,cleanup=True,prefix="Calc_"):
+def serial_calculate(systems, calculator):
+    """
+    Create the list of systems with a calculator for elastic constant calculations. This is done in serial.
 
-    if type(systems) != type([]) :
-        sysl=[systems]
-    else :
-        sysl=systems
-    basedir=os.getcwd()
-    res=[]
-    for n,s in enumerate(sysl):
-        if isinstance(calc, LAMMPSlib):
-            s.set_calculator(calc)
-        else:
-            s.set_calculator(deepcopy(calc))
-        s.get_calculator().block=False
-        place=tempfile.mkdtemp(prefix=prefix, dir=basedir)
-        os.chdir(place)
-        s.get_calculator().working_dir=place
-        #print("Start at :", place)
+    Args:
+        systems (Union[list(ASE.Atoms),ASE.Atoms]): atomic structures that contain deformations
+        calculator (ASE.calculators): Internal ASE calculator instance
+
+    Returns:
+        list(systems)
+    """
+
+    if type(systems) != type([]):
+        sysl = [systems]
+    else:
+        sysl = systems
+
+    res = []
+    for n, s in enumerate(sysl):
+        s.set_calculator(calculator)
         s.get_potential_energy()
-        os.chdir(basedir)
-        time.sleep(0.2)
-        res.append([n,s])
-        print("Workers started:", len(sysl))
+        res.append([n, s])
 
-    return [r for ns,s in enumerate(sysl) for nr,r in res if nr==ns]
+    return [r for ns, s in enumerate(sysl) for nr, r in res if nr == ns]
[…]

### 2024-05-22 src/scripts/CalculateGruneisen.py

@@ -1,134 +1,0 @@
-import paths
-import numpy as np
-import numpy.polynomial.polynomial as poly
-from ase.db import connect
-
-
-def FindIndex(data, label):
-    """
-        locates the index for the q-point value that is closest to the
-    point of interest.
-    """
-    if label not in data["bandstructure"]["labels"]:
-        return "Label not found"
-
-    label_index = data["bandstructure"]["labels"].index(label)
-    label_loc = data["bandstructure"]["label_locs"][label_index]
-    q_points = data["bandstructure"]["q_points"]
-    closest_index = min(
-        range(len(q_points)), key=lambda i: abs(q_points[i] - label_loc)
-    )
-
-    return closest_index
-
-
-def ModeGruneisen(phonons, q, n=3):
-    """
-    Calculate the specific mode Gruneisen parameter via a polynomial fit.
-    """
-
-    volumes = []
-    ef_list = []
-    for k in phonons.keys():
-        volumes.append(phonons[k]["volume"])
-        mode_index = FindIndex(phonons[k], q)
-        mode_bands = []
-        for band in phonons[k]["bandstructure"]["band_index"].values():
-            mode_bands.append(band[mode_index])
-        ef_list.append(mode_bands)
-    ef = np.array(ef_list)
-
-    v0 = phonons["0.00"]["volume"]
-    domega_dv0 = np.zeros(ef.shape[1])
-    for i in range(ef.shape[1]):
-        coefs = poly.polyfit(volumes, ef[:, i], n)
-        derivative = poly.polyder(coefs)
-        # Evaluate the derivative at the equilibrium volume
-        domega_dv0[i] = poly.polyval(v0, derivative)
-
-    return domega_dv0
-
-
-def Gruneisen(phonons, n=3):
-    """
-    Calculate the Gruneisen parameters via a polynomial fit.
-    """
-
-    volumes = []
-    ef_list = []
-    for k in phonons.keys():
-        volumes.append(phonons[k]["volume"])
-        bands = []
-        for band in phonons[k]["bandstructure"]["band_index"].values():
-            bands.append(band)
-        ef_list.append(bands)
-    ef = np.array(ef_list)
-
-    v0 = phonons["0.00"]["volume"]
-    domega_dv0 = np.zeros((ef.shape[1], ef.shape[2]))
-    for i in range(ef.shape[1]):
-        for j in range(ef.shape[2]):
-            # Fit a polynomial to the i-th mode
-            coefs = poly.polyfit(volumes, ef[:, i, j], n)
[…]

### 2024-05-22 src/scripts/CalculatePhonons.py

@@ -1,160 +1,156 @@
-import argparse
-import os
 import numpy as np
-import matplotlib.pyplot as plt
 import paths
-
+from ase import Atoms
+from ase.db import connect
 from ase.phonons import Phonons
-from ase.db.core import Database, connect
-from ase.spacegroup import get_spacegroup
-from ase.spacegroup import crystal
-from ase.io.trajectory import Trajectory
-from ase.io import write
-from ase.constraints import StrainFilter, UnitCellFilter
-from ase.spacegroup.symmetrize import FixSymmetry, check_symmetry
-
-from NiTiStructures import *
-from ConfigsUtils import *
+from ase.spacegroup import crystal, get_spacegroup
+from ase.spectrum.band_structure import BandStructure
+from Config import get_phonon_config
+from PlotPhonons import plot_default_phonons
+from Structures import get_bandpath
 
 EV_to_THz = 241.799050402293e0
 
 
-def PlotBandstructure(bs, dos, info, ymaxlim=9.0, of="."):
-    """
-    Plots a single phonon band structure and DOS.
-    """
-    fig = plt.figure(1, figsize=(7, 4))
-    ax = fig.add_axes([0.12, 0.07, 0.67, 0.85])
-    dosax = fig.add_axes([0.8, 0.07, 0.17, 0.85])
+class PhononsFromFile:
+    """Class for reading saved results generated and reprocessing phonon results.
 
-    xticks = bs.get_labels()[1].tolist()
-    xlabels = bs.get_labels()[2]
+    Notes:
 
-    for xt in xticks:
-        ax.axvline(x=xt, color="lightgray")
-    ax.axhline(y=0.0, color="black")
+        - This is not a child class of :class:`~ase.phonons.Phonons` because
+          inheriting all methods including from base
+          :class:`~ase.phonons.Displacement` is not desired, as they would be
+          invalid calls. Therefore, only suitable methods are exposed. For
+          example, see :meth:`~ase.phonons.Phonons.band_structure`.
 
-    q_points = bs.get_labels()[0]
-    energies_THz = EV_to_THz * bs.energies
-    for n in range(len(energies_THz[0, 1, :])):
-        omega = energies_THz[0, :, n]
-        ax.plot(q_points, omega, "k-", lw=2)
+    Example:
 
-    ax.set_title(
-        f"Volume: {info['vol']:.3f}Å^3\
-        Strain: {info['strain']:.2f}% \
[…]

### 2024-05-22 src/scripts/CalculationRunner.py

@@ -1,0 +1,243 @@
+import argparse
+import os
+import warnings
+
+import numpy as np
+import paths
+from ase.db import connect
+from CalculateElastic import calculate_elastic_constants
+from CalculateEOS import get_min_structure_and_calculate_eos
+from CalculatePhonons import calculate_phonons
+from Calculators import get_ase_calculator
+from Structures import get_structure
+
+# Surpress warning from dependencies for which I can do nothing about.
+warnings.filterwarnings("ignore", category=UserWarning, module="ase.spacegroup")
+warnings.filterwarnings(
+    "ignore", category=UserWarning, module="dgl.backend.pytorch.tensor"
+)
+warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_cuda'.*")
+warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_cudnn'.*")
+warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_mps'.*")
+warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_mkldnn'.*")
+
+
+def min_eos_calc(dbname, structure, potential, strain_info):
+    # NOTE: Because this adds a new entry and its hard to get Snakemake to handle
+    # checking if the entry exist directly. We check here and then skip if the model
+    # and structure have already been optimized
+    chemsys = structure.info["chemsys"]
+    structure_name = structure.info["structure_name"]
+    entry_existed = False
+    try:
+        entry = connect(dbname).get(
+            chemsys=chemsys,
+            model_name=potential[0],
+            structure_name=structure_name,
+            relaxed=True,
+        )
+        print(
+            f"""
+            -----------------------------------------------------------
+            EOS already exist for {structure.info['chemsys']} {structure.info['structure_name']} using {potential[0]}
+            -----------------------------------------------------------
+            """
+        )
+        entry_existed = True
+
+    except:
+        print(
+            f"""
+            -----------------------------------------------------------
+            Started EOS Calculation of {structure.info['chemsys']} {structure.info['structure_name']} using {potential[0]}
+            -----------------------------------------------------------
+            """
+        )
+        get_min_structure_and_calculate_eos(
+            structure, potential, dbname=dbname, strain=np.linspace(*strain_info)
+        )
+        print(
+            f"""
+            ------------------------------------------------------------
[…]

### 2024-05-22 src/scripts/Calculators.py

@@ -1,53 +1,211 @@
-import paths
 import numpy as np
+import paths
 from ase.calculators.eam import EAM
+
+# TODO: For M3GNet and CHGNet is stress_weight=1.0*GPa or stress_weight=1.0?
+from ase.units import GPa
+
 
 # No longer needed using conda LAMMPS, which provides the needed stress.
 class CustomEAM(EAM):
-   '''
-   Override EAM to zero out stress for ASE Database support
-   '''
-   def __init__(self, **kwargs):
-       super().__init__(**kwargs)
-   
-   def get_stress(self, atoms=None, **kwargs):
-       return np.zeros(6)
+    """
+    Override EAM to zero out stress for ASE Database support
+    """
 
-def GetCalculator(args):
-    if args.model == 'Mutter':
-       potential_file = str(paths.static / "NiTi_mutter.eam.fs")
-       #asecalc = CustomEAM(potential= str(paths.static / "NiTi_mutter.eam.fs"))
-       from ase.calculators.lammpslib import LAMMPSlib 
-       cmds = ["pair_style eam/fs",f"pair_coeff * *  {potential_file} Ni Ti"]
-       amds=["thermo_style custom etotal lx ly lz vol pxx pyy pzz pxy pxz pyz press"]
-       asecalc = LAMMPSlib(lmpcmds=cmds,amendments=amds,keep_alive=True)
-       #from ase.calculators.lammpsrun import LAMMPS
-       #parameters = {'pair_style': 'eam/fs','pair_coeff': [f'* * {potential_file} Ni Ti']}
-       #files = [potential_file]
-       #asecalc = LAMMPS(parameters=parameters,files=files)
-    elif args.model == 'Zhong':
-       potential_file = str(paths.static / "NiTiZhong.eam.fs")
-       #asecalc = CustomEAM(potential= str(paths.static / "NiTiZhong.eam.fs"))
-       from ase.calculators.lammpslib import LAMMPSlib
-       cmds = ["pair_style eam/fs",f"pair_coeff * *  {potential_file} Ni Ti"]
-       amds=["thermo_style custom etotal lx ly lz vol pxx pyy pzz pxy pxz pyz press"]
-       asecalc = LAMMPSlib(lmpcmds=cmds,amendments=amds,keep_alive=True)
-    elif args.model == "M3GNet":
-       import matgl
-       from matgl.ext.ase import M3GNetCalculator
-       m3gnet_params = matgl.load_model("M3GNet-MP-2021.2.8-PES")
[…]

### 2024-05-22 src/scripts/Config.py

@@ -1,0 +1,73 @@
+def get_phonon_config(structure_name, potential):
+    """Settings for the phonon calculations
+
+    - supercell
+    - atomic displacement (angstrom)
+    - q-points along band path
+
+    Settings have been manually calibrated.
+    """
+    # Supercell, displacement (ang.), band path points
+    configurations = {
+        "B2": {
+            "Mutter": ((8, 8, 8), 0.0935, 200),
+            "Zhong": ((8, 8, 8), 0.0935, 200),
+            "Ko": ((8, 8, 8), 0.01, 200),
+            "Kavousi": ((8, 8, 8), 0.01, 200),
+            "Kim": ((8, 8, 8), 0.01, 200),
+            "M3GNet": ((8, 8, 8), 0.010, 200),
+            "CHGNet": ((8, 8, 8), 0.030, 200),
+            "MACE": ((8, 8, 8), 0.010, 200),
+            "ALIGNN": ((6, 6, 6), 0.050, 200),
+            "DeepMD": ((8, 8, 8), 0.03, 200),
+            "GPAW": ((2, 2, 2), 0.03, 200),
+        },
+        "B19": {
+            "Mutter": ((7, 9, 7), 0.05, 200),
+            "Zhong": ((7, 9, 7), 0.05, 200),
+            "Ko": ((7, 9, 7), 0.03, 200),
+            "Kavousi": ((7, 9, 7), 0.03, 200),
+            "Kim": ((7, 9, 7), 0.03, 200),
+            "M3GNet": ((7, 9, 7), 0.05, 200),
+            "CHGNet": ((6, 8, 6), 0.05, 200),
+            "MACE": ((7, 9, 7), 0.05, 200),
+            "ALIGNN": ((5, 7, 5), 0.05, 200),
+            "DeepMD": ((7, 9, 7), 0.05, 200),
+        },
+        "B19P": {
+            "Mutter": ((8, 6, 4), 0.05, 200),
+            "Zhong": ((8, 6, 4), 0.05, 200),
+            "Ko": ((8, 6, 4), 0.01, 200),
+            "Kavousi": ((8, 6, 4), 0.01, 200),
+            "Kim": ((8, 6, 4), 0.01, 200),
+            "M3GNet": ((8, 6, 4), 0.05, 200),
+            "CHGNet": ((8, 6, 4), 0.05, 200),
+            "MACE": ((8, 6, 4), 0.05, 200),
[…]

### 2024-05-22 src/scripts/ConfigsUtils.py

@@ -1,62 +1,0 @@
-ORDER = {
-    "Mutter": 1,
-    "Zhong": 2,
-    "M3GNet": 3,
-    "CHGNet": 4,
-    "MACE": 5,
-    "ALIGNN": 6,
-}
-
-
-SPACEGROUP_MAP = {"B2" : 221,
-                  "B19P" : 11,
-                  "B19" : 11,
-                  "BCO" : 63,
-                  }
-
-def GetPhononCalcConfig(structure_name, potential):
-    """
-    Settings have been manually calibrated.
-    """
-    # Supercell, displacement (ang.), band path points
-    configurations = {
-        "B2": {
-            "Mutter": ((8, 8, 8), 0.0935, 200),
-            "Zhong": ((8, 8, 8), 0.0935, 200),
-            "M3GNet": ((8, 8, 8), 0.260, 200),
-            "CHGNet": ((8, 8, 8), 0.030, 200),
-            "MACE": ((8, 8, 8), 0.010, 200),
-            "ALIGNN": ((6, 6, 6), 0.050, 200),
-        },
-        "B19": {
-            "Mutter": ((7, 9, 7), 0.05, 200),
-            "Zhong": ((7, 9, 7), 0.05, 200),
-            "M3GNet": ((7, 9, 7), 0.05, 200),
-            "CHGNet": ((7, 9, 7), 0.05, 200),
-            "MACE": ((7, 9, 7), 0.05, 200),
-            "ALIGNN": ((5, 7, 5), 0.05, 200),
-        },
-        "B19P": {
-            "Mutter": ((8, 6, 4), 0.05, 200),
-            "Zhong": ((8, 6, 4), 0.05, 200),
-            "M3GNet": ((8, 6, 4), 0.05, 200),
-            "CHGNet": ((8, 6, 4), 0.05, 200),
-            "MACE": ((8, 6, 4), 0.05, 200),
-            "ALIGNN": ((7, 5, 3), 0.05, 200),
-        },
-        "BCO": {
-            "Mutter": ((8, 4, 6), 0.05, 200),
-            "Zhong": ((8, 4, 6), 0.05, 200),
-            "M3GNet": ((8, 4, 6), 0.05, 200),
-            "CHGNet": ((8, 4, 6), 0.05, 200),
-            "MACE": ((8, 4, 6), 0.05, 200),
-            "ALIGNN": ((5, 2, 3), 0.05, 200),
-        },
-    }
-
-    # Validate structure_name and potential_key
[…]

### 2024-05-22 src/scripts/Generate.py

@@ -1,151 +1,0 @@
-import warnings
-import argparse
-import os
-import paths
-
-from NiTiStructures import *
-from Calculators import *
-
-from CalculateEOS import *
-
-from CalculateElastic import *
-from CalculatePhonons import *
-
-# Surpress warning from dependencies for which I can do nothing about.
-warnings.filterwarnings("ignore", category=UserWarning, module="ase.spacegroup")
-warnings.filterwarnings(
-    "ignore", category=UserWarning, module="dgl.backend.pytorch.tensor"
-)
-warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_cuda'.*")
-warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_cudnn'.*")
-warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_mps'.*")
-warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_mkldnn'.*")
-
-
-def main():
-    """
-    Provides equation of state calculations of NiTi. Be mindful that correct python environment is
-    activated for selected model (i.e., ASE Calculator)
-    """
-    parser = argparse.ArgumentParser(
-        description="""Equation of State Calculations of NiTi.
-    Be mindful that the correct python environment is activated for the model (i.e., ASE Calculator) you
-    intend to use!!"""
-    )
-    parser.add_argument(
-        "--dbname", type=str, default="NiTi_EOS.json", help="ASE Database file name"
-    )
-    parser.add_argument("--dbfolder", type=str, default=".", help="ASE Database folder")
-    parser.add_argument(
-        "--model",
-        choices=["Mutter", "Zhong", "M3GNet", "CHGNet", "MACE", "ALIGNN"],
-        default="Mutter",
-        help="Choose ASECalculator Model",
-    )
-    parser.add_argument("--structure", default="B2")
-    parser.add_argument(
-        "--n",
-        type=int,
-        default=40,
-        help="Number of configurations in strain scan of EOS",
-    )
-    parser.add_argument(
-        "--nph",
-        type=int,
-        default=13,
-        help="Number of configurations in strain scan of Phonons",
-    )
-    parser.add_argument(
-        "--min_strain_eos",
-        type=float,
-        default=-0.10,
-        help="Minimum strain for eos calcs, fractional",
-    )
-    parser.add_argument(
-        "--max_strain_eos",
-        type=float,
-        default=0.10,
[…]

### 2024-05-22 src/scripts/GenerateElasticTable.py

@@ -1,0 +1,107 @@
+import sys
+
+import paths
+from ase.db import connect
+from TableConfig import ORDER, SPACEGROUP_MAP
+
+
+def generate_elastic_constants_table(dbname, chemsys, output_filename):
+    db = connect(dbname)
+
+    sorder = ORDER[chemsys]["structure"]
+    default_order = max(sorder.values())
+    unique_structure_types = set(
+        entry.structure_name for entry in db.select(chemsys=chemsys)
+    )
+    structure_order = sorted(
+        unique_structure_types, key=lambda x: sorder.get(x, default_order)
+    )
+
+    with open(output_filename, "w") as file:
+        file.write("\\begin{landscape}\n")
+        file.write("\\setlength{\\tabcolsep}{6pt}\n")
+        file.write("\\begin{longtable}{|p{0.25\\linewidth}|p{0.75\\linewidth}|}\n")
+        file.write(
+            "\\caption{Elastic constants for "
+            + chemsys
+            + ".} \\label{tab:elastic_"
+            + chemsys
+            + "}\\\\\n"
+        )
+        file.write("\\hline\n")
+        file.write("Structure (\#) & Properties \\\\\n")
+        file.write("\\hline\n")
+        file.write("\\endfirsthead\n")
+        file.write(
+            "\\caption[]{Elastic constants for " + chemsys + ". (Continued)}\\\\\n"
+        )
+        file.write("\\hline\n")
+        file.write("Structure & Properties \\\\\n")
+        file.write("\\endhead\n")
+        file.write("\\hline\n")
+        file.write("\\endfoot\n")
+
+        for structure_type in structure_order:
+            structure_name = structure_type.replace("_", "-")
+            spg_num = SPACEGROUP_MAP[structure_name]
+            file.write(f"{structure_name} ({spg_num}) & ")
+
+            entries = db.select(chemsys=chemsys, structure_name=structure_type)
+
+            # Filter and sort based on models defined in ORDER
+            sentries = sorted(
+                [
+                    entry
+                    for entry in entries
+                    if entry.model_name in ORDER[chemsys]["model"]
+                ],
+                key=lambda entry: ORDER[chemsys]["model"][entry.model_name],
+            )
+
+            file.write("\\begin{tabularx}{\\linewidth}{l l l l l l l l l l l l l l}\n")
+            file.write(
[…]

### 2024-05-22 src/scripts/GenerateEquilibriumTable.py

@@ -1,0 +1,144 @@
+import re
+import sys
+from functools import reduce
+from math import gcd
+
+import numpy as np
+import paths
+from ase.db import connect
+from ase.spacegroup import Spacegroup, get_basis, get_spacegroup
+from TableConfig import ORDER, SPACEGROUP_MAP
+
+
+def get_element_basis_list(structure, spacegroup, tol=1e-3):
+    """Provide the basis for a crystal structure.
+
+    Attempts to find the basis atoms for a given crystal structure.
+
+    STILL IN TESTING.
+
+    Arguments:
+       - structure (ASE.Atoms): The crystal structure.
+       - spacegroup (int): The spacegroup number
+       - tol (float,optional): The tolerance used for determining if sites are the same.
+
+    returns:
+         np.array: The basis atoms
+
+    """
+    sg = Spacegroup(spacegroup)
+    rotations = sg.rotations
+    translations = sg.translations
+
+    unique_sites = []
+    unique_elements = []
+
+    for atom in structure:
+        pos = atom.scaled_position
+        is_unique = True
+
+        # Apply symmetry operations
+        for rot, trans in zip(rotations, translations):
+            sym_pos = np.dot(rot, pos) + trans
+            sym_pos -= np.floor(
+                sym_pos
+            )  # Wrap around using periodic boundary conditions
+
+            # Check if the position is already in the unique sites within a tolerance
+            for usite in unique_sites:
+                if np.allclose(sym_pos, usite, atol=tol):
+                    is_unique = False
+                    break
+
+            if not is_unique:
+                break
+
+        if is_unique:
+            unique_sites.append(pos)
+            unique_elements.append(atom.symbol)
+
+    return list(zip(unique_elements, unique_sites))
+
+
+def generate_structure_table(dbname, chemsys, output_filename):
+    db = connect(dbname)
+
+    sorder = ORDER[chemsys]["structure"]
+    default_order = max(sorder.values())
+    unique_structure_types = set(
+        entry.structure_name for entry in db.select(chemsys=chemsys)
+    )
+    structure_order = sorted(
+        unique_structure_types, key=lambda x: sorder.get(x, default_order)
+    )
+
+    with open(output_filename, "w") as file:
+        file.write("\\begin{longtable}{|l|c|}\n")
+        file.write(
[…]

### 2024-05-22 src/scripts/GenerateMGruneisenTable.py

@@ -1,60 +1,0 @@
-import sys
-from ase.db import connect
-import paths
-from CalculateGruneisen import *
-from ConfigsUtils import *
-
-
-def GenerateMModeGruneisen(dbname, output_filename):
-    """
-    Only calculate at the M point, which for  B2 phase for NiTi is where dynamical instability occurs.
-    """
-    db = connect(dbname)
-    unique_structure_types = set(entry.structure_name for entry in db.select())
-
-    with open(output_filename, "w") as file:
-        file.write("\\begin{table}[h]\n")
-        file.write("\\centering\n")
-
-        # Begin tabular environment
-        file.write("\\begin{tabular}{|c|c|}\n")
-        file.write("\\hline\n")
-        file.write("Structure & \\textbf{M}-Mode Gr\\\"{u}neisen Parameters \\\\\n")
-        file.write("\\hline\n")
-
-        for structure_type in unique_structure_types:
-            if structure_type != "B2":
-                continue
-            structure_name = structure_type.replace("_", "-")
-            file.write(f"{structure_name} & ")
-            file.write("\\begin{tabular}{c c c c c c c}\n")  # Inner table header
-            file.write("Model & TA$_1$ & TA$_2$ & LA & TO$_1$ & TO$_2$ & LO \\\\\n")
-            file.write("\\hline\n")
-            entries = db.select(structure_name=structure_type)
-            sentries = sorted(entries, key=lambda entry: ORDER[entry.model_name])
-            for entry in sentries:
-                model_name =  entry.get('model_name', 'NA')
-                if model_name == "ALIGNN":
-                    if structure_name != "B2":
-                        UserWarning(f"Skipping: {structure_name} {model_name}")
-                        continue
-
-                _, gamma = CalculateModeGruneisen(entry)
-                file.write(
-                    f"{model_name} & {gamma[0]:.2f} & {gamma[1]:.2f} & {gamma[2]:.2f} & {gamma[3]:.2f} & {gamma[4]:.2f} & {gamma[5]:.2f} \\\\\n"
-                )
-
-            file.write("\\hline\n")
-            file.write("\\end{tabular}\\\\\n")
-            file.write("\\hline\n")
-        file.write("\\end{tabular}\n")
-        file.write(
-            "\\caption{Mode Gruneisen parameters for different NiTi structures.}\n"
-        )
-        file.write("\\label{tab:mode_gruneisen_niti}\n")
[…]

### 2024-05-22 src/scripts/GenerateNiTiEquilTable.py

@@ -1,155 +1,0 @@
-import sys
-import re
-import numpy as np
-from math import gcd
-from functools import reduce
-import paths
-
-
-from ase.db import connect
-from ase.spacegroup import get_basis
-from ase.spacegroup import Spacegroup
-from ase.spacegroup import Spacegroup, get_spacegroup
-
-from ConfigsUtils import *
-
-def GetElementBasisList(structure, spacegroup, tol=1e-3):
-    """
-    NOT CONFIRMED
-    Provides the basis/wykcoff sites with element labels
-    """
-    sg = Spacegroup(spacegroup)
-    rotations = sg.rotations
-    translations = sg.translations
-
-    unique_sites = []
-    unique_elements = []
-
-    for atom in structure:
-        pos = atom.scaled_position
-        is_unique = True
-
-        # Apply symmetry operations
-        for rot, trans in zip(rotations, translations):
-            sym_pos = np.dot(rot, pos) + trans
-            sym_pos -= np.floor(sym_pos)  # Wrap around using periodic boundary conditions
-
-            # Check if the position is already in the unique sites within a tolerance
-            for usite in unique_sites:
-                if np.allclose(sym_pos, usite, atol=tol):
-                    is_unique = False
-                    break
-
-            if not is_unique:
-                break
-
-        if is_unique:
-            unique_sites.append(pos)
-            unique_elements.append(atom.symbol)
-
-    return list(zip(unique_elements, unique_sites))
-
-def GenerateStructureTable(dbname, output_filename):
-    db = connect(dbname)
-    
-    unique_structure_types = set(entry.structure_name for entry in db.select())
-    
-    with open(output_filename, 'w') as file:
-        file.write("\\begin{longtable}{|l|c|}\n")
-        file.write("\\caption{Equilibrium structures for NiTi.} \\label{tab:equil_niti}\\\\\n")
-        file.write("\\hline\n")
-        file.write("Structure & Unit Cell \\\\\n")
-        file.write("\\hline\n")
-        file.write("\\endfirsthead\n")
-        file.write("\\caption[]{Equilibrium structures for NiTi. (Continued)}\\\\\n")
-        file.write("\\hline\n")
-        file.write("Structure\n(Spacegroup \#) & Unit Cell \\\\\n")
-        file.write("\\endhead\n")
-        file.write("\\hline\n")
-        file.write("\\endfoot\n")
[…]

### 2024-05-22 src/scripts/GruneisenParameters.py

@@ -1,0 +1,326 @@
+import sys
+
+# TODO: Move to ConfigPlots and use for all.
+import matplotlib.pyplot as plt
+import numpy as np
+import numpy.polynomial.polynomial as poly
+import paths
+import scienceplots
+
+plt.style.use(["science", "scatter", "no-latex"])
+
+from ase.db import connect
+from TableConfig import ORDER
+
+
+# Function to apply the signed log transformation
+def log_transform(val):
+    return np.sign(val) * np.log10(abs(val) + 1)
+
+
+def find_index(data, label):
+    """
+    locates the index for the q-point value that is closest to the point of interest.
+
+    Arguments:
+        data (dict): Phonon data
+        label (str): Symmetry point label
+
+    Returns:
+        int: index of the closest matched occurance.
+    """
+    if label not in data["bandstructure"]["labels"]:
+        return "Label not found"
+
+    label_index = data["bandstructure"]["labels"].index(label)
+    label_loc = data["bandstructure"]["label_locs"][label_index]
+    q_points = data["bandstructure"]["q_points"]
+    closest_index = min(
+        range(len(q_points)), key=lambda i: abs(q_points[i] - label_loc)
+    )
+
+    return closest_index
+
+
+def gruneisen(phonons, n=3):
+    """Calculates the derivative of the frequency with respect to volume for all q-points.
+
+    This function computes the derivative for the Gruneisen parameter. It fits a polynomial of degree `n`
+    to the frequency-vs-volume data for each mode at each wave vector and then evaluates the derivative
+    of this polynomial at the equilibrium volume.
+
+    Args:
+        phonons (dict): A dictionary with keys being the strain values (as strings, e.g., "0.00") and values
+            being dictionaries containing 'volume' and 'bandstructure' data.
+        n (int, optional): The degree of the polynomial to fit to the frequency-vs-volume data. Defaults to 3.
+
+    Returns:
+        np.ndarray: A 2D array where each element [i, j] is the derivative of the frequency of the i-th mode
+        at the j-th wave vector with respect to the volume, evaluated at the equilibrium volume.
+
+    """
+
+    volumes = []
+    ef_list = []
+    for k in phonons.keys():
+        volumes.append(phonons[k]["volume"])
+        bands = []
[…]

### 2024-05-22 src/scripts/NewASEDatabase.py

@@ -1,20 +1,39 @@
 import sys
+
 import paths
-from ase.db import connect
-from SetMetadata import set_db_metadata
-
-dbname = sys.argv[1]
-db = connect(paths.data / dbname, append=False)
-
-#Need to write something otherwise it doesn't work
 from ase import Atoms
 from ase.calculators.emt import EMT
-h2 = Atoms('H2', [(0, 0, 0), (0, 0, 0.7)])
-h2.calc = EMT()
-h2.get_total_energy()
-h2.get_forces()
-db.write(h2)
-db.delete(ids=[1])
-set_db_metadata(paths.data / dbname)
+from ase.db import connect
 
-exit()
+
+def new_ase_database(dbname):
+    """
+    Create a ASE database with a dummy entry then delete.
+
+    Arguments:
+        dbname (string): full path to ASE database file to be created.
+
+    Returns:
+        None
+
+    Notes:
+        A dummy entry is created then deleted so the index will
+        be shifted by 1.
+
+    """
+    db = connect(paths.data / dbname, append=False)
+
+    # NOTE: Must write something to the connected database
+    # but then delete it.
+    h2 = Atoms("H2", [(0, 0, 0), (0, 0, 0.7)])
+    h2.calc = EMT()
+    h2.get_total_energy()
+    h2.get_forces()
+    db.write(h2)
+    db.delete(ids=[1])
+
+    return None
+
+
+if __name__ == "__main__":
+    new_ase_database(sys.argv[1])

### 2024-05-22 src/scripts/PlotDensity.py

@@ -1,0 +1,98 @@
+import sys
+
+import matplotlib.pyplot as plt
+import paths
+import scienceplots
+from matplotlib.ticker import MaxNLocator
+
+plt.style.use(["science", "scatter", "no-latex"])
+
+from ase.db import connect
+from PlotConfigs import ORDER, STRUCTURE_ORDER
+
+
+def plot_density(dbname, chemsys, output_filename):
+    db = connect(dbname)
+
+    sorder = STRUCTURE_ORDER[chemsys]
+    default_order = max(sorder.values())
+    unique_structure_types = set(
+        entry.structure_name for entry in db.select(chemsys=chemsys)
+    )
+    structure_order = sorted(
+        unique_structure_types, key=lambda x: sorder.get(x, default_order)
+    )
+
+    # Data containers
+    structure_names = []
+    model_names = set()
+    density_data = {}
+
+    # Collecting data
+    for name in structure_order:
+        entries = db.select(chemsys=chemsys, structure_name=name)
+        structure_names.append(name.replace("_", "-"))
+        density_data[name] = {}
+
+        for entry in entries:
+            model_name = entry.get("model_name", "NA")
+            density = entry.get("density", 0)  # Default to 0 if not available
+            if model_name in ORDER[chemsys]:
+                density_data[name][model_name] = density
+                model_names.add(model_name)
+
+    # structure_names = sorted(structure_names)
+    model_names = sorted(model_names, key=lambda x: ORDER[chemsys].get(x, float("inf")))
+
+    # Fancy colors and patterns
+    colors = ["red", "green", "blue", "cyan", "magenta", "yellow", "black"]
+    patterns = ["//", "+", "*", "x", "o", "-", "O", "||", "++"]
+
+    # Plotting
+    fig, ax = plt.subplots()
+    bar_width = 0.125
+    space_between_bars = 0.075  # Adjust this value as needed
+    indices = list(range(len(structure_names)))
+    num_models = len(model_names)
+    group_centers = [
+        1.65 * x + (num_models / 2) * (bar_width + space_between_bars) for x in indices
+    ]
+
+    for i, model_name in enumerate(model_names):
+        offset = [(1.5 * x + i * (bar_width + space_between_bars)) for x in indices]
+        density_values = [
+            density_data[struct].get(model_name, 0) for struct in structure_order
+        ]
[…]

### 2024-05-22 src/scripts/PlotEOS.py

@@ -1,0 +1,154 @@
+import sys
+import warnings
+
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+import numpy as np
+import paths
+import scienceplots
+import seaborn as sns  # For color palette
+from ase.db import connect
+from cycler import cycler
+from matplotlib.ticker import MultipleLocator
+
+plt.style.use(["science", "scatter", "no-latex"])
+
+from PlotConfigs import SUBPLOT_ORDER
+
+
+def plot_eos(
+    dbname,
+    chemsys,
+    output_plot,
+    figsize=(7.24, 7.24),
+    dpi=600,
+    xlim=(0.75, 1.25),
+    ignore_list=["B32", "L21P", "R_Phase"],
+    color_palette="muted",
+    n_colors=10,
+    wspace=0.3,
+    hspace=0.3,
+    linewidth=0.75,
+    linestyle="-",
+    fontsize=5,
+    tick_fontsize=5,
+    legend_fontsize=6,
+    legend_ncol=5,
+    label_notes="EOS_Labels.txt",
+):
+    db = connect(dbname)
+
+    # Select unique models if in SUBPLOT_ORDER and sort
+    plot_models = SUBPLOT_ORDER[chemsys].values()
+
+    calculator_models = sorted(
+        set(
+            row.model_name
+            for row in db.select(chemsys=chemsys)
+            if row.model_name in plot_models
+        ),
+        key=lambda model: list(SUBPLOT_ORDER[chemsys].values()).index(model),
+    )
+    structure_names = set(row.structure_name for row in db.select(chemsys=chemsys))
+
+    # Create a figure with a 2x3 grid of subplots
+    fig, axs = plt.subplots(2, 3)  # , figsize=figsize)
+    axs = axs.ravel()  # Flatten the array of axes for easier indexing
+
+    # Create a color cycle iterator using a slightly modified color palette
+    color_cycle = sns.color_palette(color_palette, n_colors=n_colors)
+
+    label_file_path = paths.figures / label_notes
+    with open(label_file_path, "w") as label_file:
+        for i, calculator_model in enumerate(calculator_models):
+            if calculator_model not in SUBPLOT_ORDER[chemsys].values():
+                warnings.warn(
+                    f"{calculator_model} model results not used/expected", UserWarning
+                )
+
+            ax = axs[i]
+            alphabetic_label = f"({chr(97 + i)})"
[…]

### 2024-05-22 src/scripts/PlotNiTiCohesiveEnergy.py

@@ -1,60 +1,0 @@
-import sys
-import matplotlib.pyplot as plt
-from ase.db import connect
-import paths
-
-def GenerateCohesiveEnergyPlot(dbname, output_filename):
-    db = connect(dbname)
-    unique_structure_types = set(entry.structure_name for entry in db.select())
-
-    # Data containers
-    structure_names = []
-    model_names = set()
-    ecoh_data = {}
-
-    # Collecting data
-    for structure_type in unique_structure_types:
-        entries = db.select(structure_name=structure_type)
-        structure_names.append(structure_type.replace("_", "-"))
-        ecoh_data[structure_type] = {}
-
-        for entry in entries:
-            model_name = entry.get('model_name', 'NA')
-            ecoh = entry.get('ecoh', 0)  # Default to 0 if not available
-            ecoh_data[structure_type][model_name] = ecoh
-            model_names.add(model_name)
-
-    # Sorting for consistent plotting
-    model_names = sorted(model_names)
-    structure_names = sorted(structure_names)
-
-    # Plotting
-    fig, ax = plt.subplots()
-
-    # Width of a bar
-    bar_width = 0.35
-    # Position of bars on the x-axis
-    indices = list(range(len(structure_names)))
-
-    for i, model_name in enumerate(model_names):
-        # Offset each model's bar
-        offset = [(x + i * bar_width) for x in indices]
-        # Extract cohesive energies for each structure for the current model
-        ecoh_values = [ecoh_data[struct].get(model_name, 0) for struct in structure_names]
-        ax.bar(offset, ecoh_values, bar_width, label=model_name)
-
-    # Setting the x-axis labels
-    ax.set_xlabel('Structure Name')
-    ax.set_ylabel('Cohesive Energy (eV/atom)')
-    ax.set_title('Cohesive Energy by Structure and Model')
-    ax.set_xticks([r + bar_width for r in range(len(structure_names))])
-    ax.set_xticklabels(structure_names)
-    ax.legend()
-
-    # Save plot
-    plt.savefig(output_filename)
-
-
-if __name__ == "__main__":
-    output_plot = paths.figures / "NiTi_CohesiveEnergy.png"
-    GenerateCohesiveEnergyPlot(paths.data / sys.argv[1], output_plot)

### 2024-05-22 src/scripts/PlotNoStrainPhonons.py

@@ -1,11 +1,0 @@
-import sys
-import paths
-from PlotPhonons import *
-
-
-models = ["Mutter", "Zhong","M3GNet","CHGNet","MACE","ALIGNN"]
-dbname = paths.data / sys.argv[1]
-structures = sys.argv[2:]
-
-for s in structures:
-    PlotAllModelPhonons(dbname, models, s, strain=0.00)

### 2024-05-22 src/scripts/PlotPhonons.py

@@ -1,9 +1,15 @@
-import paths
-import numpy as np
+import argparse
+import os
+import sys
+
 import matplotlib.pyplot as plt
-from matplotlib.gridspec import GridSpec
+import numpy as np
+import paths
 from ase.db import connect
-from scipy.interpolate import interp1d
+from matplotlib.gridspec import GridSpec
+from PlotConfigs import SUBPLOT_ORDER
+
+EV_to_THz = 241.799050402293e0
 
 
 def format_xticklabels(labels, locs):
@@ -38,15 +44,85 @@ def get_color(value, cmap_name="viridis", vmin=-1.5, vmax=1.5):
     return cmap(norm(value))
 
 
-def PlotPhonons(dbname, model_name, structure_name, num_strains=None):
+def plot_default_phonons(bs, dos, info, of="."):
     """
-    This function plots individual phonon bandstructures for different strains and
-    then all the strains on one plot with the optical branches semi-transparent.
+    Plot results from ASE phonon calculation.
+    """
+    fig = plt.figure(1, figsize=(7, 4))
+    ax = fig.add_axes([0.12, 0.07, 0.67, 0.85])
+    dosax = fig.add_axes([0.8, 0.07, 0.17, 0.85])
 
-    TODO: confirm strain is in bars and thus conver to MPa
+    xticks = bs.get_labels()[1].tolist()
+    xlabels = bs.get_labels()[2]
+
+    for xt in xticks:
+        ax.axvline(x=xt, color="lightgray")
+    ax.axhline(y=0.0, color="black")
+
+    q_points = bs.get_labels()[0]
+    energies_THz = EV_to_THz * bs.energies
+    for n in range(len(energies_THz[0, 1, :])):
+        omega = energies_THz[0, :, n]
+        ax.plot(q_points, omega, "k-", lw=2)
+
+    ax.set_title(
+        f"Volume: {info['vol']:.3f}Å^3\
+        Strain: {info['strain']:.2f}% \
+        Potential: {info['potname']}",
+        fontsize=10,
+    )
+
+    ax.set_xticks(xticks)
+    ax.set_xticklabels(xlabels)
+
+    yminlim, ymaxlim = np.min(energies_THz) * 1.01, np.max(energies_THz) * 1.01
+    ax.set_xlim((0.0, q_points[-1]))
+    ax.set_ylim((yminlim, ymaxlim))
[…]

### 2024-05-22 src/scripts/TableConfig.py

@@ -1,0 +1,39 @@
+SPACEGROUP_MAP = {
+    "B2": 221,
+    "B19P": 11,
+    "B19": 51,
+    "BCO": 63,
+}
+
+ORDER = {
+    "NiTi": {
+        "model": {
+            "Mutter": 1,
+            "Zhong": 2,
+            "Ko": 3,
+            "M3GNet": 4,
+            "CHGNet": 5,
+            "MACE": 6,
+        },
+        "structure": {
+            "B2": 1,
+            "B19P": 2,
+            "B19": 3,
+            "BCO": 4,
+        },
+    },
+    "PtTi": {
+        "model": {
+            "Kim": 1,
+            "M3GNet": 2,
+            "CHGNet": 3,
+            "MACE": 4,
+        },
+        "structure": {
+            "B2": 1,
+            "B19P": 2,
+            "B19": 3,
+            "BCO": 4,
+        },
+    },
+}

### 2024-05-22 src/scripts/Testing.py

@@ -1,151 +1,0 @@
-import warnings
-import argparse
-import os
-import paths
-
-from NiTiStructures import *
-from Calculators import *
-
-from CalculateEOS import *
-
-#from CalculateElastic import *
-from CalculatePhonons import *
-
-# Surpress warning from dependencies for which I can do nothing about.
-warnings.filterwarnings("ignore", category=UserWarning, module="ase.spacegroup")
-warnings.filterwarnings(
-    "ignore", category=UserWarning, module="dgl.backend.pytorch.tensor"
-)
-warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_cuda'.*")
-warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_cudnn'.*")
-warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_mps'.*")
-warnings.filterwarnings("ignore", category=UserWarning, message=".*'has_mkldnn'.*")
-
-
-def main():
-    """
-    Provides equation of state calculations of NiTi. Be mindful that correct python environment is
-    activated for selected model (i.e., ASE Calculator)
-    """
-    parser = argparse.ArgumentParser(
-        description="""Equation of State Calculations of NiTi.
-    Be mindful that the correct python environment is activated for the model (i.e., ASE Calculator) you
-    intend to use!!"""
-    )
-    parser.add_argument(
-        "--dbname", type=str, default="NiTi_EOS.json", help="ASE Database file name"
-    )
-    parser.add_argument("--dbfolder", type=str, default=".", help="ASE Database folder")
-    parser.add_argument(
-        "--model",
-        choices=["Mutter", "Zhong", "M3GNet", "CHGNet", "MACE", "ALIGNN"],
-        default="Mutter",
-        help="Choose ASECalculator Model",
-    )
-    parser.add_argument("--structure", default="B2")
-    parser.add_argument(
-        "--n",
-        type=int,
-        default=40,
-        help="Number of configurations in strain scan of EOS",
-    )
-    parser.add_argument(
-        "--nph",
-        type=int,
-        default=13,
-        help="Number of configurations in strain scan of Phonons",
-    )
-    parser.add_argument(
-        "--min_strain_eos",
-        type=float,
-        default=-0.10,
-        help="Minimum strain for eos calcs, fractional",
-    )
-    parser.add_argument(
-        "--max_strain_eos",
-        type=float,
-        default=0.10,
[…]

### 2024-05-22 src/scripts/VisualizeStructures.py

@@ -1,0 +1,211 @@
+import io
+import sys
+from math import cos, degrees, radians, sin, sqrt
+
+import matplotlib.pyplot as plt
+import paths
+from ase.visualize.plot import plot_atoms
+from ovito.io.ase import ase_to_ovito
+from ovito.modifiers import (
+    DeleteSelectedModifier,
+    ExpressionSelectionModifier,
+    ReplicateModifier,
+)
+from ovito.pipeline import Pipeline, StaticSource
+from ovito.qt_compat import QtCore
+from ovito.vis import (
+    CoordinateTripodOverlay,
+    OpenGLRenderer,
+    OSPRayRenderer,
+    TachyonRenderer,
+    TextLabelOverlay,
+    Viewport,
+)
+from PIL import Image, ImageDraw, ImageFont
+from PySide6.QtCore import QBuffer, QIODevice
+from Structures import get_structure
+
+
+def ovito_visualize_and_save(structures, filename, renderer="OpenGL"):
+    rendered_images = []
+
+    for structure in structures:
+        a, b, c = structure.cell.lengths()
+        alpha, beta, gamma = [radians(a) for a in structure.cell.angles()]
+        # Convert the ASE structure to an OVITO DataCollection
+        data_collection = ase_to_ovito(structure)
+
+        # Create a pipeline and set the data collection as its source
+        pipeline = Pipeline(source=StaticSource(data=data_collection))
+
+        # Replicate the structure
+        n = 3
+        pipeline.modifiers.append(
+            ReplicateModifier(num_x=n, num_y=n - 1, num_z=n, adjust_box=False)
+        )
+
+        # Select and delete atoms outside the original box
+        # expression = (
+        #     "Position.X > {:.4f} || Position.Y > {:.4f} || Position.Z > {:.4f}".format(
+        #         pipeline.source.data.cell[0][0],
+        #         pipeline.source.data.cell[1][1],
+        #         pipeline.source.data.cell[2][2],
+        #     )
+        # )
+        # pipeline.modifiers.append(ExpressionSelectionModifier(expression=expression))
+        # pipeline.modifiers.append(DeleteSelectedModifier())
+
+        pipeline.compute()
+
+        vis_element = pipeline.compute().particles.vis
+        vis_element.scaling = 0.50
+        vis_cell = pipeline.source.data.cell.vis
+        vis_cell.line_width = 0.05
+
[…]

### 2024-05-22 src/scripts/WriteCodeToTex.py

@@ -1,0 +1,46 @@
+import re
+import sys
+
+import paths
+
+
+def extract_function(source_file, function_name):
+    with open(source_file, "r") as file:
+        code = file.read()
+
+    pattern = re.compile(rf"(def {function_name}\(.*?:\n(?:\s+.*\n)+)", re.MULTILINE)
+    match = pattern.search(code)
+
+    if match:
+        return match.group(1)
+    else:
+        raise ValueError(f"Function '{function_name}' not found in the file.")
+
+
+def write_to_tex(output_file, function_code):
+    """write extracted code in LaTeX listing environment"""
+    tex_content = (
+        f"\\begin{{lstlisting}}[language=Python]\n{function_code}\\end{{lstlisting}}\n"
+    )
+
+    with open(output_file, "w") as file:
+        file.write(tex_content)
+
+
+if __name__ == "__main__":
+    if len(sys.argv) != 4:
+        print(
+            "Usage: python script.py <source_python_file> <function_name> <output_tex_file>"
+        )
+        sys.exit(1)
+
+    source_python_file = paths.scripts / sys.argv[1]
+    function_name = sys.argv[2]
+    output_tex_file = paths.output / sys.argv[3]
+
+    try:
+        python_function_code = extract_function(source_python_file, function_name)
+        write_to_tex(output_tex_file, python_function_code)
+        print(f"Function {function_name} has been written to {output_tex_file}")
+    except ValueError as e:
+        print(e)

### 2024-05-24 src/tex/ms.tex
Fix script reference typo
Update to README and manuscript.

@@
 \usepackage{longtable} \usepackage{tabularx} \usepackage{subcaption}
-\usepackage{listings} \lstset{ basicstyle=\ttfamily,
+\usepackage{listings} \lstset{basicstyle=\ttfamily,
   columns=fullflexible, frame=single, breaklines=true,
-  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space}, }
+  postbreak=\mbox{\textcolor{red}{$\hookrightarrow$}\space},}
 
@@
 	Shape Memory Alloys (SMAs) such as NiTi are pivotal in vibrational
-	dampening,robotic arms, prosthetic hands, and mechanical vales as a
+	dampening, robotic arms, prosthetic hands, and mechanical vales as a
 	result of their unique displacive phase transformation
@@
 functional theory calculations to obtain properties such as equations
-of state, phonon disperisons, and elastic constants, can be done,
+of state, phonon band structures, and elastic constants, can be done,
 however, such calculations are computationally demanding and are
@@
 Recently, there as been a growing effort in the development of machine
-learning potentials. These deviate from traditional physics approaches
+learning potentials. These deviate from traditional based physics approaches
 to interatomic potentials, which have historically been inspired from
-more accurate methods like tight-binding method. In machine learning
-potentials, one describes the atoms and structure using descriptors of
-features, these are typically learned or prescribed. The features can
-be prescribed in isolation or combination of structural,
+more accurate methods like tight-binding method \cite{Finnis2010}. In machine learning
+potentials, one describes the atoms and structure using descriptors (also called
+features and are used interchangeably here on out) which are typically learned or prescribed.  If prescribed, the descriptors can
+be isolated in terms of structure or composition, or a combination of the two. If needed properties can also be used as part of the descriptor set.\
 compositional, and property characteristics of the material data
-point. Futhermore, one may choose to fined new representations via
-unsupervised learning. \par
+point. Futhermore, one may choose to fine new representations via
+unsupervised learning \cite{Choudhary2021}. \par
 
-The subsequent step once the featurization process is completed is to
-determine a suitable neural network (NN) architecture which can learn
-the target property or properties given the feature inputs. In the
-case of atomistic simulations the target is prediction of the
+The subsequent step, once the featurization process is completed, is to
+determine a suitable machine learning model which can learn
+the target property or properties given the descriptor inputs. For large data sets Neural Networks have become the preferred choice. \par
+
+In the case of atomistic simulations the target is prediction of the
 potential energy surface and corresponding gradients (i.e.,
-forces). Two of the most early attempts to do this were the works of
+forces). Two of the most early ML models to do this were the works of
 \citet{Bartok2010} and \citet{Thompson2015}, whereby projection onto
@@
 details to learn new features from which could limit
-transferability. Futhermore, the training process was typically
+transferability. Futhermore, the training process is typically
 restricted to a limited number of elemental constituents of the target
 material system. As a consequence, for each chemical system and phases
-of interest one needs to provide a suitable training dataset. To move
-beyond such limitations, groups began using graph neural network (GNN)
+of interest one needs to provide a suitable dataset to learn from (i.e., training). To move beyond such limitations, groups began using graph neural network (GNN)
 architectures and message passing scheme to represent the atomic
@@
 a generalization\footnote{Typically these are referred to as Universal
-	interatomic potentials or forcefields since they in theory
+	interatomic potentials or force fields since they in theory
 	generalize well to any chemical system or phase.}. The use of GNNs
@@
 interatomic potentials could have tremendous impact on
-technology. Therefore, the primary goal is to provide a qualitative
+technology. \par
+
+The primary goal of this work is to provide a qualitative
 assessment of pre-trained GNN interatomic potentials that are promoted
@@
 \cite{Chen2022,Deng2023,Batatia2022}. We note that the
-\citet{Riebesell2023} provides a summarized performance over these GNN
-using the materials project database, but does not specifically
+\citet{Riebesell2023} provides a summarized performance for all of these GNN
+using the materials project database\cite{Jain2013}, but does not specifically
 address SMA and phonon characteristics. \par
@@
 required by the interatomic potentials we leverage the ability to
-specify the python computing environments using
+specify virtual/containerized python computing environments using
 \href{https://github.com/showyourwork/showyourwork}{\showyourwork}
@@
 shown in fig~\ref{fig:niti_structures} and are adapted for PtTi as
-well given same space group symmetries. The energy and forces on the
+well, given same space group symmetries. The energy and forces on the
 atoms are determined using the ASE calculator interfaces for EAM,
@@
 The EAM and MEAM potential interfaces are achieved through the use of
-LAMMPS python library \cite{Thompson2022} and ASE. We have also
+LAMMPS python library \cite{Thompson2022} and ASE\@. We have also
 implemented calculators for potentials by \cite{Kavousi2019},
@@
 \begin{figure}[ht!]
-	\script{VisualizedStructures.py}
+	\script{VisualizeStructures.py}
 	\begin{centering}
 		\includegraphics[width=\linewidth]{figures/NiTi_VisualizedStructures.png}
-		\caption{ Initial \ce{NiTi} structures (i.e., unrelaxed) used for
+		\caption{Initial \ce{NiTi} structures (i.e., unrelaxed) used for
 			(a) \textit{B2}(221), (b) \textit{B19}(11), (c)
@@
 characteristics of the different phases that can occur for a given SMA
-is particular important. One of the challenges is in performing such
-calculations, which typically employ density functional theory (DFT)
-in order to determine the atomic forces, which in turn are used to
-construct the dynamical matrix. However, classical interatomic
-potentials can alleviate the computational burden. \par
+is particular important. \par
 
@@
 for sampling the irreducible Brillouin zone for the specific
-structures.\cite{Setyawan2010}. We predominately focus on
+structures \cite{Setyawan2010}. We predominately focus on
 high-symmetry points in the IBZ that correspond to the corners and
 along directions known to exhibit phonon instabilities. The IBZ and
-points symmetry point locations can be seen in appendix
-\ref{sec:appx_ibz}. The final phonon dispersion diagrams are plot
+points symmetry point locations can be seen in appendix~\ref{sec:appx_ibz}. The final phonon dispersion diagrams are plot
 along these points. \par
@@
 Similar to the phonon calculations the elastic constants within the
-harmonic approximation the elastic constants can be determined from
+harmonic approximation, can be determined from
 linear-response of strain-stress calculations. Therefore, the
@@
 
-The EAM and MEAM potentials were developed to go beyond pair
+The EAM and MEAM potentials were developed to go beyond simpler pair
 interatomic potentials to treat point and extended defects in metal
 and alloys systems. The motivation was taken from the physical results
-of density functional theory and tight-binding methods, namely to add
-to the pair potentials a energy functional which depends on the local
-electron density which in turn is affected by the atomic
+of density functional theory and tight-binding methods. The approach was to add
+to the pair potentials a energy functional which depends on the local atomic
 environment. These type of interatomic potentials are broadly referred
-to as EAM or MEAM, with the later included direction (i.e. angle)
-dependent interactions. \par
+to as EAM or MEAM, with the later included direction
+dependent interactions (i.e., angular terms). \par
 
@@
 		\includegraphics[width=\linewidth]{figures/NiTi_CohesiveEnergyPlot.png}
-		\caption{ Comparison of predicted \ce{NiTi} cohesive energy for
+		\caption{Comparison of predicted \ce{NiTi} cohesive energy for
 			respective ground-state structure and model. The experimental
@@
 \textit{B19'} structures. The MEAM potential shows a lower cohesive
-energy than all three GNN. Interestingly the difference between the
+energy than all three GNN\@. Interestingly the difference between the
 MEAM and GNN is roughly the same order, $\approx 2.1$ eV, which maybe
@@
 		\includegraphics[width=\linewidth]{figures/PtTi_CohesiveEnergyPlot.png}
-		\caption{ Comparison of predicted \ce{PtTi} cohesive energy for
+		\caption{Comparison of predicted \ce{PtTi} cohesive energy for
 			respective ground-state structure and model.  }
@@
 The MACE potential shown in fig.~\ref{fig:eos}(e) has a very different
-set of EOS curves compard to the M3GNet and CHGNet. Here many of the
-phases exhibit a stablity window, except the \textit{B33} phase, which
+set of EOS curves compared to the M3GNet and CHGNet. Here many of the
+phases exhibit a stability window, except the \textit{B33} phase, which
 was stable in M3GNet and CHGNet. Interestingly, the \text{B19} phase,
@@
 Finally, the ALIGNN potential EOS was generated, but a few issues
-arose during the calculations. The most prominant was the difficulty
+arose during the calculations. The most prominent was the difficulty
 in convergence of the structural optimzation to the specified
-tolerance (see \ref{sec:methods}. Secondly the calculated cohesive
+tolerance (see~\ref{sec:methods}). Secondly the calculated cohesive
 energies differs from the UIP-GNN by a factor $\approx 2$ and by $1$
@@
 		\includegraphics[width=\linewidth]{figures/NiTi_EquationOfStates.png}
-		\caption{ The equation of state curves for different \ce{NiTi}
+		\caption{The equation of state curves for different \ce{NiTi}
 			structures for (a) Mutter (EAM) (b) Zhong (EAM) (c) Ko
@@
 
+In the context of NiTi alloys, phonon instabilities play a crucial role in understanding the phase stability and transformations.
+The  phonon band structures for \textit{B2} phase, as shown in Fig.~\ref{fig:allmodelsB2}, reveal that the Mutter EAM potential predicts no dynamic instability at 0K.
+In contrast the Zhong EAM potential predicts onset of dynamic instability.
+In contrast, the Ko and MACE models exhibit phonon softening along the $\Gamma-M$ direction, aligning well with DFT results \cite{Wu2022}.
+The M3GNet and CHGNet models predict also predict mode instability but qualitatively differs significantly.
+This suggests that Ko and MACE models more accurately capture the intrinsic instability of the \textit{B2} phase, crucial for predicting the martensitic transformation to \textit{B19}$\prime$ or \textit{BCO} phases.
+
+For the \textit{B19} phase, Fig.~\ref{fig:allmodelsB19}, the phonon band structures indicate instability in the $\Gamma-Y/B$ directions, consistent with DFT predictions.
+The Ko, CHGNet, and MACE models display similar trends, albeit with slight discrepancies.
+Furthermore, the \textit{B19}$\prime$ and \textit{BCO} phases, illustrated in Fig.~\ref{fig:allmodels_B19P} and Fig.~\ref{fig:allmodelsBCO}, respectively, highlight the stability predicted by DFT\@.
+Notably, only the Mutter, Zhong, and Ko models predict no mode instability for \textit{B19}$\prime$ and \textit{BCO}, while the others indicate potential mode instability.
+This reinforces the necessity here for fine-tuned GNN  potentials that can accurately capture the complex phonon dynamics and phase stability in NiTi alloys.
+
 \begin{figure}[!htp]
@@
 		\includegraphics[width=\linewidth]{figures/NiTi_B2_ModelsPhononBandstructures.png}
-		\caption{ Ground-state, no isotropic strain, Phonon bandstructure
-			of B2 phase for all models. The Mutter and Zhong EAM potentials
+		\caption{Ground-state, no isotropic strain, Phonon bandstructure
+			of B2 phase for all models. The Mutter (a) and Zhong (b) EAM potentials
 			differ in that Mutter predicts no dynamic instability of the B2
-			phase at 0K.  }
-		\label{fig:allmodels_b2}
+			phase at 0K.
+The results for Ko (c), M3GNet (d), CHGNet (e), and MACE (f) all display the instability from $\Gamma-M$.
+The Ko (c) and MACE (f) models appear to agree best with DFT results \cite{Wu2022}.}
+		\label{fig:allmodelsB2}
 	\end{centering}
@@
 		\includegraphics[width=\linewidth]{figures/NiTi_B19_ModelsPhononBandstructures.png}
-		\caption{ Ground-state phonon bandstructure of B19 phase for all
-			models.  }
-		\label{fig:allmodels_B19P}
+		\caption{Ground-state phonon bandstructure of B19 phase for all
+			models.
+Plots (a-f) follow same ordering as in fig.~\ref{fig:allmodelsB2}.The DFT results in ref.~\cite{Wu2022} indicate instability in $\Gamma-Y/B$ directions.
+In (c,e-f) the results are similar to the DFT but slight differ.}
+		\label{fig:allmodelsB19}
 	\end{centering}
@@
 		\includegraphics[width=\linewidth]{figures/NiTi_B19P_ModelsPhononBandstructures.png}
-		\caption{ Ground-state phonon bandstructure of
-			\textit{B19}$^\prime$ phase for all models.  }
-		\label{fig:allmodels_B19P}
+		\caption{Ground-state phonon bandstructure of
+			\textit{B19}$^\prime$ phase for all models.
+Plots (a-f) follow same ordering as in fig.~\ref{fig:allmodelsB2}.
+As indicated by DFT this should be a potential stable structure however only (a-c) show no mode instability.}
+		\label{fig:allmodelsB19P}
 	\end{centering}
@@
 		\includegraphics[width=\linewidth]{figures/NiTi_BCO_ModelsPhononBandstructures.png}
-		\caption{ Ground-state phonon bandstructure of BCO phase for all
-			models.  }
-		\label{fig:allmodels_B19P}
+		\caption{round-state phonon bandstructure of BCO phase for all
+			models. Plots (a-f) follow same ordering as in fig.~\ref{fig:allmodelsB2}. As indicated by DFT this should be a potential stable structure however only (a-c) show no mode instability.}
+		\label{fig:allmodelsBCO}
 	\end{centering}
@@
 
-The B2 phase of NiTi exhibits notable phonon instabilities,
+The \textit{B2} phase of NiTi exhibits notable phonon instabilities,
 particularly at the \textit{M} point in the Brillouin zone. These
@@
 		\includegraphics[width=0.5\linewidth]{figures/NiTi_Ko_B2_StrainsPhononBandstructures.png}
-		\caption{ Phonon dispersion of B2 structure for the (top) EAM
+		\caption{Phonon dispersion of B2 structure for the (top) EAM
 			Mutter, (middle) EAM Zhong, and (bottom) 2NN-MEAM Ko potentials
@@
 			to tensile strain. Note that the strain percent shown is with
-			respect to isotropic changes (i.e. volume).  }
+			respect to isotropic changes (i.e., volume).}
 		\label{fig:mutter_zhong_ko_phonon_b2}
@@
 		\includegraphics[width=0.5\linewidth]{figures/NiTi_MACE_B2_StrainsPhononBandstructures.png}
-		\caption{ Phonon dispersion of B2 structure for the (top) M3GNet,
+		\caption{Phonon dispersion of B2 structure for the (top) M3GNet,
 			(middle) CHGNet, and (bottom) MACE potentials for different
-			isotropic strains. Similar transcparency is used as done in
-			\ref{fig:mutter_zhong_ko_phonon_b2}.  }
+			isotropic strains. Similar transcparency is used as done in~\ref{fig:mutter_zhong_ko_phonon_b2}.}
 		\label{fig:gnn_phonon_b2}
@@
 \label{sec:conclusion}
-This word aimed to provide an assement of the of GNN univeral
-interatomic potentials on the ability to capture the stable
-ground-state phases of binary and ternary SMA. In addition, the
-dynamic stability of these SMA is crucially evaluated to understand
-the suitability in observing the onset of displacive phase
-transformations via the softening of TA branch modes. In the case of
+This work provides a first step in the assessment of pre-trained universal GNN
+interatomic potentials on ability to model SMA materials. In addition, the
+dynamic stability of these SMA is crucially evaluated through phonon band structures to understand the suitability in observing the onset of displacive phase
+transformations via the softening modes. In the case of
 NiTi we find that all the GNN potentials predict either the Monoclinic
 or BCO phases as stable over the B2 structure. However, not all agree
-with the recent works by (cite) which have indicated that it is the
-BCO phase which is the ground-state structure.
-
-We observe that the M3GNet and CHGnet capture the M-point instability
-of the B2 phase, however, no stablizing isotropic strain is
-observed. With regard to the MACE, very good agrement is found with
-the phonon bandstructure and that from ab-initio methods. The ALIGNN
-potential was found not to be suitable of NiTi.
+with relatively recent works \cite{Haskins2016,Wu2022} that indicate \textit{BCO} or \textit{B33} phase is the ground-state structure.
 
@@
 high-symmetry points in the IBZ are shown in the tables below. In
-Ref. xxx the specific implementation of the band path sampling and
+Ref.~xxx the specific implementation of the band path sampling and
 definitions are provided.

### 2024-05-24 src/tex/ms.tex
Tweaks

@@
 \begin{abstract}
-	Shape Memory Alloys (SMAs) such as NiTi are pivotal in vibrational
-	dampening, robotic arms, prosthetic hands, and mechanical vales as a
-	result of their unique displacive phase transformation
-	properties. However, assessing their phonon characteristics, which
-	provide insight into the displacive phase transformation, by using
-	conventional Density Functional Theory (DFT) methods is
-	computationally challenging. In this paper we assess the use of
-	recent Graph Neural Network (GNN) potentials as a viable
-	alternative. These GNN potentials have leveraged large DFT-computed
-	databases for training and provide a good approximation to the
-	potential energy surfaces which enable inference on energies and
-	forces in lattice and atomic dynamic calculations. We compare GNN
-	potentials with the Embedded Atom Method (EAM) and Modified EAM
-	style potentials in predicting phonon properties of NiTi and
-	PtTi. We find that in general the EAM/MEAM potentials are slightly
-	more performant than the GNN in that they appear to more accurately
-	describe equation of state, phonon dispersions, and elastic
-	constants. However, the MACE potential shows significant promise as
-	an exploratory potential for designing and characterizing more
-	complex SMAs. Our findings suggest GNN potentials as a promising
-	initial avenue for advancing SMA research, in particular ternary and
-	quaternary systems, upon fine-tuning of pre-trained GNN models with
-	more relevant DFT data (i.e., force-constants) to permit better
-	outcomes.
+	Shape Memory Alloys (SMAs), such as NiTi, are crucial in vibration dampening, robotic arms, prosthetic hands, and mechanical valves due to their unique displacive phase transformation properties.
+However, assessing their phonon characteristics using conventional Density Functional Theory (DFT) methods is computationally expensive.
+This paper evaluates the use of recent Graph Neural Network (GNN) potentials as a viable alternative.
+These GNN potentials, trained on extensive DFT-computed databases, approximate potential energy surfaces, enabling inference on energies and forces in lattice and atomic dynamics calculations.
+We compare GNN potentials with the Embedded Atom Method (EAM) and Modified EAM (MEAM) potentials in predicting phonon properties of NiTi and PtTi.
+Generally, EAM/MEAM potentials appear to slightly outperform GNNs phonon, providing more accurate descriptions of the equation of state, phonon dispersions, and elastic constants.
+However, the MACE potential shows significant promise as an exploratory tool for designing and characterizing more complex SMAs.
+Our findings suggest GNN potentials as a promising initial avenue for advancing SMA research, particularly for ternary and quaternary systems, with improved outcomes possible by fine-tuning pre-trained GNN models with more relevant DFT data (e.g., force-constants).
 \end{abstract}
@@
 \section*{Acknowledgements}
+The author would like to thank Prof.Venkateswara Manga at the University of Arizona for fruitful discussions.
 
